### 2024-05-21 paper/figures/numerical-rank.tex
finish first draft of paper

\begin{tikzpicture}
    \fill[lightblue] (-2, -0.15) rectangle (1.5, 0.15);
    \draw[thick, ->] (-5, 0) to (5, 0);
    \fill[darkblue] (-4, 0) circle (0.075) node[above] {$\lambda_{(7)}$};
    \fill[darkblue] (-3, 0) circle (0.075) node[above] {$\lambda_{(5)}$};
    \fill[darkblue] (-2.3, 0) circle (0.075) node[above] {$\lambda_{(4)}$};
    \fill[darkblue] (-1, 0) circle (0.075) node[above] {$\lambda_{(2)}$};
    \fill[darkblue] (0.3, 0) circle (0.075) node[above] {$\lambda_{(1)}$};
    \fill[darkblue] (1, 0) circle (0.075) node[above] {$\lambda_{(3)}$};
    \fill[darkblue] (2.8, 0) circle (0.075) node[above] {$\lambda_{(6)}$};
    \fill[darkblue] (3.6, 0) circle (0.075) node[above] {$\lambda_{(8)}$};
    \fill[darkblue] (4.4, 0) circle (0.075) node[above] {$\lambda_{(9)}$};
    \draw[darkblue, ultra thick] (-0.155, 0.15) to (-0.155, -0.15) node[below] {$t$};
    \draw[darkblue, thick] (-2, 0.15) to (-2, -0.15) node[below] {$t - d_{\varepsilon, \sigma}$};
    \draw[darkblue, thick] (1.5, 0.15) to (1.5, -0.15) node[below] {$t + d_{\varepsilon, \sigma}$};
\end{tikzpicture}