### 2024-06-04 paper/main.tex
make corrections after meeting

\documentclass[12pt]{article}

\usepackage{stylesheet}

\title{Randomized trace estimation for parameter-dependent matrices applied to spectral density approximation}

\begin{document}

\maketitle

%\todo{
%\begin{itemize}
%    \item Unify index notation
%    \item Double-check proofs
%    \item Add all references
%    \item Run numerical experiments 
%    \item Write introduction, abstract, and discussion
%    \item ArXiv HTML support (check if packages)
%\end{itemize}
%}

%\todo{
%    \tableofcontents
%}

\todo{

\paragraph{TODO}
\begin{itemize}
    \item Change citation style to \textbf{[1]}, \textbf{[2]}, \dots
    \item Re-run numerical experiments with simplified and improved implementation (test pinvh against lstsq)
\end{itemize}
}

\begin{abstract}
    \todo{Write abstract.}
\end{abstract}

\section{Introduction}
\label{sec:introduction}

In numerous problems across physics, chemistry, engineering, and data science, how the eigenvalues of certain matrices are distributed can be indicative of what properties a system has: In electronic structure calculations eigenvalues represent the energy levels which electrons occupy \cite{lin-2017-randomized-estimation, drabold-1993-maximum-entropy, ducastelle-1970-moments-developments, haydock-1972-electronic-structure}, in neural network optimization they are indicative of the topology of the loss landscape \cite{yao-2020-pyhessian-neural, ghorbani-2019-investigation-neural}, and in graph processing they can uncover hidden graph motifs \cite{huang-2021-density-states}. The distribution of the eigenvalues $\lambda_1, \dots, \lambda_n \in \mathbb{R}$ of a symmetric matrix $\mtx{A} \in \mathbb{R}^{n \times n}$ can be represented by the spectral density $\phi(t) = n^{-1} \sum_{i=1}^{n} \delta(t - \lambda_i)$. Clearly, assembling this expression amounts to computing all eigenvalues of the matrix; an operation which is often prohibitively expensive. Therefore, many techniques have been proposed for approximating $\phi$. For example, by matching its moments with a discrete density function \cite{cohen-steiner-2018-approximating-spectrum, braverman-2022-sublinear-time}, by deriving a quadrature rule with the Lanczos procedure \cite{lin-2016-approximating-spectral, chen-2021-analysis-stochastic}, or by explicit polynomial expansion \cite{weisse-2006-kernel-polynomial, lin-2016-approximating-spectral}. Further, because $\delta$ is hard to approximate and in many applications, a rough estimate of the spectral density is already acceptable, a standard approach is to instead approximate the smooth spectral density
\begin{equation}
    \phi_{\sigma}(t) = \frac{1}{n} \sum_{i=1}^{n} g_{\sigma}(t - \lambda_i) = \frac{1}{n} \Trace(g_{\sigma}(t \mtx{I}_n - \mtx{A}))
\end{equation}
for a smoothing kernel $g_{\sigma}$, typically a Gaussian \cite{lin-2016-approximating-spectral, lin-2017-randomized-estimation} or a Lorentzian \cite{haydock-1972-electronic-structure, lin-2016-approximating-spectral}. This translates the problem to computing the trace of a continuous parameter-dependent matrix $\mtx{B}(t) \equiv g_{\sigma}(t \mtx{I}_n - \mtx{A}) \in \mathbb{R}^{n \times n}$. Directly assembling the matrix function requires diagonalizing the matrix $\mtx{A}$, which again is prohibitively expensive. Therefore, we estimate its trace implicitly by approximating products of the matrix function with vectors. For constant matrices, the most widespread methods for implicit trace estimation are the Girard-Hutchinson estimator \cite{girard-1989-fast-montecarlo, hutchinson-1990-stochastic-estimator} and the variance reduced versions thereof \cite{gambhir-2017-deflation-method, saibaba-2017-randomized-matrixfree, lin-2017-randomized-estimation, meyer-2021-hutch-optimal, persson-2022-improved-variants, epperly-2024-xtrace-making, chen-2023-krylovaware-stochastic}. Some of these estimators have already been successfully applied to parameter-dependent matrices \cite{lin-2017-randomized-estimation}, but they have not yet been analyzed. \todo{A related technique is dynamic trace estimation for sequences $\mtx{B}_1, \dots, \mtx{B}_m$ of similar matrices \cite{woodruff-2022-optimal-query, dharangutte-2021-dynamic-trace}, which achieves guarantees for the estimate $b_i$ of $\Trace(\mtx{B}_i)$ of the form $\mathbb{P}(|\Trace(\mtx{B}_i) - b_i| \geq \varepsilon) \leq \delta$ for $i=1,\dots, m$. A drawback of this type of bound is that when we aim for $\varepsilon$ accuracy for all $i=1,\dots, m$ simultaneously, the probability will grow linearly with $m$ by the union bound. Therefore, instead of discretizing $[a, b]$ with $t_1, \ldots, t_m$ and looking at the error $|\Trace(\mtx{B}(t_i)) - b(t_i)|$ for each $i$, a more natural way to quantify the error is to use $\int_{a}^{b} | \Trace(\mtx{B}(t)) - b(t) |~\mathrm{d}t$, where $b(t)$ is the trace estimate for each $t \in [a, b]$.}

\paragraph{Contribution}
We analyze three well-established randomized trace estimators when they are applied to parameter-dependent matrices. In particular, we reuse the same randomization for each value of the parameter $t$ in the estimators, which makes them scale favorably with the number of parameter evaluations. Subsequently, we propose a simple, fast, robust, and provable procedure in which these estimators are employed for approximating the smooth spectral density $\phi_{\sigma}$ of a matrix.

\paragraph{Reproducibility} This document, along with all included plots and tables, was generated from commit \href{https://github.com/FMatti/Rand-SD/tree/ca391a8}{ca391a8} at \href{https://github.com/FMatti/Rand-TRACE}{github.com/FMatti/Rand-TRACE} on 2024-02-05 at 18:44:29 UTC.

\section{Analysis of trace estimators for parameter-dependent matrices}
\label{sec:analysis}

We consider parameter-dependent matrices of the form
\begin{equation}
    \mtx{B}(t) = \begin{bmatrix}
        b_{11}(t) & b_{12}(t) & \dots & b_{1n}(t) \\
        b_{21}(t) & b_{22}(t) & \dots & b_{2n}(t) \\
        \vdots & \vdots & \ddots & \vdots \\
        b_{n1}(t) & b_{n2}(t) & \dots & b_{nn}(t) \\
    \end{bmatrix} \in \mathbb{R}^{n \times n}
\end{equation}
where $b_{ij}(t)$ are functions depending continuously on the parameter $t$ which takes values in the interval $[a,b]$. The trace of such a matrix is defined as
\begin{equation}
    \Trace(\mtx{B}(t)) = \sum_{i=1}^{n} b_{ii}(t).
    \label{equ:definition-trace}
\end{equation}
However, we assume that we only have access to products of this matrix with vectors for each $t \in [a, b]$, so this definition will not be directly useful for computing the trace.

\paragraph{Girard-Hutchinson estimator} We can estimate the trace with the Girard-Hutchinson estimator \cite{girard-1989-fast-montecarlo,hutchinson-1990-stochastic-estimator}: We take $n_{\mtx{\Psi}}$ stochastically independent standard Gaussian random vectors $\vct{\psi}_1,\dots, \vct{\psi}_{n_{\mtx{\Psi}}} \in \mathbb{R}^{n}$ to form
\begin{equation}
    \Hutch{\mtx{\Psi}}(\mtx{B}(t))
    = \frac{1}{n_{\mtx{\Psi}}} \sum_{j=1}^{n_{\mtx{\Psi}}} \vct{\psi}_j^{\top} \mtx{B}(t) \vct{\psi}_j
    = \frac{1}{n_{\mtx{\Psi}}} \Trace( \mtx{\Psi}^{\top} \mtx{B}(t) \mtx{\Psi})
    \label{equ:hutchinson-trace-estimator}
\end{equation}
where $\mtx{\Psi} = [\vct{\psi}_1 ~ \cdots ~ \vct{\psi}_{n_{\mtx{\Psi}}}] \in \mathbb{R}^{n \times n_{\mtx{\Psi}}}$. Other choices for the distribution of the random vectors are possible, for example by uniformly sampling from $\{-1, +1\}$ or from the $(n-1)$-sphere. However, our theoretical developments only hold in the Gaussian case.

\paragraph{Nyström approximation} Alternatively, the trace of a symmetric matrix whose rank is significantly lower than its size can be approximated well by using a Gaussian sketching matrix $\mtx{\Omega} \in \mathbb{R}^{n \times n_{\mtx{\Omega}}}$ to form the Nyström approximation \cite{gittens-2013-revisiting-nystrom}
\begin{equation}
    \Nystr{\mtx{\Omega}}{\mtx{B}}(t) = (\mtx{B}(t) \mtx{\Omega}) (\mtx{\Omega}^{\top} \mtx{B}(t) \mtx{\Omega})^{\dagger} (\mtx{B}(t) \mtx{\Omega})^{\top}.
    \label{equ:nystrom-approximation}
\end{equation}
Then we can estimate the trace as $\Trace(\Nystr{\mtx{\Omega}}{\mtx{B}}(t))$. Thanks to the invariance of the trace under cyclic permutation of its arguments and the symmetry of the matrix, we may rewrite this estimator as

\begin{equation}
    \Trace(\Nystr{\mtx{\Omega}}{\mtx{B}}(t)) = \Trace( (\mtx{\Omega}^{\top} \mtx{B}(t) \mtx{\Omega})^{\dagger} ( \mtx{\Omega}^{\top} \mtx{B}(t)^2 \mtx{\Omega})).
    \label{equ:nystrom-trace-estimator}
\end{equation}

\paragraph{Nyström++ estimator} Finally, \cite{lin-2017-randomized-estimation} proposes an estimator which corrects for inaccuracies in the Nyström approximation \refequ{equ:nystrom-trace-estimator} by estimating the trace of its residual using the Girard-Hutchinson estimator \refequ{equ:hutchinson-trace-estimator}
\begin{equation}
    \Nystrpp{\mtx{\Omega}}{\mtx{\Psi}}(\mtx{B}(t)) = \Trace(\Nystr{\mtx{\Omega}}{\mtx{B}}(t)) + \Hutch{\mtx{\Psi}}(\mtx{B}(t) - \Nystr{\mtx{\Omega}}{\mtx{B}}(t)).
    \label{equ:nystrompp-trace-estimator}
\end{equation}
This is the parameter-dependent analogue of the Nyström++ estimator \cite{persson-2022-improved-variants}, which is based on the Hutch++ estimator \cite{meyer-2021-hutch-optimal}. %We can interpret this estimator as an interpolation between the trace of the Nyström approximation and the Girard-Hutchinson estimator.

\subsection{Girard-Hutchinson estimator for parameter-dependent matrices}
\label{subsec:hutchinson}

We first consider the estimator \refequ{equ:hutchinson-trace-estimator}. \todo{Introduce $L^1$ norm and Frobenius norm}

\begin{theorem}{Girard-Hutchinson estimator for parameter-dependent matrices}{hutchinson}
    Let $\mtx{B}(t) \in \mathbb{R}^{n \times n}$ be symmetric and continuously depend on $t \in [a, b]$. Then for any $k \in \mathbb{N}$ and $\gamma \geq 1$ it holds with probability at least $1 - \gamma^{-2k}$
    \begin{equation}
        \int_{a}^{b} \left| \Trace(\mtx{B}(t)) - \Hutch{\mtx{\Psi}}(\mtx{B}(t)) \right| ~\mathrm{d}t < 2^4 \gamma k \frac{1}{\sqrt{n_{\mtx{\Psi}}}} \int_{a}^{b} \lVert \mtx{B}(t) \rVert _F  ~\mathrm{d}t.
    \end{equation}
    In particular, if we choose $n_{\mtx{\Psi}} = \mathcal{O}(\varepsilon^{-2} \log(\delta^{-1})^2)$, then with probability at least $1-\delta$ for $\delta \in (0, e^{- \sfrac{1}{2}})$ and any $\varepsilon > 0$ we have $\int_{a}^{b} | \Trace(\mtx{B}(t)) - \Hutch{\mtx{\Psi}}(\mtx{B}(t)) | ~\mathrm{d}t < \varepsilon \int_{a}^{b} \lVert \mtx{B}(t) \rVert _F~\mathrm{d}t$.
\end{theorem}

The idea behind the proof is to bound the higher order moments for fixed parameter values and then use Markov's inequality to transition to the continuous parameter case. We first only consider the $1$-query estimator and subsequently use a diagonal embedding trick to carry the result over to the general case.
\begin{proof} 
    We define the 1-query Girard-Hutchinson estimate 
    \begin{equation}
        r(\mtx{B}(t), \vct{\psi}) = \Trace(\mtx{B}(t)) - \vct{\psi}^{\top} \mtx{B}(t) \vct{\psi}
    \end{equation}
    for a symmetric parameter-dependent matrix $\mtx{B}(t)$ over the real numbers with the Gaussian random vector $\vct{\psi}$.

    First, we consider $r(\mtx{B}(t), \vct{\psi})$ for a fixed $t \in [a,b]$ and therefore temporarily ignore the parameter-dependence. From the proof of \cite[lemma 3]{cortinovis-2022-randomized-trace} we know that $r(\mtx{B}, \vct{\psi})$ is sub-Gamma with parameters $(v, c) = (2 \lVert \mtx{B} \rVert _F^2, 2 \lVert \mtx{B} \rVert _2)$ \todo{introduce spectral norm}. Thus, by \cite[theorem 2.3]{boucheron-2013-basic-inequalities} this implies that for every $k \in \mathbb{N}$
    \begin{align}
        \mathbb{E}\left[ r(\mtx{B}, \vct{\psi})^{2 k} \right]
        &\stackrel{\text{\cite{boucheron-2013-basic-inequalities}}}{\leq} k! \left( 16 \lVert \mtx{B} \rVert _F^2 \right)^k + (2 k)! \left( 8 \lVert \mtx{B} \rVert _2 \right)^{2 k} \notag \\
        &= k! 2^{4 k} \lVert \mtx{B} \rVert _F^{2 k} + (2 k)! 2^{6 k} \lVert \mtx{B} \rVert _2^{2 k}.
    \end{align}
    Since $\lVert \mtx{B} \rVert _2 \leq \lVert \mtx{B} \rVert _F$ and $k! 2^{4 k} + (2 k)! 2^{6 k} \leq \frac{9}{8}(2 k)! 2^{6 k}$ for any $k \in \mathbb{N}$, we can upper bound 
    \begin{equation}
        \mathbb{E}\left[ r(\mtx{B}, \vct{\psi})^{2 k} \right] \leq \frac{9}{8} (2 k)! 2^{6 k} \lVert \mtx{B} \rVert _F^{2k}.
    \end{equation}
    Stirling's approximation \cite{robbins-1955-remark-stirling} bonuds $(2 k)! < 2 \sqrt{\pi k}  e^{\sfrac{1}{24 k}} ( 2 k / e )^{2 k}$. Consequently, the moments $\mathbb{E}^{k}[\cdot] = \left(\mathbb{E}\left[ | \cdot |^{k} \right] \right)^{\sfrac{1}{k}}$ of $r(\mtx{B}, \vct{\psi})$ are limited by
    \begin{equation}
        \mathbb{E}^{2k}\left[ r(\mtx{B}, \vct{\psi}) \right]
        < \left( \frac{9}{4} \sqrt{\pi k} e^{\sfrac{1}{24 k}} \right)^{\sfrac{1}{2k}} \left( \frac{2 k}{e}\right) 2^{3} \lVert \mtx{B} \rVert _F < 2^4 k \lVert \mtx{B} \rVert _F,
        \label{equ:hutchinson-trace-onequery-fixed}
    \end{equation}
    since it can be checked that $(\frac{9}{4} \sqrt{\pi k} e^{\sfrac{1}{24 k}})^{\sfrac{1}{2k}} < e$, due to the monotonous decrease of this expression in $k \in \mathbb{N}$.

    Now we make the transition to the continuous. Minkowski's integral inequality \cite[theorem 202]{hardy-1952-inequalities} allows us to apply \refequ{equ:hutchinson-trace-onequery-fixed} in the continuous setting: 
    \begin{equation}
        \mathbb{E}^{2 k}\left[ \int_{a}^{b} |r(\mtx{B}(t), \vct{\psi})|~\mathrm{d}t  \right]
        \stackrel{\text{\cite{hardy-1952-inequalities}}}{\leq} \int_{a}^{b} \mathbb{E}^{2 k}\left[ r(\mtx{B}(t), \vct{\psi}) \right]~\mathrm{d}t
        < 2^4 k \int_{a}^{b} \lVert \mtx{B}(t) \rVert _F~\mathrm{d}t.
    \end{equation}
    Consequently, by Markov's inequality, for any $k \in \mathbb{N}$ and $\gamma \geq 1$, it holds with probability at least $1 - \gamma^{-2 k}$
    \begin{equation}
        \int_{a}^{b} |r(\mtx{B}(t), \vct{\psi})|~\mathrm{d}t < 2^4 k \gamma \int_{a}^{b} \lVert \mtx{B}(t) \rVert _F~\mathrm{d}t.
        \label{equ:hutchinson-trace-onequery-uniform}
    \end{equation}

    To extend the bound to the $n_{\mtx{\Psi}}$-query quadratic trace estimator, we use a technique from the proof of \cite[theorem 1]{cortinovis-2022-randomized-trace}. Let
    \begin{equation}
        \widetilde{\mtx{B}}(t)
        = \frac{1}{n_{\mtx{\Psi}}} \begin{pmatrix}
            \mtx{B}(t) & & \\
            & \ddots & \\
            & & \mtx{B}(t)
        \end{pmatrix}
        \quad \text{and} \quad
        \widetilde{\vct{\psi}} = \begin{pmatrix}
            \vct{\psi}_1 \\
            \vdots \\
            \vct{\psi}_{n_{\mtx{\Psi}}}
        \end{pmatrix}
    \end{equation}
    where $\vct{\psi}_1, \dots, \vct{\psi}_{n_{\mtx{\Psi}}} \in \mathbb{R}^{n}$ are independent Gaussian random vectors. Then $\lVert \widetilde{\mtx{B}}(t) \rVert _F = \lVert \mtx{B}(t) \rVert _F / \sqrt{n_{\mtx{\Psi}}}$, and
    \begin{equation}
        r(\widetilde{\mtx{B}}(t), \widetilde{\vct{\psi}}) = \Trace(\mtx{B}(t)) - \frac{1}{n_{\mtx{\Psi}}} \sum_{j=1}^{n_{\mtx{\Psi}}} \vct{\psi}_j^{\top} \mtx{B}(t) \vct{\psi}_j.
    \end{equation}
    Hence, by applying \refequ{equ:hutchinson-trace-onequery-uniform} to $r(\widetilde{\mtx{B}}(t), \widetilde{\vct{\psi}})$ we conclude that for any $k \in \mathbb{N}$ and $\gamma \geq 1$ with probability at least $1 - \gamma^{-2 k}$
    \begin{equation}
        \int_{a}^{b} \left| \Trace(\mtx{B}(t)) - \frac{1}{n_{\mtx{\Psi}}} \sum_{j=1}^{n_{\mtx{\Psi}}} \vct{\psi}_j^{\top} \mtx{B}(t) \vct{\psi}_j \right| ~ \mathrm{d}t
        < 2^4 k \gamma \frac{1}{\sqrt{n_{\mtx{\Psi}}}} \int_{a}^{b} \lVert \mtx{B}(t) \rVert _F~\mathrm{d}t.
    \end{equation}

    Setting $\delta = \gamma^{-2 k}$ and choosing $k = \lceil \log(\delta^{-1}) \rceil$, where $\lceil x \rceil$ is the ceiling function which returns the greatest integer larger than or equal to $x$, we get that for all $\delta \in (0, e^{-\sfrac{1}{2}})$
    \begin{equation}
        k \gamma = \lceil \log(\delta^{-1}) \rceil \delta^{-\sfrac{1}{2\lceil \log(\delta^{-1}) \rceil}}
        = \lceil \log(\delta^{-1}) \rceil e^{\sfrac{1}{2}\frac{\log(\delta^{-1})}{\lceil \log(\delta^{-1}) \rceil}}
        \leq 2 \log(\delta^{-1}) e^{\sfrac{1}{2}},
    \end{equation}
    from which follows the second part of the theorem.
    %Consequently, with probability $1 - \delta$ we have
    %\begin{equation}
    %    \int_{a}^{b} \left| \frac{1}{n_{\mtx{\Psi}}} \sum_{i=1}^{n_{\mtx{\Psi}}} \vct{\psi}_i^{\top} \mtx{B}(t) \vct{\psi}_i - \Trace(\mtx{B}(t)) \right| ~ \mathrm{d}t \leq 32 e^{\sfrac{1}{2}} \frac{\log(\delta^{-1})}{\sqrt{n_{\mtx{\Psi}}}} \int_{a}^{b} \lVert \mtx{B}(t) \rVert _F~\mathrm{d}t
    %\end{equation}
\end{proof}

\begin{remark}
    If additionally $\mtx{B}(t)$ is nonzero and symmetric positive semi-definite for all $t \in [a, b]$, then it follows from the proof of \refthm{thm:hutchinson} that with probability at least $1 - \delta$
    \begin{equation}
        \int_{a}^{b} \frac{| \Trace(\mtx{B}(t)) - \Hutch{\mtx{\Psi}}(\mtx{B}(t)) |}{\Trace(\mtx{B}(t))} ~\mathrm{d}t < \varepsilon,
    \end{equation}
    if $n_{\mtx{\Psi}} \geq 2^{10} \varepsilon^{-2} \log(\delta^{-1})^2 (\int_{a}^{b} \mu(t)^{\sfrac{1}{2}}~\mathrm{d}t)^2$ with $\mu(t) = \lVert \mtx{B}(t) \rVert _2 / \Trace(\mtx{B}(t))$. Compared to the constant matrix equivalent \cite[remark 2]{cortinovis-2022-randomized-trace} we mainly observe a larger constant and an additional factor of $\log(\delta^{-1})$.
\end{remark}

The Girard-Hutchinson estimator distinguishes itself for its simplicity and provability for all symmetric matrices. However, the $n_{\mtx{\Psi}} = \mathcal{O}(\varepsilon^{-2})$ requirement implies that if we aim to increase the accuracy of the estimate by one digit, we would need to increase the number of queries by a factor of $100$. The next two estimators improve on this by exploiting the structure of $\mtx{B}(t)$.

\subsection{Nyström approximation for parameter-dependent matrices}
\label{subsec:nystrom}

Next we consider \refequ{equ:nystrom-trace-estimator}. \todo{Motivate Nyström approximation if matrix is low-rank.}

\begin{theorem}{Nyström estimator for parameter-dependent matrices}{nystrom}
    Suppose $\mtx{B}(t) \in \mathbb{R}^{n \times n}$ is continuous and positive semi-definite in $t \in [a, b]$. Let $n_{\mtx{\Omega}} \geq 6, n_{\mtx{\Omega}} - 4 \geq r \geq 2$ be two integers. Then for all $\gamma \geq 1$ with probability at least $1 - \gamma^{-(n_{\mtx{\Omega}} - r)}$
    \begin{equation}
        \int_{a}^{b} \left| \Trace(\mtx{B}(t)) - \Trace(\Nystr{\mtx{\Omega}}{\mtx{B}}(t)) \right| ~\mathrm{d}t
        \leq \gamma^2 (1 + r) \int_{a}^{b} \sum_{i = r+1}^{n} \sigma_i(\mtx{B}(t)) ~\mathrm{d}t.
    \end{equation}
\end{theorem}

\begin{proof}
    Since $f$ is non-negative, $\mtx{B}(t)$ is positive semi-definite for all $t$. By \cite[lemma 2.1]{frangella-2023-randomized-nystrom}, $\mtx{B}(t) - \Nystr{\mtx{\Omega}}{\mtx{B}}(t)$ is also positive semi-definite for all $t$. Therefore, for fixed $t \in [a, b]$
    \begin{equation}
        \left| \Trace(\mtx{B}(t)) - \Trace(\Nystr{\mtx{\Omega}}{\mtx{B}}(t)) \right|
        = \left| \Trace(\mtx{B}(t) - \Nystr{\mtx{\Omega}}{\mtx{B}}(t)) \right|
        = \lVert \mtx{B}(t) - \Nystr{\mtx{\Omega}}{\mtx{B}}(t) \rVert _{\ast},
    \end{equation}
    \todo{introduce nuclear norm}.  From the proof of \cite[corollary 8.2]{tropp-2023-randomized-algorithms} it follows that
    \begin{equation}
        %\lVert f(\mtx{A}, t) - \widehat{f}(\mtx{A}, t) \rVert _{\ast} = \lVert (\mtx{I}_n - \mtx{\Pi}_{f(\mtx{A}, t)^{\sfrac{1}{2}} \mtx{\Omega}}) f(\mtx{A}, t)^{\sfrac{1}{2}} \rVert _2^2 \leq \lVert (\mtx{I}_n - \mtx{\Pi}_{f(\mtx{A}, t)^{\sfrac{1}{2}} \mtx{\Omega}}) f(\mtx{A}, t)^{\sfrac{1}{2}} \rVert _F^2
        \lVert \mtx{B}(t) - \Nystr{\mtx{\Omega}}{\mtx{B}}(t) \rVert _{\ast} = \lVert (\mtx{I}_n - \mtx{\Pi}_{\mtx{B}(t)^{\sfrac{1}{2}} \mtx{\Omega}}) \mtx{B}(t)^{\sfrac{1}{2}} \rVert _F^2.
    \end{equation}
    Hence, by \cite[theorem 9]{kressner-2023-randomized-lowrank}, with $n_{\mtx{\Omega}} = p + r$, the result follows by noting $\sigma_i(\mtx{B}(t)^{\sfrac{1}{2}})^{2} = \sigma_i(\mtx{B}(t))$.
\end{proof}

\subsection{Nyström++ estimator for parameter-dependent matrices}
\label{subsec:nystrom-pp}

\todo{Motivate and explain proof technique (to mimic Hutch++)}

For the coming pages, we employ the following notation. Let
\begin{equation}
    \mtx{B}(t) 
    = \mtx{U}(t) \mtx{\Lambda}(t) \mtx{U}(t)^{\top} 
\end{equation}
be the eigenvalue decomposition of $\mtx{B}(t)$ at a fixed $t$, i.e. $\mtx{\Lambda}(t) = \operatorname{diag}(\lambda_1(t), \dots, \lambda_n(t))$ and $\mtx{U}(t)^{\top} \mtx{U}(t) = \mtx{I}_n$. We define the matrix partitions
\begin{equation}
    \rule[\dimexpr-2ex-\ht\strutbox]{0pt}{\dimexpr2ex+4ex+\baselineskip}
    \mtx{U}(t) = \begin{bmatrix}
        \smash{\underbrace{\mtx{U}_1(t)}_{n \times k}} & \smash{\underbrace{\mtx{U}_2(t)}_{n \times (n-k)}}
    \end{bmatrix}
    \quad \text{and} \quad
    \mtx{\Lambda}(t) =
    \begin{bmatrix}
        \smash{\overbrace{\mtx{\Lambda}_1(t)}^{k \times k}} & \\ & \smash{\underbrace{\mtx{\Lambda}_2(t)}_{(n-k) \times (n-k)}}
    \end{bmatrix}
\end{equation}
and let $\mtx{\Omega}_1(t) = \mtx{U}_1(t)^{\top} \mtx{\Omega} \in \mathbb{R}^{k \times 2k}$ and $\mtx{\Omega}_2(t) = \mtx{U}_2(t)^{\top} \mtx{\Omega} \in \mathbb{R}^{(n - k) \times 2k}$ with $\mtx{\Omega} \in \mathbb{R}^{n \times 2k}$.

\begin{lemma}{Moments of tail bound}{ose}
    \todo{A matrix $\mtx{A} \in \mathbb{R}^{l \times l}$ and }independent standard Gaussian random matrices $\mtx{\Omega}_1 \in \mathbb{R}^{k \times 2k}, \mtx{\Omega}_2 \in \mathbb{R}^{l \times 2k}$ satisfy for all $k \in \mathbb{N}$ and a universal constant $c > 0$
    \begin{equation}
        \mathbb{E}^{k}\left[ \lVert \mtx{A} \mtx{\Omega}_2 \mtx{\Omega}_1^{\dagger} \rVert _2 \right] \leq c \lVert \mtx{A} \rVert _2 .
    \end{equation}
\end{lemma}

\begin{proof}
    We first use the submultiplicativity of the spectral norm and the stochastic independendence of the random matrices
    \begin{equation}
        \mathbb{E}^{k}\left[ \lVert \mtx{A} \mtx{\Omega}_2 \mtx{\Omega}_1^{\dagger} \rVert _2 \right]
        \leq \mathbb{E}^{k}\left[ \lVert \mtx{A} \rVert _2 \lVert \mtx{\Omega}_2 \rVert _2 \lVert \mtx{\Omega}_1^{\dagger} \rVert _2 \right]
        = \lVert \mtx{A} \rVert _2 \mathbb{E}^{k}\left[ \lVert \mtx{\Omega}_2 \rVert _2 \right] \mathbb{E}^{k}\left[  \lVert \mtx{\Omega}_1^{\dagger} \rVert _2 \right].
        \label{equ:OSE-moment-bound-initial-new}
    \end{equation}

    To bound the moments of $\lVert \mtx{\Omega}_2 \rVert _2$ we define $\widetilde{\mtx{\Omega}}_2 = \mtx{\Omega}_2^{\top} / \sqrt{2k}$ and rewrite
    \begin{align}
        &\mathbb{E}^{k}\left[ \lVert \mtx{\Omega}_2 \rVert _2 \right] \notag \\
        &= \sqrt{2k} \cdot \mathbb{E}^{k}\left[ \lVert \widetilde{\mtx{\Omega}}_2^{\top} \rVert _2 \right] && \text{(definition of $\widetilde{\mtx{\Omega}}_2$)} \notag \\
        &= \sqrt{2k} \cdot \sqrt{ \mathbb{E}^{\sfrac{k}{2}}\left[ \lVert \widetilde{\mtx{\Omega}}_2^{\top} \widetilde{\mtx{\Omega}}_2 \rVert _2 \right] } && \text{($\lVert \mtx{B} \rVert _2 = \lVert \mtx{B}^{\top} \mtx{B} \rVert _2^{\sfrac{1}{2}} $ )} \notag \\
        &\leq \sqrt{2k} \cdot \sqrt{ \mathbb{E}^{\sfrac{k}{2}}\left[ \lVert \widetilde{\mtx{\Omega}}_2^{\top} \widetilde{\mtx{\Omega}}_2 - \mtx{I}_l \rVert _2 \right] + \mathbb{E}^{\sfrac{k}{2}}\left[ \lVert \mtx{I}_l \rVert _2 \right]} && \text{(triangle inequality)} \notag \\
        &\leq \sqrt{2k} \cdot \sqrt{ \mathbb{E}^{\sfrac{k}{2}}\left[ \lVert \widetilde{\mtx{\Omega}}_2^{\top} \widetilde{\mtx{\Omega}}_2 - \mtx{I}_l \rVert _2 \right]} + \sqrt{2k} &&
         \text{($\sqrt{a+b} \leq \sqrt{a} + \sqrt{b}$, $\lVert \mtx{I}_l \rVert _2 = 1$)}
    \end{align}
    By \cite[section 2.2]{nelson-2017-sketching-algorithms} $\widetilde{\mtx{\Omega}}_2 \in \mathbb{R}^{2k \times l}$ satisfies the $(c_1 \sqrt{q/2k}, e^{-q}, q)$-JL moment property for any $q > 0$ and a universal constant $c_1 > 0$. Hence, by \cite[lemma 4]{cohen-2016-optimal-approximate} it satisfies the $(2 c_1 \sqrt{q/2k}, 9^d e^{-q}, d, q)$-OSE moment property for any $d \in \mathbb{N}$. For convenience, we choose $d=2k$ and $q=k/2$, such that $\widetilde{\mtx{\Omega}}_2$ has $(c_1, (3^4 e^{-\sfrac{1}{2}})^k, 2k, k/2)$-OSE moment property, i.e. \cite[definition 3]{cohen-2016-optimal-approximate}
    \begin{equation}
        \mathbb{E}^{\sfrac{k}{2}} \left[ \lVert \widetilde{\mtx{\Omega}}_2 \widetilde{\mtx{\Omega}}_2^{\top} - \mtx{I}_l \rVert _2 \right] \leq c_1 (3^4 e^{-\sfrac{1}{2}})^2 = c_1 3^8 e^{-1}.
        %\mathbb{E}^{\sfrac{k}{2}}\left[ \lVert (\widetilde{\mtx{\Omega}}_2 \mtx{A} )^{\top} (\widetilde{\mtx{\Omega}}_2 \mtx{A}) - \mtx{A}^{\top} \mtx{A} \rVert _2\right] \leq c_1 c_2^2 3^8e^{-1} \left( \lVert \mtx{A} \rVert _2^2 + \frac{\lVert \mtx{A} \rVert _F^2}{k} \right).
    \end{equation}
    Inserting this bound into the above inequality we get
    \begin{equation}
        \mathbb{E}^{k}\left[ \lVert \mtx{\Omega}_2 \rVert _2 \right]
        \leq \sqrt{2k} \cdot (3^4 \sqrt{c_1 e^{-1}} + 1).
        \label{equ:OSE-moment-bound-first}
    \end{equation}

    The moments of $\lVert \mtx{\Omega}_1^{\dagger} \rVert _2$ can be bound with help of the proof of \cite[lemma B.3]{tropp-2023-randomized-algorithms}
    \begin{equation}
        \mathbb{E}\left[ \lVert \mtx{\Omega}_1^{\dagger} \rVert _2^{k} \right]
        = \mathbb{E}\left[ \lVert ( \mtx{\Omega}_1 \mtx{\Omega}_1^{\top} )^{-1} \rVert _2^{\sfrac{k}{2}} \right]
        \leq \left( k + 1 \right) \left( \frac{1}{(k + 1)!}\right)^{\sfrac{k}{k + 1}} \left( \frac{3 k}{2}\right)^{\sfrac{k}{2}}.
    \end{equation}
    With the Taylor series expansion of the exponential function it can be shown that $e^n \geq 1 + n$ and $e^n \geq \frac{n^n}{n!}$ for all $n \in \mathbb{N}$, from which $(n + 1)^{\sfrac{1}{n}} \leq e$ and $\left( \frac{1}{n!} \right)^{\sfrac{1}{n}} \leq \frac{e}{n}$ follow respectively. Hence,
    \begin{equation}
        \mathbb{E}^{k}\left[ \lVert \mtx{\Omega}_1^{\dagger} \rVert _2 \right]
        \leq \frac{e^2}{k + 1}\sqrt{ \frac{3k}{2} }
        \leq e^2 \sqrt{\frac{3}{2k}}.
        \label{equ:OSE-moment-bound-second}
    \end{equation}

    Putting \refequ{equ:OSE-moment-bound-first} and  \refequ{equ:OSE-moment-bound-second} together, we get
    \begin{equation}
        \mathbb{E}^{k}\left[ \lVert \mtx{A} \mtx{\Omega}_2 \mtx{\Omega}_1^{\dagger} \rVert _2 \right]
        \leq \underbrace{e^2\sqrt{3} (3^4 \sqrt{c_1 e^{-1}} + 1)}_{\equiv c} \lVert \mtx{A} \rVert _2.
    \end{equation}

\end{proof}

\begin{lemma}{Nyström approximation for parameter-dependent matrices}{nystrom}
    Let $\mtx{B}(t) \in \mathbb{R}^{n \times n}$ be symmetric positive semi-definite and continuously depend on $t \in [a, b]$. Then its Nyström approximation $\Nystr{\mtx{\Omega}}{\mtx{B}}(t)$ with Gaussian random matrix $\mtx{\Omega} \in \mathbb{R}^{n \times n_{\mtx{\Omega}}}$ for \emph{even} $n_{\mtx{\Omega}} \in \mathbb{N}$ and $\gamma \geq 1$ verifies with probability at least $1 - \gamma^{-\sfrac{n_{\mtx{\Omega}}}{4}}$
    \begin{equation}
        \int_{a}^{b} \lVert \mtx{B}(t) - \Nystr{\mtx{\Omega}}{\mtx{B}}(t) \rVert _F~\mathrm{d}t \leq \gamma \sqrt{2} \cdot \frac{1 + c}{\sqrt{n_{\mtx{\Omega}}}} \int_{a}^{b} \Trace(\mtx{B}(t))~\mathrm{d}t.
    \end{equation}
    for a universal constant $c > 0$. In particular, if we choose $n_{\mtx{\Omega}} = \mathcal{O}(\varepsilon^{-2} \log(\delta^{-1})), n_{\mtx{\Omega}} \geq 2$, then with probability at least $1-\delta$ for $\delta \in (0, e^{-e^{4 \varepsilon^2}})$ and $\varepsilon > 0$ we have $\int_{a}^{b} \lVert \mtx{B}(t) - \Nystr{\mtx{\Omega}}{\mtx{B}}(t) \rVert _F~\mathrm{d}t \leq \varepsilon \int_{a}^{b} \Trace(\mtx{B}(t))~\mathrm{d}t$.
\end{lemma}

%\todo{Proof idea: structural bound, then higher order moment bound to apply Markov's inequality.}

\begin{proof}
    We let $n_{\mtx{\Omega}} = 2k$ for some $k \in \mathbb{N}$, i.e. $\mtx{\Omega} \in \mathbb{R}^{n \times 2k}$. After temporarily omitting the parameter dependence, we can apply \cite[theorem B.1]{persson-2023-randomized-lowrank} with $f(x) = x$ to bound the integrand with% \todo{maybe explain where $\mtx{\Omega}_i$ come from...}
    \begin{equation}
        \lVert \mtx{B} - \Nystr{\mtx{\Omega}}{\mtx{B}} \rVert _F 
        \leq  \lVert \mtx{\Lambda}_2 \rVert _F + \lVert (\mtx{\Lambda}_2^{\sfrac{1}{2}} \mtx{\Omega}_2 \mtx{\Omega}_1^{\dagger} )^2 \rVert _F.
        \label{equ:nystrom-proof-persson-bonud}
    \end{equation}

    As in the proof of \cite[lemma 3]{meyer-2021-hutch-optimal}, the first term can be bonud with
    \begin{equation}
        \lVert \mtx{\Lambda}_2 \rVert _F
        = \sqrt{\sum_{i=k+1}^{n} \lambda_i^2}
        \leq \sqrt{ \lambda_{k+1} \sum_{i=k+1}^{n} \lambda_i}
        \leq \sqrt{ \frac{\Trace(\mtx{B})}{k} \Trace(\mtx{B})}
        \leq \frac{1}{\sqrt{k}} \Trace(\mtx{B}).
        \label{equ:nystrom-proof-frobenius-trace}
    \end{equation}

    The second term is first processed with standard matrix-norm inequalities:
    \begin{equation}
        \lVert (\mtx{\Lambda}_2^{\sfrac{1}{2}} \mtx{\Omega}_2 \mtx{\Omega}_1^{\dagger} )^2 \rVert _F 
        \leq \lVert \mtx{\Lambda}_2^{\sfrac{1}{2}} \mtx{\Omega}_2 \mtx{\Omega}_1^{\dagger} \rVert _2 \lVert \mtx{\Lambda}_2^{\sfrac{1}{2}} \mtx{\Omega}_2 \mtx{\Omega}_1^{\dagger} \rVert _F
        \leq \sqrt{k} \lVert \mtx{\Lambda}_2^{\sfrac{1}{2}} \mtx{\Omega}_2 \mtx{\Omega}_1^{\dagger} \rVert _2^2.
        \label{equ:nystrom-proof-processed-tail}
    \end{equation}
    Its higher order moments are therefore bounded by
    \begin{align}
        \mathbb{E}^{\sfrac{k}{2}}\left[ \lVert (\mtx{\Lambda}_2^{\sfrac{1}{2}} \mtx{\Omega}_2 \mtx{\Omega}_1^{\dagger} )^2 \rVert _F \right]
        &\leq \sqrt{k} \cdot \mathbb{E}^{\sfrac{k}{2}}\left[ \lVert \mtx{\Lambda}_2^{\sfrac{1}{2}} \mtx{\Omega}_2 \mtx{\Omega}_1^{\dagger} \rVert _2^2 \right] && \text{(using \refequ{equ:nystrom-proof-processed-tail})} \notag \\
        &= \sqrt{k} \cdot \mathbb{E}^{k}\left[ \lVert \mtx{\Lambda}_2^{\sfrac{1}{2}} \mtx{\Omega}_2 \mtx{\Omega}_1^{\dagger} \rVert _2 \right]^2 && \text{(by definition of $\mathbb{E}^{k}$)} \notag \\
        &\leq \sqrt{k} c \cdot \lVert \mtx{\Lambda}_2^{\sfrac{1}{2}} \rVert _2^2 && \text{(\reflem{lem:ose} with $\mtx{A}=\mtx{\Lambda}_2^{\sfrac{1}{2}}$)} \notag \\
        &= \sqrt{k} c \cdot \lambda_{k+1} && \text{(from the definition of $\lVert \cdot \rVert _2$)} \notag \\
        &\leq \frac{c}{\sqrt{k}} \Trace(\mtx{B}) && \text{($\lambda_{k+1} \leq \Trace(\mtx{B})/k$)}
        \label{equ:nystrom-proof-tail-bound}
    \end{align}
    Thus, using the triangle inequality for $\mathbb{E}^{\sfrac{k}{2}}$ and inserting \refequ{equ:nystrom-proof-tail-bound} along with \refequ{equ:nystrom-proof-frobenius-trace} in \refequ{equ:nystrom-proof-persson-bonud} we obtain
    \begin{equation}
        \mathbb{E}^{\sfrac{k}{2}} \left[\lVert \mtx{B} - \Nystr{\mtx{\Omega}}{\mtx{B}} \rVert _F \right]
        \leq \frac{1 + c}{\sqrt{k}} \Trace(\mtx{B})
    \end{equation}

    We use Minkowski's integral inequality \cite[theorem 202]{hardy-1952-inequalities} to get
    \begin{equation}
        \mathbb{E}^{\sfrac{k}{2}}\left[ \int_{a}^{b} \lVert \mtx{B}(t) - \Nystr{\mtx{\Omega}}{\mtx{B}}(t) \rVert _F~\mathrm{d}t \right] 
        \stackrel{\text{\cite{hardy-1952-inequalities}}}{\leq} \int_{a}^{b} \mathbb{E}^{\sfrac{k}{2}} \left[\lVert \mtx{B} - \Nystr{\mtx{\Omega}}{\mtx{B}} \rVert _F \right] ~\mathrm{d}t% \notag \\
        %&\leq \mathbb{E}^{\sfrac{k}{2}} \left[ \frac{1}{\sqrt{k}} \int_{a}^{b} \Trace(\mtx{B}(t)) + \sqrt{k} \lVert \mtx{\Lambda}_2(t)^{\sfrac{1}{2}} \mtx{\Omega}_2(t) \mtx{\Omega}_1(t)^{\dagger} \rVert _2^2~\mathrm{d}t  \right] && \text{(\refequ{xy})} \notag \\
        %&\leq \frac{1}{\sqrt{k}} \int_{a}^{b} \Trace(\mtx{B}(t))~\mathrm{d}t + \mathbb{E}^{\sfrac{k}{2}} \left[ \int_{a}^{b}  \sqrt{k} \lVert \mtx{\Lambda}_2(t)^{\sfrac{1}{2}} \mtx{\Omega}_2(t) \mtx{\Omega}_1(t)^{\dagger} \rVert _2^2~\mathrm{d}t  \right] \notag \\
        %&\leq \frac{1}{\sqrt{k}} \int_{a}^{b} \Trace(\mtx{B}(t))~\mathrm{d}t + \sqrt{k} \int_{a}^{b} \mathbb{E}^{\sfrac{k}{2}} \left[ \lVert \mtx{\Lambda}_2(t)^{\sfrac{1}{2}} \mtx{\Omega}_2(t) \mtx{\Omega}_1(t)^{\dagger} \rVert _2^2 \right]  ~\mathrm{d}t \notag \\
        \leq \frac{1 + c}{\sqrt{k}} \int_{a}^{b} \Trace(\mtx{B}(t))~\mathrm{d}t.
    \end{equation}
    From Markov's inequality follows with probability at least $1 - \gamma^{-\sfrac{k}{2}}$
    \begin{equation}
        \int_{a}^{b} \lVert \mtx{B}(t) - \Nystr{\mtx{\Omega}}{\mtx{B}}(t) \rVert _F~\mathrm{d}t \leq \gamma \frac{1 + c}{\sqrt{k}} \int_{a}^{b} \Trace(\mtx{B}(t))~\mathrm{d}t,
    \end{equation}
    from which follows the first statement by replacing $k = \frac{n_{\mtx{\Omega}}}{2}$.

    Setting $\delta = \gamma^{-\sfrac{k}{2}}$ and choosing $k = \lceil \varepsilon^{-2}\log(\delta^{-1}) \rceil$ we get
    \begin{align}
        \gamma \frac{1}{\sqrt{k}}
        &= \delta^{-\sfrac{2}{\lceil \varepsilon^{-2} \log(\delta^{-1}) \rceil}} \frac{1}{\sqrt{\lceil \varepsilon^{-2} \log(\delta^{-1}) \rceil}} && \text{(definition of $\delta$ and choice of $k$)} \notag \\
        &= e^{2\varepsilon^{2} \frac{\varepsilon^{-2}\log(\delta^{-1})}{\lceil \varepsilon^{-2} \log(\delta^{-1}) \rceil}} \frac{1}{\sqrt{\lceil \varepsilon^{-2} \log(\delta^{-1}) \rceil}} && \text{($\delta = e^{-\log(\delta^{-1})}$)} \notag \\ 
        &\leq e^{2\varepsilon^{2}} \frac{\varepsilon}{\sqrt{ \log(\delta^{-1})}} && \text{($\lceil x \rceil \geq x$ if $x \geq 0$)}
    \end{align}
    which is smaller than $\varepsilon$ if $\delta \leq e^{-e^{4 \varepsilon^2}}$from which follows the second part of the theorem.
\end{proof}

Finally, we combine the analysis of the two estimators from \refsec{subsec:hutchinson} and \refsec{subsec:nystrom} into a more efficient, general-purpose trace estimator.

\begin{theorem}{Nyström++ trace estimator for parameter-dependent matrices}{nystrom-pp}
    Let $\mtx{B}(t) \in \mathbb{R}^{n \times n}$ be symmetric positive semi-definite and continuously depend on $t \in [a, b]$. If $n_{\mtx{\Psi}} = n_{\mtx{\Omega}} = \mathcal{O}(\varepsilon^{-1} \log(2\delta^{-1})^2)$ with \emph{even} $n_{\mtx{\Omega}} \geq 2$, then for $\delta \in (0, 2^{-1} e^{-e^{4\varepsilon}})$ with probability at least $1 - \delta$
    \begin{equation}
        \int_{a}^{b} | \Trace(\mtx{B}(t)) - \Nystrpp{\mtx{\Omega}}{\mtx{\Psi}}(\mtx{B}(t)) |~\mathrm{d}t
        \leq \varepsilon \int_{a}^{b} \Trace(\mtx{B}(t))~\mathrm{d}t.
    \end{equation}
\end{theorem}

\begin{proof}
    By choosing $n_{\mtx{\Psi}} = n_{\mtx{\Omega}} = \mathcal{O}(\tilde{\varepsilon}^{-2} \log(\tilde{\delta}^{-1})^2)$ we get
    \begin{align}
        &\int_{a}^{b} | \Trace(\mtx{B}(t)) - \Nystrpp{\mtx{\Omega}}{\mtx{\Psi}}(\mtx{B}(t)) |~\mathrm{d}t \notag \\
        &= \int_{a}^{b} | \Hutch{\mtx{\Psi}}(\mtx{B}(t) - \Nystr{\mtx{\Omega}}{\mtx{B}}(t)) - \Trace(\mtx{B}(t) - \Nystr{\mtx{\Omega}}{\mtx{B}}(t)) |~\mathrm{d}t && \text{(by definition of estimators)} \notag \\
        &\leq \tilde{\varepsilon} \int_{a}^{b} \lVert \mtx{B}(t) - \Nystr{\mtx{\Omega}}{\mtx{B}}(t) \rVert _F ~\mathrm{d}t && \text{(\refthm{thm:hutchinson} w.p. $\geq 1 - \tilde{\delta}$)} \notag \\
        &\leq \tilde{\varepsilon}^2 \int_{a}^{b} \Trace(\mtx{B}(t)) ~\mathrm{d}t && \text{(\reflem{lem:nystrom} w.p. $\geq 1 - \tilde{\delta}$)} 
    \end{align}
    with probability at least $1 - 2\tilde{\delta}$. Taking $\varepsilon = \tilde{\varepsilon}^2$ and $\delta = 2 \tilde{\delta}$
\end{proof}

Comparing \refthm{thm:nystrom-pp} with its equivalent for constant matrices \cite[theorem 3.4]{persson-2022-improved-variants}, we notice that our choice of $\delta$ is more restricted and the dependence of $n_{\mtx{\Psi}}$ and $n_{\mtx{\Omega}}$ on $\delta$ is slightly worse.

% \begin{proof}
%     Combining with union bound.
%     \begin{align}
%         &\int_{a}^{b} | \Trace(\mtx{B}(t)) - \Nystrpp{\mtx{\Omega}}{\mtx{\Psi}}(\mtx{B}(t)) |~\mathrm{d}t \notag \\
%         &= \int_{a}^{b} | \Hutch{\mtx{\Psi}}(\mtx{B}(t) - \Nystr{\mtx{\Omega}}{\mtx{A}}(t)) - \Trace(\mtx{B}(t) - \Nystr{\mtx{\Omega}}{\mtx{A}}(t)) |~\mathrm{d}t && \text{(definition of estimators)} \notag \\
%         &= \todo{c} \frac{1}{\sqrt{n_{\mtx{\Psi}}}} \int_{a}^{b} \lVert \mtx{B}(t) - \Nystr{\mtx{\Omega}}{\mtx{A}}(t) \rVert _F ~\mathrm{d}t && \text{(with probability $1 - \delta$ if $n_{\mtx{\Psi}} = \mathcal{O}(\varepsilon^{-1})$)} \notag \\
%         &= \todo{c} \frac{1}{\sqrt{n_{\mtx{\Psi}} n_{\mtx{\Omega}}}} \int_{a}^{b} \Trace(\mtx{B}(t)) ~\mathrm{d}t
%     \end{align}
% \end{proof}

\section{Application to spectral density approximation}
\label{sec:application}

We consider the task of computing the spectral density of a real symmetric matrix $\mtx{A} \in \mathbb{R}^{n \times n}$ with eigenvalues $\lambda_1, \dots, \lambda_n$, which is defined as
\begin{equation}
    \phi(t) = \frac{1}{n} \sum_{i=1}^{n} \delta(t - \lambda_i),
    \label{equ:spectral-density}
\end{equation}
where $\delta$ is the Dirac delta distribution. Finding approximations to this distribution is challenging. Furthermore, in most applications the exact location of every eigenvalue is not important, but more so their approximate locations relative to each other, such as eigenvalue clusters, spectral gaps, or outliers. In the following we will introduce a smooth version of the spectral density, which is significantly easier to approximate at the cost of losing some of the finer characteristics of the spectrum.

\subsection{Smooth spectral density}
\label{subsec:spectral-density}

The smooth spectral density $\phi_{\sigma}$ is defined as
\begin{equation}
    \phi_{\sigma}(t) = \frac{1}{n} \sum_{i=1}^{n} g_{\sigma}(t - \lambda_i) = \frac{1}{n} \Trace(g_{\sigma}(t\mtx{I}_n - \mtx{A}))
    \label{equ:smooth-spectral-density}
\end{equation}
for a smoothing kernel $g_{\sigma}$ parametrized by a smoothing parameter $\sigma > 0$, which controls by how much $\phi$ is smoothed. We choose a Gaussian smoothing kernel of width $\sigma > 0$, given by
\begin{equation}
    g_{\sigma}(s) = \frac{1}{\sigma \sqrt{2\pi}} e^{-\sfrac{s^2}{2\sigma^2}}.
    \label{equ:smoothing-kernel}
\end{equation}
The task in \refequ{equ:smooth-spectral-density} is to approximate the trace of the parameter-dependent matrix $g_{\sigma}(t\mtx{I}_n - \mtx{A})$. Hence, we can apply the results derived in \refsec{sec:analysis}.

Choosing the smoothing parameter $\sigma$ results in the following trade off: Typically, larger $\sigma$ make it easier to approximate the matrix function $g_{\sigma}(t\mtx{I}_n - \mtx{A})$, whereas smaller $\sigma$ allow us to stay closer to the original spectral density \refequ{equ:spectral-density}. To see the latter, we measure the error between spectral density $\phi$ and its smoothened version $\phi_{\sigma}$ with
\begin{equation}
    \sup_{f \in \mathcal{S}} \int_{-1}^{1} f(t) (\phi(t) - \phi_{\sigma}(t))~\mathrm{d}t,
    \label{equ:error-metric}
\end{equation}
where $\mathcal{S}$ represents an appropriately chosen space of test functions. Among others, \cite{lin-2016-approximating-spectral} proposes $\mathcal{S} = \{ f: f(t) \equiv g_{\sigma}(t - \lambda), \lambda \in [\lambda_{\text{min}}, \lambda_{\text{max}}]\}$, \cite{chen-2021-analysis-stochastic} uses $\mathcal{S} = \{f : f(t) = \Theta(s - t), s \in [-1, 1] \}$ where $\Theta$ is the Heaviside step function, and \cite{braverman-2022-sublinear-time} uses $\mathcal{S} = \{f : |f(x) - f(y)| \leq |x - y| \}$ for which \refequ{equ:error-metric} is equivalent to the Wasserstein-1 distance between $\phi$ and $\phi_{\sigma}$. In this last metric, if we assume the spectrum to be somewhat uniform and $\sigma$ to be rather small, a simple calculation allows us to roughly estimate the smoothing error \refequ{equ:error-metric} to be $\sigma$. Consequently, if we can only allow an approximation to deviate by at most a factor of $\varepsilon > 0$ from the original spectral density, we should choose a smoothing parameter $\sigma \approx \varepsilon$.

\subsection{Chebyshev expansion of smoothing kernel}
\label{subsec:chebyshev-expansion}

To estimate the trace in \refequ{equ:smooth-spectral-density}, we first need to evaluate the involved matrix function $g_{\sigma}(t\mtx{I}_n - \mtx{A})$. However, doing this exactly would require us to first diagonalize $\mtx{A}$, which is often a prohibitively expensive operation. For algorithmic reasons, we seek for approximations of this matrix function in terms of affine expansions, of which the Chebyshev expansion is most suitable due to its approximation error guarantees and ease of manipulation. In every $t \in \mathbb{R}$ we expand 
\begin{equation}
    g_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A}) = \sum_{l=0}^{m} \mu_l(t) T_l(\mtx{A})
    \label{equ:matrix-expansion}
\end{equation}
in terms of a set of $m+1$ coefficients $\mu_0(t), \dots, \mu_m(t)$ which linearly combine the Chebyshev polynomials
\begin{equation}
    \begin{cases}
        T_l : [-1, 1] \to \mathbb{R} \\
        T_l(s) = \cos(l \cdot \arccos(s))
    \end{cases}
    \label{equ:chebyshev-polynomial}
\end{equation}
evaluated at the matrix $\mtx{A}$. It is easy to verify that these polynomials satisfy the three-term recurrence relation
\begin{equation}
    T_l(s) =
    \begin{cases}
        s^l, & \text{if $l \in \{0, 1\}$}; \\
        T_l(s) = 2 s T_{l-1}(s) - T_{l-2}(s), & \text{if $l \geq 2$.}
    \end{cases}
    \label{equ:chebyshev-recurrence}
\end{equation}
With the help of \refequ{equ:matrix-expansion} we can now define the expanded spectral density
\begin{equation}
    \phi_{\sigma}^{(m)}(t) =  \frac{1}{n} \Trace(g_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A})).
    \label{equ:expanded-spectral-density}
\end{equation}

However, the Chebyshev polynomials are only well-defined for matrices whose spectrum is contained in $[-1, 1]$. When working with general matrices of bounded spectrum, we will first have to estimate a lower bound $a$ and upper bound $b$ on the spectrum \cite{zhou-2011-bounding-spectrum}. A spectral transform
\begin{equation}
    \begin{cases}
        \tau : [a, b] \to [-1, 1],\\
        \tau(t) = \frac{2t - a - b}{b - a}.
    \end{cases}
\end{equation}
then lets us define the matrix $\bar{\mtx{A}} = \tau(\mtx{A})$ whose spectrum is identical to the one of $\mtx{A}$ but shrunk to the interval $[-1, 1]$, for which \refequ{equ:matrix-expansion} is a valid expansion. Note that through this transformation we will also have to adjust the smoothing parameter $\bar{\sigma} = 2 \sigma (b - a)^{-1}$ to obtain an undistorted approximation of the original spectrum when reverting the spectral transform $\tau$.

\subsubsection{Fast and consistent squaring}
\label{subsubsec:dct}

It is well known that the coefficients of a Chebyshev expansion of a function can be computed which a discrete cosine transform (DCT) of the function evaluations at the Chebyshev nodes $s_i = \cos(\pi i / m)$ for $i=0,\dots,m$
\begin{equation}
    g_{\sigma}(t - \cos(s_i)) = \sum_{l=0}^{m} \mu_l(t) \cos\left(\frac{\pi i l}{m} \right).
    \label{equ:discrete-cosine-transform}
\end{equation}
This can be accomplished in $\mathcal{O}(m \log(m))$ operations.

A highly useful property of the Chebyshev expansion -- of which we will make use twice in the following pages -- is the fact that we can (algebraically) square such an expansion in $\mathcal{O}(m \log(m))$ operations, which will be shown in the below algorithm.

\begin{algo}{Fast squaring of Chebyshev expansion}{chebyshev-squaring}
    \begin{algorithmic}[1]
        \Statex \textbf{Input:} Coefficients $\{ \mu_l \}_{l=0}^{m}$ of the expansion $\sum_{l=0}^{m} \mu_l T_l$
        \Statex \textbf{Output:} Coefficients $\{ \nu_l \}_{l=0}^{2m}$ such that $\sum_{l=0}^{2m} \nu_l T_l = (\sum_{l=0}^{m} \mu_l T_l)^2$
        \State Define $\vct{\mu} \in \mathbb{R}^{2m + 1}$ whose first $m + 1$ entries are the coefficients $\mu_l$ and the remaining $m$ are $0$
    \State Compute $\vct{f} = \DCT(\vct{\mu})$
    \State Compute $\vct{\nu} = \DCT^{-1}(\vct{f}^{2})$ whose entries are the coefficients $\nu_l$
    \end{algorithmic}
    \end{algo}

\subsubsection{Expansion error}
\label{subsubsec:expansion-error}

We now analyze the error of the Chebyshev expansion of a Gaussian smoothing kernel $g_{\sigma}$.

\begin{lemma}{Chebyshev expansion error}{chebyshev-error}
    The expansion $g_{\sigma}^{(m)}$ of the Gaussian smoothing kernel $g_{{\sigma}}$ \refequ{equ:smoothing-kernel} satisfies
    \begin{equation}
        \sup_{t \in [-1, 1]} \left| g_{\sigma}(t - s) - g_{\sigma}^{(m)}(t - s) \right| \leq \sqrt{\frac{2e}{\pi}} \frac{1}{\sigma^2} (1 + \sigma)^{-m} \equiv E_{\sigma, m}
        \label{equ:2-chebyshev-interpolation-sup-error-kernel}
        %\lVert  \phi_{\sigma} - \phi_{\sigma}^{(m)} \rVert _1 &\leq \frac{2\sqrt{2}}{\sigma^2} (1 + \sigma)^{-m}.
        %\label{equ:2-chebyshev-interpolation-error}
    \end{equation}
    for all degrees $m \in \mathbb{N}$, smoothing parameters $\sigma > 0$, $t \in \mathbb{R}$, and $s \in [-1, 1]$.
\end{lemma}

The proof specializes Bernstein's theorem for Chebyshev interpolation to the case of a Gaussian smoothing kernel \refequ{equ:smoothing-kernel}. It loosely follows the proof give in the preprint of \cite[theorem 2]{lin-2017-randomized-estimation}, but reaches slightly different conclusion.

\begin{proof}
    From Bernstein's theorem \cite[theorem 4.3]{trefethen-2008-gauss-quadrature} it follows that if $f$ is analytic within the Bernstein ellipse $\mathcal{E}_{\chi} \subset \mathbb{C}$ with foci $\{-1, +1\}$ and sum of semi-axes $\chi > 1$, then for any $m \in \mathbb{N}$
    \begin{equation}
        \sup_{s \in [-1, 1]} \left| f(s) - f^{(m)}(s) \right| \leq \frac{2}{\chi^m (\chi - 1)} \max_{z \in \mathcal{E}_{\chi}} |f(z)|.
        \label{equ:bernstein-bound}
    \end{equation}
    In particular, when $f = g_{\sigma}(t - \cdot)$ we observe for $z = x + iy \in \mathcal{E}_{\chi}$
    \begin{equation}
    | g_{\sigma}(t - z) | 
    = \frac{1}{\sigma \sqrt{2 \pi}} \left| e^{-\sfrac{(t - z)^2}{2\sigma^2}} \right|
    = \frac{1}{\sigma \sqrt{2 \pi}} e^{-\sfrac{(t - x)^2}{2\sigma^2}}e^{\sfrac{y^2}{2\sigma^2}}.
    %\leq \frac{1}{n \sigma \sqrt{2 \pi}} \max_{x + iy \in \mathcal{E}_{\chi}} e^{\sfrac{y^2}{2\sigma^2}} 
    \end{equation}
    Because $e^{-\sfrac{(t - x)^2}{2\sigma^2}} \leq 1$ for all $x, t \in \mathbb{R}$, and the maximum absolute value of $y$ is limited by the length of the semi-axis of the Bernstein ellipse in the direction of the imaginary axis, which is $(\chi - \chi^{-1}) / 2$, we upper bound
    \begin{equation}
        \max_{z \in \mathcal{E}_{\chi}} | g_{\sigma}(t - z) | 
        \leq \frac{1}{\sigma \sqrt{2 \pi}} e^{\sfrac{(\chi - \chi^{-1})^2}{8 \sigma^2}}.
    \end{equation}

    Choosing $\chi = 1 + \sigma$ gives a simple and good bound, since in this case $\chi - \chi^{-1} \leq 2\sigma$, which implies
    \begin{equation}
        \sup_{t \in [-1, 1]} \left| g_{\sigma}(t - s) - g_{\sigma}^{(m)}(t - s) \right| \leq \sqrt{\frac{2e}{\pi}} \frac{1}{\sigma^2} (1 + \sigma)^{-m}.
    \end{equation}
\end{proof}

\Reflem{lem:chebyshev-error} now also allows us to bound the error we make with the expanded spectral density $\phi_{\sigma}^{(m)}$ \refequ{equ:expanded-spectral-density}.

\begin{theorem}{$L^1$-error of Chebyshev expansion of spectral density}{chebyshev-error}
    Let $\phi_{\sigma}$ be the smooth spectral density of a symmetric matrix $\mtx{A} \in \mathbb{R}^{n \times n}$ whose spectrum is contained in $[-1, 1]$. Then the spectal density with Chebyshev expansion $\phi_{\sigma}^{(m)}$ satisfies
    \begin{equation}
        \int_{-1}^{1} \left| \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) \right|~\mathrm{d}t \leq 2 E_{\sigma, m}
        \label{equ:chebyshev-interpolation-sup-error-kernel}
        %\lVert  \phi_{\sigma} - \phi_{\sigma}^{(m)} \rVert _1 &\leq \frac{2\sqrt{2}}{\sigma^2} (1 + \sigma)^{-m}.
        %\label{equ:2-chebyshev-interpolation-error}
    \end{equation}
    for all degrees $m \in \mathbb{N}$, smoothing parameters $\sigma > 0$, and $t \in \mathbb{R}$. %The same result holds for the expansion of a non-negative Chebyshev expansion of \emph{even} degree $m \in \mathbb{N}$ but with $\underline{E}_{\sigma, m}$.
\end{theorem}

\begin{proof}
    We observe that for all $t \in [-1, 1]$
    \begin{align}
        &\left| \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) \right| \notag \\
        &= \left| \frac{1}{n} \Trace(g_{\sigma}(t\mtx{I}_n - \mtx{A})) - \frac{1}{n} \Trace(g_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A})) \right|
        && \text{(definitions \refequ{equ:smooth-spectral-density} and \refequ{equ:expanded-spectral-density})} \notag \\
        &= \left| \frac{1}{n} \sum_{i=1}^n \left(g_{\sigma}(t - \lambda_i) - g_{\sigma}^{(m)}(t - \lambda_i)\right) \right|
        && \text{($\lambda_1, \dots, \lambda_n$ eigenvalues of $\mtx{A}$)} \notag \\
        &\leq \max_{i = 1, \dots, n} \left| g_{\sigma}(t - \lambda_i) - g_{\sigma}^{(m)}(t - \lambda_i) \right|
        && \text{(conservative upper bound)} \notag \\
        &\leq E_{\sigma, m}.% ~ \text{(or $\underline{E}_{\sigma, m}$)}.
        && \text{(using \reflem{lem:chebyshev-error})} \notag \\
    \end{align}
    Then, Hölder's inequality allows us to conclude
    \begin{equation}
        \int_{-1}^{1} | \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) | ~\mathrm{d}t
        \leq 2 \sup_{t \in [-1, 1]} \left| \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) \right|
        \leq 2 E_{\sigma, m}.% ~ \text{(or $2 \underline{E}_{\sigma, m}$)}.
    \end{equation}
\end{proof}

\subsection{Chebyshev-Nyström++ method for spectral density approximation}
\label{subsec:chebyshev-nystrom}

Our goal is to apply the trace estimator analyzed in \refsec{subsec:nystrom-pp} to the Chebyshev expansion of the smoothing kernel $g_{\sigma}$ to approximate the smooth spectral density $\phi_{\sigma}$. However, due to the oscillatory nature of the Chebyshev expansion, we can no longer guarantee that $g_{\sigma}^{(m)} \geq 0$, and consequently, the matrix function $g_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A})$ may be indefinite. To still apply \refthm{thm:nystrom-pp} in this scenario, we will have to ensure that the Chebyshev expansion is non-negative. We can think of many ways in which this can be achieved: by using Jackson damping \cite{jackson-1912-approximation-trigonometric,braverman-2022-sublinear-time}, albeit at a significantly worse rate of convergence; by shifting up the expansion $g_{\sigma}^{(m)} + \rho$ with a large enough $\rho$, at the loss of most of the low-rank structure of $g_{\sigma}(t \mtx{I}_n - \mtx{A})$; or by directly minimizing the approximation error over all non-negative Chebyshev polynomials \cite{fejer-1916-uber-trigonometrische}, which realistically only works for small $m$. In the end, we propose the following algorithm to compute a non-negative Chebyshev expansion. First, expand the square root of the smoothing kernel in the Chebyshev polynomials up to degree $m/2$
\begin{equation}
    \sqrt{g_{\sigma}(t - s)}^{(\sfrac{m}{2})} = \sum_{l=0}^{m/2} \xi_l(t) T_l(s).
\end{equation}
Second, square this expansion with \refalg{alg:chebyshev-squaring} to obtain the non-negative Chebyshev expansion
\begin{equation}
    \underline{g}^{(m)}_{\sigma}(t - s) = \sum_{l=0}^{m} \mu_l(t) T_l(s).
\end{equation}
With a complexity of $\mathcal{O}(m \log(m))$ this procedure is computationally inexpensive and allows for very similar convergence bounds to \reflem{lem:chebyshev-error}, which we will derive in \refsec{subsubsec:chebyshev-nystrom-analysis}.

\todo{
Maybe mention indefinite Nyström [Nakatsukasa], which won't give us desirable guarantees, and refinement of Chebyshev expansion [Francis Bach], which only slightly improves, but is much harder to implement. Side-note: it is possible to refine the coefficients of the Chebyshev expansion to get a better approximation in the supremum-norm over the set of all non-negative polynomials \cite{fejer-1916-uber-trigonometrische}. The non-convex nature of this problem and the relatively small improvement we get with this technique made us decide to not pursue this further.
}

In the end, we propose the following approximation for the smooth spectral density
\begin{equation}
    \widetilde{\phi}_{\sigma}^{(m)}(t) = \frac{1}{n} \Nystrpp{\mtx{\Omega}}{\mtx{\Psi}}(\underline{g}_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A})).
    \label{equ:chebyshev-nystrom-formula}
\end{equation}

\subsubsection{Implementation}
\label{subsubsec:chebyshev-nystrom-implementation}

An efficient implementation of \refequ{equ:chebyshev-nystrom-formula} can be achieved thanks to the invariance of the trace under cyclic permutation of its arguments and the recurrence relation \refequ{equ:chebyshev-recurrence} which the Chebyshev polynomials satisfy. To illustrate this, consider the estimator \refequ{equ:chebyshev-nystrom-formula} when $\mtx{\Psi}$ is zero. Using the definition of the Nyström approximation \refequ{equ:nystrom-approximation}, the cyclic invariance of the trace, and the symmetry of $\mtx{A}$, we can rewrite it as
\begin{equation}
    \frac{1}{n} \Trace((\mtx{\Omega}^{\top} \underline{g}^{(m)}_{\sigma}(t \mtx{I}_n - \mtx{A}) \mtx{\Omega})^{\dagger} (\mtx{\Omega}^{\top} \underline{g}^{(m)}_{\sigma}(t \mtx{I}_n - \mtx{A})^2 \mtx{\Omega})).
    \label{equ:cyclic-property}
\end{equation}
Thanks to the affine form of the Chebyshev expansion, we can rewrite the first matrix as
\begin{equation}
    \mtx{\Omega}^{\top} \underline{g}^{(m)}_{\sigma}(t \mtx{I}_n - \mtx{A}) \mtx{\Omega} = \sum_{l=0}^{m} \mu_l(t) (\mtx{\Omega}^{\top} T_l(\mtx{A}) \mtx{\Omega})
\end{equation}
for which we first precompute the $n_{\mtx{\Omega}} \times n_{\mtx{\Omega}}$ matrices $\mtx{\Omega}^{\top} T_l(\mtx{A}) \mtx{\Omega}$ for $l=0, \dots, m$ using the recurrence \refequ{equ:chebyshev-recurrence} and then inexpensively sum them up with the corresponding coefficients $\mu_l(t)$ for all $t$ in which we want to evaluate the approximation. Thanks to \refalg{alg:chebyshev-squaring}, the second part of \refequ{equ:cyclic-property} can be approximated in an equivalent way \todo{(which is consistent, unlike \cite{lin-2017-randomized-estimation})}.

Similar observations for the case when $\mtx{\Psi}$ is nonzero lead to the following algorithm.

\begin{algo}{Chebyshev-Nystrom++}{nystrom-chebyshev-pp}
\begin{algorithmic}[1]
    \Statex \textbf{Input:} Symmetric matrix $\mtx{A} \in \mathbb{R}^{n \times n}$, evaluation points $\{t_i\}_{i=1}^{n_t}$
    \Statex \textbf{Parameters:} Degree $m$, allocation $(n_{\mtx{\Psi}}, n_{\mtx{\Omega}})$,  smoothing parameter $\sigma$
    \Statex \textbf{Output:} Approximate evaluations of the spectral density $\{\widetilde{\phi}_{\sigma}(t_i)\}_{i=1}^{n_t}$
    \State Compute $\{\mu_l(t_i)\}_{l=0}^{m}$ for all $t_i$ using the $\DCT$ \refequ{equ:discrete-cosine-transform}
    \State Compute $\{\nu_l(t_i)\}_{l=0}^{2m}$ using \refalg{alg:chebyshev-squaring}%\refalg{alg:2-chebyshev-chebyshev-expansion}
    %\State Compute  for all $t_i$ using %\refalg{alg:3-nystrom-chebyshev-exponentiation}
    \State Generate Gaussian random matrices $\mtx{\Omega} \in \mathbb{R}^{n \times n_{\mtx{\Omega}}}$ and $\mtx{\Psi} \in \mathbb{R}^{n \times n_{\mtx{\Psi}}}$%\glsfirst{sketching-matrix} $\in \mathbb{R}^{n \times n_{\mtx{\Omega}}}$
    %\State Generate standard Gaussian %\glsfirst{random-matrix} $\in \mathbb{R}^{n \times n_{\mtx{\Psi}}}$
    \State Initialize $[\mtx{V}_1, \mtx{V}_2, \mtx{V}_3] \gets [\mtx{0}_{n \times n_{\mtx{\Omega}}}, \mtx{\mtx{\Omega}}, \mtx{0}_{n \times n_{\mtx{\Omega}}}]$
    \State Initialize $[\mtx{W}_1, \mtx{W}_2, \mtx{W}_3] \gets [\mtx{0}_{n \times n_{\mtx{\Psi}}}, \mtx{\Psi}, \mtx{0}_{n \times n_{\mtx{\Psi}}}]$
    \State Initialize $[\mtx{K}_1(t_i), \mtx{K}_2(t_i)] \gets [\mtx{0}_{n_{\mtx{\Omega}} \times n_{\mtx{\Omega}}}, \mtx{0}_{n_{\mtx{\Omega}} \times n_{\mtx{\Omega}}}]$ for all $t_i$
    \State Initialize $[\mtx{L}_1(t_i), \ell(t_i)] \gets [\mtx{0}_{n_{\mtx{\Omega}} \times n_{\mtx{\Psi}}}, 0]$ for all $t_i$
    %\State Set $\breve{\phi}_{\sigma}^{(m)}(t_i) \gets 0$ for all $t_i$
    \For {$l = 0, \dots, 2m$}
    \State $[\mtx{X}, \mtx{Y}] \gets \mtx{\mtx{\Omega}}^{\top} [\mtx{V}_2, \mtx{W}_2]$  
    %\State $\mtx{X} \gets \mtx{\mtx{\Omega}}^{\top} \mtx{V}_2$
      %\State $\mtx{Y} \gets \mtx{\mtx{\Omega}}^{\top} \mtx{W}_2$
      \State $z \gets \Trace(\mtx{\Psi}^{\top} \mtx{W}_2)$
      \For {$i = 1, \dots, n_t$}
        \If {$l \leq m$}
            \State $\mtx{K}_1(t_i) \gets \mtx{K}_1(t_i) + \mu_l(t_i) \mtx{X}$ \Comment{assemble $\mtx{\Omega}^{\top} g_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A}) \mtx{\Omega}$}
            \State $\mtx{L}_1(t_i) \gets \mtx{L}_1(t_i) + \mu_l(t_i) \mtx{Y}$ \Comment{assemble $\mtx{\Omega}^{\top} g_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A}) \mtx{\Psi}$}
            \State $\ell(t_i) \gets \ell(t_i) + \mu_l(t_i) z$ \Comment{assemble $\Trace(\mtx{\Psi}^{\top} g_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A}) \mtx{\Psi})$}
        \EndIf
        \State $\mtx{K}_2(t_i) \gets \mtx{K}_2(t_i) + \nu_l(t_i) \mtx{X}$ \Comment{assemble $\mtx{\Omega}^{\top} g_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A})^2 \mtx{\Omega}$}
      \EndFor
      \State $[\mtx{V}_3, \mtx{W}_3] \gets (2 - \delta_{l0}) \mtx{A} [\mtx{V}_2, \mtx{W}_2] - [\mtx{V}_1, \mtx{W}_1]$ \Comment{Chebyshev recurrence}
      \State $[\mtx{V}_1, \mtx{W}_1] \gets [\mtx{V}_2, \mtx{W}_2]$
      \State $[\mtx{V}_2, \mtx{W}_2] \gets [\mtx{V}_3, \mtx{W}_3]$
      %\State $\mtx{V}_3 \gets (2 - \delta_{l0}) \mtx{A} \mtx{V}_2 - \mtx{V}_1$ %\Comment{Chebyshev recurrence \refequ{equ:2-chebyshev-chebyshev-recursion}}
      %\State $\mtx{V}_1 \gets \mtx{V}_2, \mtx{V}_2 \gets \mtx{V}_3$
      %\State $\mtx{W}_3 \gets (2 - \delta_{l0}) \mtx{A} \mtx{W}_2 - \mtx{W}_1$ %\Comment{Chebyshev recurrence \refequ{equ:2-chebyshev-chebyshev-recursion}}
      %\State $\mtx{W}_1 \gets \mtx{W}_2, \mtx{W}_2 \gets \mtx{W}_3$
    \EndFor
    \For {$i = 1, \dots, n_t$}
      \State $\widetilde{\phi}_{\sigma}(t_i) \gets \frac{1}{n} \Trace\left( \mtx{K}_1(t_i)^{\dagger}\mtx{K}_2(t_i) \right) + \frac{1}{n n_{\mtx{\Psi}}} \left( \ell(t_i) + \Trace\left( \mtx{L}_1(t_i)^{\top} \mtx{K}_1(t_i)^{\dagger} \mtx{L}_1(t_i) \right)  \right) $ \label{lin:4-nystromchebyshev-nystrom-pp}
    \EndFor
\end{algorithmic}
\end{algo}

The dominating cost stems from the $(2m + 1)(n_{\mtx{\Psi}} + n_{\mtx{\Omega}})$ multiplications of $\mtx{A}$ with a vector and the cost associated with computing the estimate in each evaluation point $t_1, \dots, t_{n_t}$ given by $\mathcal{O}(m \log(m) + m (n_{\mtx{\Omega}}^2 + n_{\mtx{\Omega}} n_{\mtx{\Psi}} + 1) + (n_{\mtx{\Omega}}^3 + n_{\mtx{\Omega}} n_{\mtx{\Psi}}^2 + 1))$.

The algorithm also accomodates the two other estimators introduced in \refsec{sec:analysis}: For $n_{\mtx{\Omega}} = 0$ we recover \refequ{equ:hutchinson-trace-estimator}, and the complexity for each evaluation of the estimate decreases significantly. For $n_{\mtx{\Psi}} = 0$ we end up with \refequ{equ:nystrom-trace-estimator}.

\todo{Clearly need not divide by $n_{\mtx{\Psi}}=0$, hence just ignore this term. }
%Chebyshev $\mathcal{O}(n_t (m \log(m) + m (n_{\mtx{\Omega}}^2 + n_{\mtx{\Omega}} n_{\mtx{\Psi}} + 1) + n_{\mtx{\Omega}}^3) + mn^2(n_{\mtx{\Psi}} + n_{\mtx{\Omega}}))$.

%\todo{Complexity in terms of both $n_{\mtx{\Omega}}$ and $n_{\mtx{\Psi}}$.}
%With the cost of a matrix-vector product denoted by $c(n)$, and supposing $n_{\mtx{\Omega}} \approx n_{\mtx{\Psi}}$, we determine the computational complexity of the Chebyshev-Nyström++ method to be $\mathcal{O}(m \log(m) n_t + m n_{\mtx{\Omega}}^2 n + m n_t n_{\mtx{\Omega}}^2 +  m c(n) n_{\mtx{\Omega}} + n_t n_{\mtx{\Omega}}^3)$, with $\mathcal{O}(m n_t + n n_{\mtx{\Omega}} + n_{\mtx{\Omega}}^2 n_t)$ required additional storage.

%\todo{Explain modifications for other estimators, and improved complexity for quadratic trace estimator}

\todo{Discuss computation of pseudo inverse: Generalized eigenvalue problem with thresholding \cite{lin-2017-randomized-estimation, epperly-2022-theory-quantuma}, }

\todo{Explain short-circuit mechanism -> very crucial if matrix can be zero}

\subsubsection{Analysis}
\label{subsubsec:chebyshev-nystrom-analysis}

\begin{lemma}{Non-negative Chebyshev expansion error}{non-negative-chebyshev-error}
    The non-negative expansion $\underline{g}_{\sigma}^{(m)}$ of the Gaussian smoothing kernel satisfies
    \begin{equation}
        \sup_{t \in [-1, 1]} \left| g_{\sigma}(t - s) - \underline{g}_{\sigma}^{(m)}(t - s) \right| \leq 2\sqrt{2} \left(1 + \sigma \sqrt{\pi} \cdot E_{\sqrt{2}\sigma, \sfrac{m}{2}}\right) E_{\sqrt{2}\sigma, \sfrac{m}{2}} \equiv \underline{E}_{\sigma, m}
        \label{equ:2-chebyshev-interpolation-sup-error-kernel}
        %\lVert  \phi_{\sigma} - \phi_{\sigma}^{(m)} \rVert _1 &\leq \frac{2\sqrt{2}}{\sigma^2} (1 + \sigma)^{-m}.
        %\label{equ:2-chebyshev-interpolation-error}
    \end{equation}
    for all \emph{even} degrees $m \in \mathbb{N}$, smoothing parameters $\sigma > 0$, $t \in \mathbb{R}$, and $s \in [-1, 1]$.
\end{lemma}

\begin{proof}
    For any numbers $a, b \in \mathbb{R}$ it holds
    \begin{equation}
    | a^2 - b^2 | = | (a + b)(a - b) | = | a + b | | a - b | \leq (2 | a | + | a - b |)  | a - b |
    \end{equation}
    Therefore, omitting the arguments and using $g_{\sigma} = (\sqrt{g_{\sigma}})^2$ and $\underline{g}_{\sigma}^{(m)} = (\sqrt{g_{\sigma}}^{(\sfrac{m}{2})})^2$.
    we have
    \begin{equation}
        \left| g_{\sigma} - \underline{g}_{\sigma}^{(m)} \right| \leq \left( 2 \left| \sqrt{g_{\sigma}} \right| + \left| \sqrt{g_{\sigma}} - \sqrt{g_{\sigma}}^{(\sfrac{m}{2})} \right| \right) \left| \sqrt{g_{\sigma}} - \sqrt{g_{\sigma}}^{(\sfrac{m}{2})} \right|.
    \end{equation}
    Since $\sqrt{g_{\sigma}} = \sqrt{2 \sigma \sqrt{2 \pi}} \cdot g_{\sqrt{2}\sigma}$ and $\sqrt{g_{\sigma}} \leq 1/\sqrt{\sigma \sqrt{2 \pi}}$ we can apply the result from \refthm{thm:chebyshev-error} to get
    \begin{align}
        &\sup_{t \in [-1, 1]} \left| g_{\sigma}(t - s) - \underline{g}_{\sigma}^{(m)}(t - s) \right| \notag \\
        &\leq \left( 2 \frac{1}{\sqrt{ \sigma \sqrt{2\pi}}} + \sqrt{2 \sigma \sqrt{2 \pi}} \cdot E_{\sqrt{2}\sigma, \sfrac{m}{2}}\right)\sqrt{2 \sigma \sqrt{2 \pi}} \cdot E_{\sqrt{2}\sigma, \sfrac{m}{2}},
    \end{align}
    from which follows the result with some minor simplifications.
\end{proof}

\begin{theorem}{Chebyshev-Nyström++ method}{chebyshev-nystrom}
    Let $\phi_{\sigma}$ be the smooth spectral density of a symmetric matrix $\mtx{A} \in \mathbb{R}^{n \times n}$ whose spectrum is contained in $[-1, 1]$. Then the Chebyshev-Nyström++ estimate $\widetilde{\phi}_{\sigma}^{(m)}$ with \emph{even} degree $m \in \mathbb{N}$, $n_{\mtx{\Omega}} = n_{\mtx{\Psi}} = \mathcal{O}(\varepsilon^{-1}\log(2\delta^{-1})^2)$, and smoothing parameter $\sigma > 0$ satisfies with probability at least $1 - \delta$
    \begin{equation}
        \int_{-1}^{1} \left| \phi_{\sigma}(t) - \widetilde{\phi}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t < 2 (1 + \varepsilon) \underline{E}_{\sigma, m} + \varepsilon
        \label{equ:chebyshev-nystrom-error}
        %\lVert  \phi_{\sigma} - \phi_{\sigma}^{(m)} \rVert _1 &\leq \frac{2\sqrt{2}}{\sigma^2} (1 + \sigma)^{-m}.
        %\label{equ:2-chebyshev-interpolation-error}
    \end{equation}
    for all $\delta \in (0, 2^{-1} \min\{e^{\sfrac{1}{2}}, e^{-e^{8\varepsilon}}\})$.
\end{theorem}

\begin{proof}
    The previously derived results can be applied to
    \begin{align}
        &\int_{-1}^{1} \left| \phi_{\sigma}(t) - \widetilde{\phi}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t \notag \\
        &\leq \int_{-1}^{1} \left| \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) \right|~\mathrm{d}t + \int_{-1}^{1} \left| \phi^{(m)}_{\sigma}(t) - \widetilde{\phi}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t && \text{(triangle inequality)} \notag \\
        &\leq \int_{-1}^{1} \left| \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) \right|~\mathrm{d}t + \varepsilon \int_{-1}^{1} \left| \phi^{(m)}_{\sigma}(t) \right|~\mathrm{d}t && \text{(\refthm{thm:nystrom-pp} w.p. $\geq 1 - \delta$)} \notag \\
        &\leq (1 + \varepsilon) \int_{-1}^{1} \left| \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) \right|~\mathrm{d}t + \varepsilon \int_{-1}^{1} \left| \phi_{\sigma}(t) \right|~\mathrm{d}t && \text{(triangle inequality)} \notag \\
        &\leq 2 (1 + \varepsilon) \underline{E}_{\sigma, m} + \varepsilon. && \text{(\reflem{lem:non-negative-chebyshev-error} and normalization)} \notag \\
    \end{align}
\end{proof}

\todo{Maybe specialize \refthm{thm:nystrom} to spectral densities?}

We now discuss what value to choose for $n_{\mtx{\Omega}}$ in the Nyström trace estimator introduced in \refsec{subsec:nystrom}. In order to guarantee a good approximation error based on \refthm{thm:nystrom}, we analyze the numerical rank of the matrix function $g_{\sigma}(t\mtx{I}_n - \mtx{A})$.

We can express the numerical rank of a matrix with respect to the nuclear norm in terms of its singular values $\sigma_1 \geq \dots \geq \sigma_n \geq 0$
\begin{equation}
    r_{\varepsilon} = \min \{1 \leq r \leq n: \sum_{j=r+1}^n \sigma_{j} \leq \varepsilon \},
    \label{equ:numerical-rank}
\end{equation}
$\varepsilon$ is usually taken to be the double machine precision, i.e. $10^{-16}$. When using a Gaussian smoothing-kernel, we can explicitly write the singular values of $g_{\sigma}(t\mtx{I}_n - \mtx{A})$ as
\begin{equation}
    \sigma_i(t) = g_{\sigma}(t - \lambda_{(i)}) = \frac{1}{\sqrt{2 \pi \sigma^2}} e^{-\sfrac{(t - \lambda_{(i)})^2}{2 \sigma^2}},
    \label{equ:gaussian-kernel-eigenvalues}
\end{equation}
where $\lambda_{(1)}, \dots, \lambda_{(n)}$ denote the eigenvalues of $\mtx{A}$ sorted by increasing distance from spectral-parameter, such that $\sigma_1(t) \geq \dots \geq \sigma_n(t)$. Consequently, by using the closed-form expression of the eigenvalues \refequ{equ:gaussian-kernel-eigenvalues}, we may upper bound the numerical rank of a Gaussian smoothing kernel as
\begin{equation}
    r_{\varepsilon}(g_{\sigma}(t\mtx{I}_n - \mtx{A})) \leq \#\left\{1\leq i\leq n: |t - \lambda_i| < d_{\varepsilon, \sigma} \right\},
    \label{equ:gaussian-kernel-numerical-rank}
\end{equation}
with $d_{\varepsilon, \sigma} = \sigma \sqrt{-2 \log(\sqrt{2 \pi} \sigma \varepsilon)}$ and the eigenvalues $\lambda_1, \dots, \lambda_n$ of $\mtx{A}$ and the constants. The expression \refequ{equ:gaussian-kernel-numerical-rank} has a nice visual interpretation: The numerical rank of a matrix is at most equal to the number of eigenvalues which are closer to the parameter $t$ than $d_{\varepsilon, \sigma}$. This is illustrated in \reffig{fig:numerical-rank}.
\begin{figure}[ht]
    \centering
    \input{figures/numerical-rank.tex}
    \caption{The numerical rank of $g_{\sigma}(t\mtx{I}_n - \mtx{A})$ can be
        approximately computed by counting the number of eigenvalues
        $\lambda_{(1)}, \dots, \lambda_{(n)}$ of the matrix $\mtx{A}$ which lie less than
        a constant $d_{\varepsilon, \sigma}$ away from $t$.}
    \label{fig:numerical-rank}
\end{figure}

If we additionally assume the eigenvalues of the matrix $\mtx{A}$ to be evenly distributed within $[a, b]$, that is, in any subinterval of fixed length in $[a, b]$ we can expect to find roughly the same number of eigenvalues, then we can estimate the numerical rank of $g_{\sigma}(t\mtx{I}_n - \mtx{A})$ to be
\begin{equation}
    r_{\varepsilon}(g_{\sigma}(t\mtx{I}_n - \mtx{A})) \lessapprox \frac{2 n}{b - a} \cdot d_{\varepsilon, \sigma}.
    \label{equ:gaussian-kernel-numerical-rank-uniform}
\end{equation}

\todo{Discuss how this informs choice of $n_{\mtx{\Omega}}$.}

\section{Numerical results}
\label{sec:results}

For our first example, we consider the matrix which arises from the second order
finite difference discretization of the Laplace operator $\Delta$ in a potential
field $V$,
\begin{equation}
    \mathcal{A} u(\vct{x}) = - \Delta u(\vct{x}) + V(\vct{x}) u(\vct{x}),
    \label{equ:5-experiments-electronic-hamiltonian}
\end{equation}
for a uniform mesh of size $h=0.6$. The potential $V$ results from a
lattice whose primitive cell is of side-length $L=6$ and in whose center a
potential
\begin{equation}
    \alpha e^{-\sfrac{\lVert \vct{x} \rVert _2^2}{ 2 \beta^2 }}
    \label{equ:5-experiments-gaussian-cell}
\end{equation}
with $\alpha = -4$, $\beta = 2$ is located. The computational domain is chosen
to span $n_c \in \mathbb{N}$ primitive cells in every spatial dimension, hence, yielding
discretization matrices which are growing in size with $n_c$. In our experiments
we consider the three-dimensional case.%, but for visualization purposes, we illustrate the potential in \reffig{fig:gaussian-well} in two dimensions.

%\begin{figure}[ht]
%    \begin{subfigure}[b]{0.32\columnwidth}
%        \input{plots/gaussian-well-1.pgf}
%        \caption{$n_c=1$}
%        \label{fig:gaussian-well-1}
%    \end{subfigure}
%    \begin{subfigure}[b]{0.32\columnwidth}
%        \input{plots/gaussian-well-2.pgf}
%        \caption{$n_c=2$}
%        \label{fig:gaussian-well-2}
%    \end{subfigure}
%    \begin{subfigure}[b]{0.32\columnwidth}
%        \input{plots/gaussian-well-5.pgf}
%        \caption{$n_c=5$}
%        \label{fig:gaussian-well-5}
%    \end{subfigure}
%    \caption{Two dimensional periodic potential $V$ for different sizes $n_c$ of the computational domain.}
%    \label{fig:gaussian-well}
%\end{figure}

\begin{figure}[ht]
    \centering
    \input{plots/convergence.pgf}
    \caption{For increasing values of $n_{\mtx{\Omega}} + n_{\mtx{\Psi}}$ but fixed $m$ we plot the $L^1$ relative approximation error for the model problem with $\sigma=0.005$.}
    \label{fig:convergence}
\end{figure}

\begin{figure}[ht]
    \centering
    \input{plots/distribution.pgf}
    \caption{The NC++ method for different ways of allocations a total of $n_{\mtx{\Omega}} + n_{\mtx{\Psi}}=80$ random vectors to either the Nystr\"om low-rank approximation or the Girard-Hutchinson trace estimation for the Gaussian smoothing kernel with multiple different values of the smoothing parameter. We make the approximation error made in the Chebyshev expansion negligible by rescaling $m=16 / \sigma$ (based on \refthm{thm:chebyshev-error}).}
    \label{fig:distribution}
\end{figure}

\todo{[Comparison plots with other estimators]}

\section{Discussion}
\label{sec:discussion}

\todo{
\begin{itemize}
    \item Robustness (can choose $m$ and $n_{\mtx{\Omega}}$, $n_{\mtx{\Psi}}$ as large as one wants, and not break algorithm, unlike Krylov-aware Lanczos)
\end{itemize}
}

% Acknowledge advice from David and Data/Parameters from Lin Lin

\clearpage
\bibliography{bibliography.bib}

%\appendix

%\clearpage
%\section{Proofs}

\end{document}