### 2024-06-27 paper/figures/numerical-rank.tex
generate plots which illustrate improvements

\begin{tikzpicture}
    \fill[black!15!white] (-2, -0.15) rectangle (1.5, 0.15);
    \draw[thick, ->] (-5, 0) to (5, 0);
    \fill[black] (-4, 0) circle (0.075) node[above] {$\lambda_{(7)}$};
    \fill[black] (-3, 0) circle (0.075) node[above] {$\lambda_{(5)}$};
    \fill[black] (-2.3, 0) circle (0.075) node[above] {$\lambda_{(4)}$};
    \fill[black] (-1, 0) circle (0.075) node[above] {$\lambda_{(2)}$};
    \fill[black] (0.3, 0) circle (0.075) node[above] {$\lambda_{(1)}$};
    \fill[black] (1, 0) circle (0.075) node[above] {$\lambda_{(3)}$};
    \fill[black] (2.8, 0) circle (0.075) node[above] {$\lambda_{(6)}$};
    \fill[black] (3.6, 0) circle (0.075) node[above] {$\lambda_{(8)}$};
    \fill[black] (4.4, 0) circle (0.075) node[above] {$\lambda_{(9)}$};
    \draw[black, ultra thick] (-0.155, 0.15) to (-0.155, -0.15) node[below] {$t$};
    \draw[black, thick] (-2, 0.15) to (-2, -0.15) node[below] {$t - d_{\varepsilon, \sigma}$};
    \draw[black, thick] (1.5, 0.15) to (1.5, -0.15) node[below] {$t + d_{\varepsilon, \sigma}$};
\end{tikzpicture}