### 2025-02-18 paper/re-pro-badge.tex

The code for our numerical experiments can be found on \url{https://github.com/FMatti/Rand-TRACE}. The figures and tables were generated from the commit \href{https://github.com/FMatti/Rand-TRACE/tree/c6a997e}{c6a997e} on 2025-02-18 at 14:03:19 UTC.