### 2025-02-20 paper/spectraldensity.tex
review Chebyshev coefficient computation

\section{Application to spectral density approximation}
\label{sec:application}

We now get back to the task of computing the spectral density $1/n \sum_{i=1}^{n} \delta(t - \lambda_i)$ of a real symmetric matrix $\mtx{A} \in \mathbb{R}^{n \times n}$ with eigenvalues $\lambda_1, \dots, \lambda_n$, where $\delta(\cdot)$ denotes the Dirac delta distribution. As explained in the introduction, finding accurate approximations to this distribution is challenging and we will therefore work with
a smoothed version. %This still yields enough information
%. Furthermore, in most applications the exact location of each individual eigenvalue is not important, but more so their approximate locations relative to each other, such as eigenvalue clusters, spectral gaps, or outliers. In the following we will work with a smoothed version of the spectral density, which is significantly easier to approximate at the cost of losing some of the finer characteristics of the spectrum.

\subsection{Smoothed spectral density}
\label{subsec:spectral-density}

We recall that the smoothed spectral density $\phi_{\sigma}$ is defined as
\begin{equation}
    \phi_{\sigma}(t) = \frac{1}{n} \sum_{i=1}^{n} g_{\sigma}(t - \lambda_i) = \frac{1}{n} \Trace(g_{\sigma}(t\mtx{I}_n - \mtx{A}))
    \label{equ:smooth-spectral-density}
\end{equation}
for a smoothing kernel $g_{\sigma}$ parametrized by a smoothing parameter $\sigma > 0$. In this work, we choose a Gaussian smoothing kernel of width $\sigma > 0$:
\begin{equation}
    g_{\sigma}(s) = \frac{1}{\sigma \sqrt{2\pi}} e^{-\sfrac{s^2}{2\sigma^2}}.
    \label{equ:smoothing-kernel}
\end{equation}
The expression~\refequ{equ:smooth-spectral-density} gives rise to the task of estimating the trace of the parameter-dependent SPSD matrix $g_{\sigma}(t\mtx{I}_n - \mtx{A})$, which has been analyzed in \refsec{sec:analysis}.

Choosing the smoothing parameter $\sigma$ results in the following trade off: A larger $\sigma$ usually makes it  easier to approximate the matrix function $g_{\sigma}(t\mtx{I}_n - \mtx{A})$, whereas a smaller $\sigma$ allow one to stay closer to the original spectral density $\phi$. To quantify the latter, one can measure the error between $\phi$ and its smoothened version $\phi_{\sigma}$ by
\begin{equation}
    \sup_{f \in \mathcal{S}} \int_{a}^{b} f(t) (\phi(t) - \phi_{\sigma}(t))~\mathrm{d}t,
    \label{equ:error-metric}
\end{equation}
where $\mathcal{S}$ represents an appropriately chosen space of test functions. Among others, the reference~\cite{lin-2016-approximating-spectral} uses $\mathcal{S} = \{ f: f(t) \equiv g_{\sigma}(t - s), s \in [a, b]\}$, \cite{chen-2021-analysis-stochastic} uses $\mathcal{S} = \{f : f(t) = \Theta(s - t), s \in [a, b] \}$ with the Heaviside step function $\Theta$, and \cite{braverman-2022-sublinear-time,bhattacharjee-2024-improved-spectral} use $\mathcal{S} = \{f : |f(t) - f(s)| \leq |t - s| \}$ for which the error \refequ{equ:error-metric} is equivalent to the Wasserstein-1 distance between $\phi$ and $\phi_{\sigma}$.
When choosing 
the last metric,
assuming that the eigenvalues are somewhat uniformly distributed and $\sigma$ is rather small, a simple calculation allows one to conclude that the smoothing error \refequ{equ:error-metric} is of the order $\sigma$. Consequently, to ensure that the approximation deviates by at most a factor of $\varepsilon > 0$, one needs to choose $\sigma \sim \varepsilon$.

\subsection{Chebyshev expansion of smoothing kernel}
\label{subsec:chebyshev-expansion}

To estimate the trace in \refequ{equ:smooth-spectral-density} efficiently, we approximate the function 
$s\mapsto g_{\sigma}(t- s)$ by a truncated
Chebyshev expansion $g_{\sigma}^{(m)}(t-s) = 
\sum_{l=0}^{m} \mu_l(t) T_l(s)$
with the expansion coefficients $\mu_l(t) \in \mathbb R$ and the Chebyshev polynomials
$T_l(s) = \cos(l \cdot \arccos(s))$ for $l = 0,1,\ldots,m$.
Assuming that the spectrum of $\mtx{A}$ is contained in $[-1,1]$, this gives rise to
%need to evaluate the involved matrix function $g_{\sigma}(t\mtx{I}_n - \mtx{A})$. However, doing this exactly would require us to first diagonalize $\mtx{A}$, which is often a prohibitively expensive operation. For algorithmic reasons (see \refsec{subsubsec:chebyshev-nystrom-implementation}), we seek for approximations of this matrix function in terms of affine linear expansions, of which the was determined to be most suitable due to its approximation error guarantees and ease of manipulation. In every $t \in \mathbb{R}$ we expand 
\begin{equation}
    g_{\sigma}(t\mtx{I}_n - \mtx{A}) \approx g_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A}) = \sum_{l=0}^{m} \mu_l(t) T_l(\mtx{A}),
    \label{equ:matrix-expansion}
\end{equation}
Inserting~\refequ{equ:matrix-expansion} into~\refequ{equ:smooth-spectral-density} yields the approximate smoothed spectral density
\begin{equation}
    \phi_{\sigma}(t) \approx \phi_{\sigma}^{(m)}(t) =  \frac{1}{n} \Trace(g_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A})).
    \label{equ:expanded-spectral-density}
\end{equation}
\begin{remark}
For a general symmetric matrix $\mtx{A}$, one first estimates an interval $[a,b]$ containing the spectrum of $A$~\cite{zhou-2011-bounding-spectrum}. The usual affine transform
$\tau(t) = \frac{2t - a - b}{b - a}$
then lets us define a matrix $\bar{\mtx{A}} = \tau(\mtx{A})$ for which the spectrum is contained in $[-1, 1]$ and, hence, the expansion \refequ{equ:matrix-expansion} is valid.\footnote{Note that one has to adjust the smoothing parameter $\bar{\sigma} = 2 \sigma (b - a)^{-1}$ to obtain an undistorted approximation of the original spectrum when inverting the transformation.} This allows to assume, from now on, without loss of generality that the spectrum of $\mtx{A}$ is contained in $[-1, 1]$.
\end{remark}

\subsubsection{Expansion error}
\label{subsubsec:expansion-error}

We now analyze the impact on truncation on the accuracy of the spectral density $\phi_{\sigma}$. We start with a standard error estimate concerning the approximation of the Gaussian smoothing kernel.

\begin{lemma}{Chebyshev expansion error}{chebyshev-error}
    The truncated Chebyshev expansion $g_{\sigma}^{(m)}$ of the Gaussian smoothing kernel $g_{{\sigma}}$ from  \refequ{equ:smoothing-kernel} satisfies
    \begin{equation}
        \sup_{s,t \in [-1, 1]} \big| g_{\sigma}(t - s) - g_{\sigma}^{(m)}(t - s) \big| \leq \sqrt{\frac{2e}{\pi}} \frac{1}{\sigma^2} (1 + \sigma)^{-m} =: E_{\sigma, m}
        \label{equ:chebyshev-interpolation-sup-error-kernel}
        %\lVert  \phi_{\sigma} - \phi_{\sigma}^{(m)} \rVert _1 &\leq \frac{2\sqrt{2}}{\sigma^2} (1 + \sigma)^{-m}.
        %\label{equ:2-chebyshev-interpolation-error}
    \end{equation}
    for every degree $m \in \mathbb{N}$ and smoothing parameter $\sigma > 0$.
\end{lemma}
\begin{proof}
The proof of this result is standard and follows the one given in the arXiv preprint of \cite[Theorem 2]{lin-2017-randomized-estimation}, but attains a slightly different dependence on $\sigma$.

Let $\mathcal{E}_{\chi} \subset \mathbb{C}$ denote the Bernstein ellipse  with foci $\{-1, +1\}$ and the sum of semi-axes equal to $\chi > 1$. By Bernstein's theorem \cite[Theorem 4.3]{trefethen-2008-gauss-quadrature},     \begin{equation}
        \sup_{s \in [-1, 1]} \big| g_{\sigma}(t - s) - g_{\sigma}^{(m)}(t - s) \big| \leq \frac{2}{\chi^m (\chi - 1)} \max_{z \in \mathcal{E}_{\chi}} |g_{\sigma}(t - z)|,
        \label{equ:bernstein-bound}
    \end{equation}
    where used the analyticity of $g_{\sigma}(t- \cdot)$.
    For $z = x + \mathrm{i} y \in \mathcal{E}_{\chi}$, we have
    \begin{equation}
    | g_{\sigma}(t - z) | 
    = \frac{1}{\sigma \sqrt{2 \pi}} \big| e^{-\sfrac{(t - z)^2}{2\sigma^2}} \big|
    = \frac{1}{\sigma \sqrt{2 \pi}} e^{-\sfrac{(t - x)^2}{2\sigma^2}}e^{\sfrac{y^2}{2\sigma^2}}
    \le \frac{1}{\sigma \sqrt{2 \pi}} e^{\sfrac{(\chi - \chi^{-1})^2}{8 \sigma^2}},
    %\leq \frac{1}{n \sigma \sqrt{2 \pi}} \max_{x + iy \in \mathcal{E}_{\chi}} e^{\sfrac{y^2}{2\sigma^2}} 
    \end{equation}
    where the last inequality uses 
    $e^{-\sfrac{(t - x)^2}{2\sigma^2}} \leq 1$ and $|y| \le (\chi - \chi^{-1}) / 2$.
    Choosing $\chi = 1 + \sigma$ and using $\chi - \chi^{-1} \leq 2\sigma$ completes the proof of~\eqref{equ:chebyshev-interpolation-sup-error-kernel}.
\end{proof}

For $0 < \sigma \ll 1$, \reflem{lem:chebyshev-error} implies that choosing a degree $m = \mathcal{O}(\sigma^{-1}(\log(\varepsilon^{-1}) + \log(\sigma^{-1})))$ ensures an error $\mathcal{O}(\varepsilon)$. The lemma also allows us to bound the $L^1$-error of the spectral density approximation
$\phi_{\sigma}^{(m)}$ defined in~\eqref{equ:expanded-spectral-density}:
\begin{align}
    \int_{-1}^{1} \left| \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) \right|~\mathrm{d}t 
    &= \int_{-1}^{1} \left| \frac{1}{n} \sum_{i=1}^n \big(g_{\sigma}(t - \lambda_i) - g_{\sigma}^{(m)}(t - \lambda_i)\big) \right|~\mathrm{d}t \notag \\
    &\leq \int_{-1}^{1} \max_{i = 1, \dots, n} \big| g_{\sigma}(t - \lambda_i) - g_{\sigma}^{(m)}(t - \lambda_i) \big|~\mathrm{d}t    \leq 2 E_{\sigma, m}.
    \label{equ:chebyshev-interpolation-spectral-density}
\end{align}

%\begin{theorem}{$L^1$-error of Chebyshev expansion of spectral density}{chebyshev-error}
%    Let $\phi_{\sigma}$ be the smoothed spectral density of a symmetric matrix $\mtx{A} \in \mathbb{R}^{n \times n}$ whose spectrum is contained in $[-1, 1]$. Then the spectal density with Chebyshev expansion $\phi_{\sigma}^{(m)}$ satisfies
%    \begin{equation}
%        \int_{-1}^{1} \left| \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) \right|~\mathrm{d}t \leq 2 E_{\sigma, m}
%        \label{equ:chebyshev-interpolation-spectral-density}
%        %\lVert  \phi_{\sigma} - \phi_{\sigma}^{(m)} \rVert _1 &\leq \frac{2\sqrt{2}}{\sigma^2} (1 + \sigma)^{-m}.
%        %\label{equ:2-chebyshev-interpolation-error}
%    \end{equation}
%    for all degrees $m \in \mathbb{N}$, smoothing parameters $\sigma > 0$, and the error $E_{\sigma, m}$ from \reflem{lem:chebyshev-error}.%, and $t \in \mathbb{R}$. %The same result holds for the expansion of a non-negative Chebyshev expansion of \emph{even} degree $m \in \mathbb{N}$ but with $\underline{E}_{\sigma, m}$.
%\end{theorem}
%
%\begin{proof}
%    We observe that for all $t \in [-1, 1]$
%    \begin{align}
%        &\left| \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) \right| \notag \\
%        &= \left| \frac{1}{n} \Trace(g_{\sigma}(t\mtx{I}_n - \mtx{A})) - \frac{1}{n} \Trace(g_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A})) \right|
%        && \text{(definitions \refequ{equ:smooth-spectral-density} and \refequ{equ:expanded-spectral-density})} \notag \\
%        &= \left| \frac{1}{n} \sum_{i=1}^n \left(g_{\sigma}(t - \lambda_i) - g_{\sigma}^{(m)}(t - \lambda_i)\right) \right|
%        && \text{($\lambda_1, \dots, \lambda_n$ eigenvalues of $\mtx{A}$)} \notag \\
%        &\leq \max_{i = 1, \dots, n} \left| g_{\sigma}(t - \lambda_i) - g_{\sigma}^{(m)}(t - \lambda_i) \right|
%        && \text{(conservative upper bound)} \notag \\
%        &\leq E_{\sigma, m}.% ~ \text{(or $\underline{E}_{\sigma, m}$)}.
%        && \text{(using \reflem{lem:chebyshev-error})} \notag \\
%    \end{align}
%    Then, Hölder's inequality allows us to conclude
%    \begin{equation}
%        \int_{-1}^{1} | \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) | ~\mathrm{d}t
%        \leq 2 \sup_{t \in [-1, 1]} \left| \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) \right|
%        \leq 2 E_{\sigma, m}.% ~ \text{(or $2 \underline{E}_{\sigma, m}$)}.
%    \end{equation}
%\end{proof}

\subsubsection{Preservation of non-negativity} \label{sec:preservnonneg}

In general, the truncated Chebyshev expansion $g_{\sigma}^{(m)}$ does not inherit the  non-negativity of the smoothing kernel $g_{\sigma}$. Consequently, the matrix $g_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A})$ may become 
(slightly) indefinite. This causes issues when attempting to apply the Nyström++ trace estimator (Algorithm~\ref{equ:nystrompp-trace-estimator}). While we have never observed this in our numerical experiments, the Nyström approximation may fail when the SPSD condition is not met; see also the discussion in~\cite{nakatsukasa-2023-randomized-lowrank}. On the theoretical side,~\refthm{thm:nystrom-pp}
critically relies on the SPSD condition.

To ensure a non-negative Chebyshev expansion, we suggest the following simple method~\textcolor{red}{[REFERENCE NEEDED]}: Given an even degree $m$, expand the square root of the smoothing kernel up to degree $m/2$:
\begin{equation}
    \sqrt{g_{\sigma}}(t - s) \approx \sqrt{g_{\sigma}}^{(\sfrac{m}{2})}(t - s) = \sum_{l=0}^{m/2} \xi_l(t) T_l(s).
    \label{equ:square-root-chebyshev-expansion}
\end{equation}
When squaring this expansion, once obtains a polynomial of degree $m$, which admits an exact expansion
of the form 
$(\sum_{l=0}^{m/2} \xi_l(t) T_l(s))^2 = \sum_{l=0}^{m} \underline{\mu}_l(t) T_l(s)$.
Thus, we obtain the non-negative Chebyshev expansion
\begin{equation}
g_{\sigma}(t - s) \approx    \underline{g}_{\sigma}^{(m)}(t - s) = \sum_{l=0}^{m} \underline{\mu}_l(t) T_l(s).
    \label{equ:non-negative-chebyshev-expansion}
\end{equation}
%With a complexity of $\mathcal{O}(m \log(m))$ this procedure is computationally inexpensive and allows for very similar convergence bounds to \reflem{lem:chebyshev-error}, which we will derive \textcolor{red}{in the following?} in \refsec{subsubsec:chebyshev-nystrom-analysis}.
%To this extent, we will first analyze the non-negative Chebyshev expansion.
The corresponding non-negative approximation of the smoothed spectral density,
\begin{equation}
    \phi_{\sigma}(t) \approx \underline{\phi}_{\sigma}^{(m)}(t) := \frac{1}{n} \Trace(\underline{g}_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A})),
    \label{equ:non-negative-expanded-density}
\end{equation}
satisfies the following bound.

\begin{lemma}{$L^1$-error of non-negative approximate smooth spectral density}{non-negative-chebyshev-error}
Let $\phi_{\sigma}$ be the smoothed spectral density defined in~\eqref{equ:smooth-spectral-density}
for a symmetric matrix $\mtx{A} \in \mathbb{R}^{n \times n}$ with spectrum contained in $[-1, 1]$. Then the approximation $\underline{\phi}_{\sigma}^{(m)}$ defined in~\eqref{equ:non-negative-expanded-density} satisfies
    \begin{equation}
        \int_{-1}^{1} \big| \phi_{\sigma}(t) - \underline{\phi}_{\sigma}^{(m)}(t) \big|~\mathrm{d}t \leq 4 \sqrt{2} \left(1 + \sigma \sqrt{\pi} \cdot E_{\sqrt{2}\sigma, \sfrac{m}{2}}\right) E_{\sqrt{2}\sigma, \sfrac{m}{2}} =: \underline{E}_{\sigma, m}
        \label{equ:chebyshev-interpolation-sup-error-kernel-nonneg}
        %\lVert  \phi_{\sigma} - \phi_{\sigma}^{(m)} \rVert _1 &\leq \frac{2\sqrt{2}}{\sigma^2} (1 + \sigma)^{-m}.
        %\label{equ:2-chebyshev-interpolation-error}
    \end{equation}
    for every \emph{even} degree $m \in \mathbb{N}$, smoothing parameter $\sigma > 0$, and with $E_{\sigma, m}$ defined in \reflem{lem:chebyshev-error}.
\end{lemma}
\begin{proof}
    Using that the inequality $| a - b | \leq (2 \sqrt{a} + | \sqrt{a} - \sqrt{b} |)  | \sqrt{a} - \sqrt{b} |$ holds for any non-negative numbers $a, b$, it follows that
    \begin{equation}
        \big| g_{\sigma} - \underline{g}_{\sigma}^{(m)} \big| \leq \big( 2 \big| \sqrt{g_{\sigma}} \big| + \big| \sqrt{g_{\sigma}} - \sqrt{g_{\sigma}}^{(\sfrac{m}{2})} \big| \big) \big| \sqrt{g_{\sigma}} - \sqrt{g_{\sigma}}^{(\sfrac{m}{2})} \big|.
    \end{equation}
    Because of $\sqrt{g_{\sigma}} = \sqrt{2 \sigma \sqrt{2 \pi}} \cdot g_{\sqrt{2}\sigma}$,
we can apply \reflem{lem:chebyshev-error} with the substitutions
    $\sigma \gets \sqrt{2}\sigma$ and $m \gets m/2$ to bound 
    $\big| \sqrt{g_{\sigma}} - \sqrt{g_{\sigma}}^{(\sfrac{m}{2})} \big|$.
    Together with  $|\sqrt{g_{\sigma}}| \leq 1/\sqrt{\sigma \sqrt{2 \pi}}$, we get
    \begin{align}
        \big| g_{\sigma}(t - s) - \underline{g}_{\sigma}^{(m)}(t - s) \big| \notag 
        &\leq \Big( 2 \frac{1}{\sqrt{ \sigma \sqrt{2\pi}}} + \sqrt{2 \sigma \sqrt{2 \pi}} \cdot E_{\sqrt{2}\sigma, \sfrac{m}{2}}\Big)\sqrt{2 \sigma \sqrt{2 \pi}} \cdot E_{\sqrt{2}\sigma, \sfrac{m}{2}} \notag \\
        &= 2\sqrt{2} \big(1 + \sigma \sqrt{\pi} \cdot E_{\sqrt{2}\sigma, \sfrac{m}{2}}\big) E_{\sqrt{2}\sigma, \sfrac{m}{2}}
    \end{align}
    for every $s,t \in [-1, 1]$.
    Bounding the integral as in~\refequ{equ:chebyshev-interpolation-spectral-density} then yields the desired result.
\end{proof}

As the non-negative expansion~\refequ{equ:non-negative-chebyshev-expansion} ``wastes'' half of the degrees of freedom, it can be expected to be less accurate than the standard Chebyshev expansion. On the other hand, $\sqrt{g_{\sigma}} \propto g_{\sqrt{2}\sigma}$ is somewhat easier to approximate because it is less \enquote{pointy} compared to $g_{\sigma}$. Overall, it turns out that one needs to choose the degree of the non-negative expansion roughly a factor $\sqrt{2}$ larger to match the approximation error of the standard Chebyshev expansion (see also \reffig{fig:interpolation-issue}).

%In the following, we will show how the coefficients $\underline{\mu}_0(t), \dots, \underline{\mu}_m(t)$ of the non-negative Chebyshev expansion \refequ{equ:non-negative-chebyshev-expansion} can efficiently be computed and that this expansion exhibits similar convergence properties to the standard Chebyshev expansion.
%\todo{
%Maybe mention indefinite Nyström [Nakatsukasa], which won't give us desirable guarantees, and refinement of Chebyshev expansion [Francis Bach], which only slightly improves, but is much harder to implement. Side-note: it is possible to refine the coefficients of the Chebyshev expansion to get a better approximation in the supremum-norm over the set of all non-negative polynomials \cite%{fejer-1916-uber-trigonometrische}. The non-convex nature of this problem and the relatively small improvement we get with this technique made us decide to not pursue this further.
%}

%\subsection{Computation of expansion coefficients}

\subsubsection{Computation of Chebyshev coefficients}
\label{subsubsec:dct}

\textcolor{red}{??? I don't understand this derivation. Don't you think that~\eqref{equ:discrete-cosine-transform} only holds approximately for the exact $g_\sigma$ and it should be something like $g_\sigma \approx g^{(m)} = ...$. I think there is a major confusion of two different approaches here:
\begin{itemize}
 \item The Chebyshev \emph{expansion} of $f$ is the $\ell^2$-orthogonal projection of $f$ on the first $m+1$ Chebyshev polynomials. The coefficients are computed from $\ell^2$ inner products between the Chebyshev polynomials and $f$. This is probably what Lin is doing. 
 \item The Chebyshev \emph{approximation} used by Chebfun and possibly others proceeds differently. It first replaces (or rather pretends to replace) $f$ by the interpolation of $f$ in (some) Chebyshev nodes. This interpolation admits an exact truncated expansion and the coefficients can be computed from a Fourier transform (or even DCT?) of the interpolation data. This might match what is written below, but I am not sure whether the interpolation nodes are the same. Please check with Trefethen's book.
\end{itemize}
There are two issues now
\begin{enumerate}
 \item I am now unsure how $g_\sigma^{(m)}$ is precisely defined. This should be spelled out clearer in the beginning of Section 3.2.
 \item There could be a mismatch between the $g_\sigma^{(m)}$ of Lemma 3.1 and the one we are actually using. This mismatch can most likely be resolved, because a result of the form of Lemma 3.1 also holds for the Chebyshev interpolation. If it is easy to resolve, please do so. If not, at least point this out here. Have a look at Trefethen's book; it might have the right error bound.
\end{enumerate}
???
It is well known that the coefficients of a Chebyshev expansion of a function can --- in exact arithmetic --- be computed by inverting the discrete cosine transform (DCT)
\begin{equation}
    g_{\sigma}(t - \cos(s_i)) = \sum_{l=0}^{m} \mu_l(t) \cos\left(\frac{\pi i l}{m} \right)
    \label{equ:discrete-cosine-transform}
\end{equation}
between the expansion coefficients $\mu_l(t),~i=0,\dots,m$ and the function evaluations at the Chebyshev nodes of the second kind $s_i = \cos(\pi i / m),~i=0,\dots,m$ \cite{vanloan-1992-computational-frameworks, baszenski-1997-fast-polynomial}. This can be accomplished in $\mathcal{O}(m \log(m))$ operations.}

A Chebyshev expansion can be squared in $\mathcal{O}(m \log(m))$ operations by using the DCT and its inverse \cite{baszenski-1997-fast-polynomial}; see Algorithm~\ref{lin:inverse-DCT}. This allows one to efficiently compute the non-negative Chebyshev expansion \refequ{equ:non-negative-chebyshev-expansion}, and it will find further use in \refsec{subsec:chebyshev-nystrom}.

\begin{algo}{Fast squaring of Chebyshev expansions}{chebyshev-squaring}
    \begin{algorithmic}[1]
        \Statex \textbf{Input:} Coefficients $\{ \mu_l \}_{l=0}^{m}$ of Chebyshev expansion $\sum_{l=0}^{m} \mu_l T_l$
        \Statex \textbf{Output:} Coefficients $\{ \nu_l \}_{l=0}^{2m}$ such that $\sum_{l=0}^{2m} \nu_l T_l = (\sum_{l=0}^{m} \mu_l T_l)^2$
        \State By appending $m$ zeros, define vector $\vct{\mu} = [\mu_0, \ldots, \mu_m,0,\ldots,0 ] \in \mathbb{R}^{2m + 1}$.
    \State Compute $\vct{f} = \operatorname{DCT}(\vct{\mu})$
    \State Compute $[\nu_0, \dots, \nu_{2m}] = \operatorname{DCT}^{-1}(\vct{f} \odot \vct{f})$
    \label{lin:inverse-DCT}
    \end{algorithmic}
\end{algo}
On \reflin{lin:inverse-DCT} of Algorithm~\ref{lin:inverse-DCT}, the symbol $\odot$ denotes the entrywise product.

\textcolor{red}{??? To be modified after the notion of Chebyshev expansion is clarified ???
In \cite{lin-2017-randomized-estimation}, a slightly different approach for computing the expansion coefficients of a standard Chebyshev  is taken. They express the coefficients $\mu_l(t)$ in terms of an integral, which they subsequently approximate with a quadrature supported in at least $2m + 1$ nodes. They evaluate the resulting approximation with the fast Fourier transform (FFT). Besides not yielding the exact expansion coefficients $\mu_l(t)$, this procedure is also significantly slower due to the larger number of quadrature nodes and the implementational benefits of the DCT over the FFT. In \reftab{tab:chebyshev-timing-interpolation}, we compare a benevolent implementation of \cite[Algorithm 1]{lin-2017-randomized-estimation} with our approach, both for the standard and the non-negative Chebyshev expansion.}

\begin{table}[ht]
    \caption{Runtime in milliseconds for computing the coefficients of a Chebyshev expansion of the smoothing kernel $g_{\sigma}$ in three different ways: Using Algorithm 1 from~\cite{lin-2017-randomized-estimation} (\emph{FFT}), applying the discrete cosine transform direclty to $g_\sigma$ (\emph{DCT}), and applying the discrete cosine transform to $\sqrt{g_\sigma}$ together with \refalg{alg:chebyshev-squaring} (\emph{non-negative DCT}). \textcolor{red}{??? It is not clear whether the degree from the non-negative DCT is $m$ or $2m$ ???} We use $\sigma=0.005$, $n_t=1000$ parameter values, and various degrees $m$. For stability, we average over 7 runs of the algorithms and repeat this $1000$ times to form the mean and standard deviation reported below.}
    \label{tab:chebyshev-timing-interpolation}
   \input{tables/interpolation.tex}
\end{table}

\subsection{Chebyshev-Nyström++ method for spectral density approximation}
\label{subsec:chebyshev-nystrom}

Combining the results of \refsec{sec:analysis} and \refsec{subsec:chebyshev-expansion}, we apply the Nyström++ estimator $\Nystrpp{\mtx{\Omega}}{\mtx{\Psi}}$ from~\refequ{equ:nystrompp-trace-estimator}
to the non-negative 
Chebyshev expansion  $\mtx{B}(t) = \underline{g}_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A})$, which yields the approximation
\begin{equation}
    \phi_{\sigma}(t) \approx \widetilde{\underline{\phi}}_{\sigma}^{(m)}(t) := \frac{1}{n} \Nystrpp{\mtx{\Omega}}{\mtx{\Psi}}(\underline{g}_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A})).
    \label{equ:chebyshev-nystrom-formula}
\end{equation}

\subsubsection{Implementation}
\label{subsubsec:chebyshev-nystrom-implementation}

One usually aims at evaluating~\eqref{equ:chebyshev-nystrom-formula} for many values of $t$. In the following, we explain how this can be done efficiently by exploiting the 
the well-known three-term recurrence relation of Chebyshev polynomials:
\begin{equation}
    T_0(s) = 1,\quad  T_1(s) = s,\quad 
    T_l(s) = 2 s T_{l-1}(s) - T_{l-2}(s) \text{ for $l \geq 2$.}
    \label{equ:chebyshev-recurrence}
\end{equation}

We start with the first term in the defintion~\refequ{equ:nystrompp-trace-estimator} of $\Nystrpp{\mtx{\Omega}}{\mtx{\Psi}}$, which requires evaluating the trace of the 
Nyström approximation \refequ{equ:nystrom-approximation}:
\begin{equation}
    \frac{1}{n} \Trace \big((\mtx{B}(t) \mtx{\Omega}) (\mtx{\Omega}^{\top} \mtx{B}(t) \mtx{\Omega})^{\dagger} (\mtx{B}(t) \mtx{\Omega})^{\top} \big) =
    \frac{1}{n} \Trace \big((\underbrace{\mtx{\Omega}^{\top} \mtx{B}(t) \mtx{\Omega}}_{=: \mtx{K}_1(t)})^{\dagger} (\underbrace{\mtx{\Omega}^{\top} \mtx{B}(t)^2 \mtx{\Omega}}_{=: \mtx{K}_2(t)}) \big).
    %\frac{1}{n} \Trace((\mtx{\Omega}^{\top} \underline{g}_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A}) \mtx{\Omega})^{\dagger} (\mtx{\Omega}^{\top} \underline{g}_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A})^2 \mtx{\Omega})),
    \label{equ:cyclic-property}
\end{equation}
where we used the invariance of the trace under cyclic permutations.
Inserting the Chebyshev expansion  $\mtx{B}(t) = \underline{g}_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A})$ gives
\begin{equation}
    \mtx{K}_1(t) = \mtx{\Omega}^{\top} \underline{g}_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A}) \mtx{\Omega} = \sum_{l=0}^{m} \underline{\mu}_l(t) (\mtx{\Omega}^{\top} T_l(\mtx{A}) \mtx{\Omega}).
\end{equation}
This allows us to precompute the small $n_{\mtx{\Omega}} \times n_{\mtx{\Omega}}$ matrices $\mtx{\Omega}^{\top} T_l(\mtx{A}) \mtx{\Omega}$ for $l=0, \dots, m$ using the Chebyshev recurrence \refequ{equ:chebyshev-recurrence}. For each value of $t$, we then only need to sum these matrices (multiplied by the corresponding coefficient $\underline{\mu}_l(t)$).

The squaring of Chebyshev expansions discussed in Section~\ref{sec:preservnonneg} allows us to express $\mtx{B}(t)^2 = \underline{g}_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A})^2 = \sum_{l=0}^{2m} \underline{\nu}_l(t) T_l(\mtx{A})$, where the coefficients $\underline{\nu}_l(t)$ can be cheaply computed from $\underline{\mu}_l(t)$ using~\refalg{alg:chebyshev-squaring}). Hence, if we precompute
$\mtx{\Omega}^{\top} T_l(\mtx{A}) \mtx{\Omega}$
also for $l=m+1, \dots, 2m$, this allows us to evaluate $\mtx{K}_2(t)$ cheaply, in an analogous way to $\mtx{K}_1(t)$. Alternatively, \cite{lin-2017-randomized-estimation} proposes to expand the squared smoothing kernel $g_{\sigma}^2$ in a separate Chebyshev expansion of degree $2m$. While this approximation of $g_{\sigma}^2$ might be more accurate, it is \emph{inconsistent} in the sense that computing a Chebyshev expansion of $g_{\sigma}^2$ is in general not the same as squaring an existing Chebyshev expansion of $g_{\sigma}$. \textcolor{red}{??? The discussion in this whole paragraph suffers a bit from that the notion of Chebyshev expansions is ambiguous. ???} We have observed that ensuring consistency between the approximations of $g_{\sigma}$ and $g_{\sigma}^2$ makes a visible difference in terms of accuracy (cf. \reffig{fig:interpolation-issue}). This consistency will also allow us to derive bounds for the estimator \refequ{equ:chebyshev-nystrom-formula}.

\begin{figure}[ht]
    \centering
    \input{plots/interpolation.pgf}
    \caption{Difference in the accuracy when computing the estimate \refequ{equ:cyclic-property} once by separately expanding the squared smoothing kernel $g_{\sigma}^2$ as proposed in~\cite{lin-2017-randomized-estimation} (\emph{inconsistent}), once by squaring the Chebyshev expansion $g_{\sigma}^{(m)}$ using \refalg{alg:chebyshev-squaring} (\emph{consistent}), and once with a non-negative Chebyshev expansion $\underline{g}_{\sigma}^{(m)}$ \refequ{equ:non-negative-chebyshev-expansion} which is also squared with \refalg{alg:chebyshev-squaring} (\emph{non-negative consistent}) \textcolor{red}{??? Please fix legend in Figure. ???}. We use \textcolor{red}{a small example matrix ??? which one? ???} from \refsec{sec:results}, fix $n_{\mtx{\Omega}} = 80$, $\sigma = 0.005$, and compute the $L^1$ error from the approximation of the spectral density at $n_t = 100$ uniformly spaced values of $t$ \textcolor{red}{??? In which interval? ???}.}
    \label{fig:interpolation-issue}
\end{figure}

\color{black}
Analogous observations for the \enquote{Girard-Hutchinson} term which appears in $\Nystrpp{\mtx{\Omega}}{\mtx{\Psi}}$ when $n_{\mtx{\Psi}}$ is nonzero show that also then, the estimator can be expressed in terms of small quadratic forms involving the matrix $\mtx{B}(t) = \underline{g}_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A})$:
\begin{align}
    &\frac{1}{n n_{\mtx{\Psi}}} \Trace \big(\mtx{\Psi}^{\top} (\mtx{B}(t) - (\mtx{B}(t) \mtx{\Omega}) (\mtx{\Omega}^{\top} \mtx{B}(t) \mtx{\Omega})^{\dagger} (\mtx{B}(t) \mtx{\Omega})^{\top}) \mtx{\Psi}^{\top} \big) \notag \\
    &= \frac{1}{n n_{\mtx{\Psi}}} \big( \underbrace{\Trace(\mtx{\Psi}^{\top} \mtx{B}(t)\mtx{\Psi})}_{= \ell(t)} - \Trace((\underbrace{\mtx{\Psi}^{\top} \mtx{B}(t) \mtx{\Omega}}_{= \mtx{L}_1(t)^{\top}}) (\underbrace{\mtx{\Omega}^{\top} \mtx{B}(t) \mtx{\Omega}}_{= \mtx{K}_1(t)^{\top}})^{\dagger} (\underbrace{\mtx{\Omega}^{\top} \mtx{B}(t) \mtx{\Psi}^{\top}}_{= \mtx{L}_1(t)})) \big).
    \label{equ:cyclic-property-hutchinson}
\end{align}

An efficient implementation of the whole procedure described in this section can be found in \refalg{alg:nystrom-chebyshev-pp}.

\begin{algo}{Chebyshev-Nyström++ method}{nystrom-chebyshev-pp}
\begin{algorithmic}[1]
    \Statex \textbf{Input:} Symmetric $\mtx{A} \in \mathbb{R}^{n \times n}$ with spectrum in $[-1, 1]$, points $\{t_i\}_{i=1}^{n_t} \subset \mathbb{R}$
    \Statex \textbf{Parameters:} Even degree of expansion $m \in \mathbb{N}$, number of Girard-Hutchinson queries $n_{\mtx{\Psi}} \in \mathbb{N}_0$, Nyström sketch size $n_{\mtx{\Omega}} \in \mathbb{N}_0$,  smoothing parameter $\sigma > 0$
    \Statex \textbf{Output:} Approximate evaluations of the spectral density $\{\widetilde{\underline{\phi}}_{\sigma}^{(m)}(t_i)\}_{i=1}^{n_t}$
    \For {$i = 1, \dots, n_t$}
        \State Compute the coefficients $\{\xi_l(t_i)\}_{l=0}^{m/2}$ of the Chebyshev expansion of $\sqrt{g_{\sigma}}$ \refequ{equ:square-root-chebyshev-expansion}
        \State Compute the coefficients $\{\underline{\mu}_l(t_i)\}_{l=0}^{m}$ of the non-negative Chebyshev expansion by applying \refalg{alg:chebyshev-squaring} to $\{\xi_l(t_i)\}_{l=0}^{m/2}$
        \State Compute the coefficients $\{\underline{\nu}_l(t_i)\}_{l=0}^{2m}$ of the square of the non-negative Chebyshev expansion by applying \refalg{alg:chebyshev-squaring} to $\{\underline{\mu}_l(t_i)\}_{l=0}^{m}$
    \EndFor
    %\State Compute the coefficients $\{\underline{\mu}_l(t_i)\}_{l=0}^{m}$ of the non-negative Chebyshev expansion \refequ{equ:non-negative-chebyshev-expansion} for all points $t_i$ \textcolor{red}{maybe more details needed (maybe compute $\xi$ for m/2, then twice Alg 3.3)}
    %\State Compute the coefficients $\{\nu_l(t_i)\}_{l=0}^{2m}$ of the square of the non-negative Chebyshev expansion with coefficients $\{\underline{\mu}_l(t_i)\}_{l=0}^{m}$ using \refalg{alg:chebyshev-squaring}%\refalg{alg:2-chebyshev-chebyshev-expansion}
    %\State Compute  for all $t_i$ using %\refalg{alg:3-nystrom-chebyshev-exponentiation}
    \State Generate Gaussian random matrices $\mtx{\Omega} \in \mathbb{R}^{n \times n_{\mtx{\Omega}}}$ and $\mtx{\Psi} \in \mathbb{R}^{n \times n_{\mtx{\Psi}}}$%\glsfirst{sketching-matrix} $\in \mathbb{R}^{n \times n_{\mtx{\Omega}}}$
    %\State Generate standard Gaussian %\glsfirst{random-matrix} $\in \mathbb{R}^{n \times n_{\mtx{\Psi}}}$
    \State Initialize $[\mtx{V}_1, \mtx{V}_2, \mtx{V}_3] \gets [\mtx{0}_{n \times n_{\mtx{\Omega}}}, \mtx{\mtx{\Omega}}, \mtx{0}_{n \times n_{\mtx{\Omega}}}]$
    \State Initialize $[\mtx{W}_1, \mtx{W}_2, \mtx{W}_3] \gets [\mtx{0}_{n \times n_{\mtx{\Psi}}}, \mtx{\Psi}, \mtx{0}_{n \times n_{\mtx{\Psi}}}]$
    \State Initialize $[\mtx{K}_1(t_i), \mtx{K}_2(t_i)] \gets [\mtx{0}_{n_{\mtx{\Omega}} \times n_{\mtx{\Omega}}}, \mtx{0}_{n_{\mtx{\Omega}} \times n_{\mtx{\Omega}}}]$ for all $t_i$
    \State Initialize $[\mtx{L}_1(t_i), \ell(t_i)] \gets [\mtx{0}_{n_{\mtx{\Omega}} \times n_{\mtx{\Psi}}}, 0]$ for all $t_i$
    %\State Set $\breve{\phi}_{\sigma}^{(m)}(t_i) \gets 0$ for all $t_i$
    \For {$l = 0, \dots, 2m$}
    \State $[\mtx{X}, \mtx{Y}] \gets \mtx{\mtx{\Omega}}^{\top} [\mtx{V}_2, \mtx{W}_2]$  
    %\State $\mtx{X} \gets \mtx{\mtx{\Omega}}^{\top} \mtx{V}_2$
      %\State $\mtx{Y} \gets \mtx{\mtx{\Omega}}^{\top} \mtx{W}_2$
      \State $z \gets \Trace(\mtx{\Psi}^{\top} \mtx{W}_2)$
      \For {$i = 1, \dots, n_t$}
        \If {$l \leq m$}
            \State $\mtx{K}_1(t_i) \gets \mtx{K}_1(t_i) + \underline{\mu}_l(t_i) \mtx{X}$ \Comment{assembles $\mtx{\Omega}^{\top} \underline{g}_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A}) \mtx{\Omega}$}
            \State $\mtx{L}_1(t_i) \gets \mtx{L}_1(t_i) + \underline{\mu}_l(t_i) \mtx{Y}$ \Comment{assembles $\mtx{\Omega}^{\top} \underline{g}_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A}) \mtx{\Psi}$}
            \State $\ell(t_i) \gets \ell(t_i) + \underline{\mu}_l(t_i) z$ \Comment{assembles $\Trace(\mtx{\Psi}^{\top} \underline{g}_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A}) \mtx{\Psi})$}
        \EndIf
        \State $\mtx{K}_2(t_i) \gets \mtx{K}_2(t_i) + \underline{\nu}_l(t_i) \mtx{X}$ \Comment{assembles $\mtx{\Omega}^{\top} \underline{g}_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A})^2 \mtx{\Omega}$}
      \EndFor
      \State $[\mtx{V}_3, \mtx{W}_3] \gets (2 - \delta_{l0}) \mtx{A} [\mtx{V}_2, \mtx{W}_2] - [\mtx{V}_1, \mtx{W}_1]$ \Comment{Chebyshev recurrence}
      \State $[\mtx{V}_1, \mtx{W}_1] \gets [\mtx{V}_2, \mtx{W}_2]$
      \State $[\mtx{V}_2, \mtx{W}_2] \gets [\mtx{V}_3, \mtx{W}_3]$
      %\State $\mtx{V}_3 \gets (2 - \delta_{l0}) \mtx{A} \mtx{V}_2 - \mtx{V}_1$ %\Comment{Chebyshev recurrence \refequ{equ:2-chebyshev-chebyshev-recursion}}
      %\State $\mtx{V}_1 \gets \mtx{V}_2, \mtx{V}_2 \gets \mtx{V}_3$
      %\State $\mtx{W}_3 \gets (2 - \delta_{l0}) \mtx{A} \mtx{W}_2 - \mtx{W}_1$ %\Comment{Chebyshev recurrence \refequ{equ:2-chebyshev-chebyshev-recursion}}
      %\State $\mtx{W}_1 \gets \mtx{W}_2, \mtx{W}_2 \gets \mtx{W}_3$
    \EndFor
    \For {$i = 1, \dots, n_t$}
      \State $\widetilde{\underline{\phi}}_{\sigma}(t_i) \gets \frac{1}{n} \Trace\left( \mtx{K}_1(t_i)^{\dagger}\mtx{K}_2(t_i) \right) + \frac{1}{n n_{\mtx{\Psi}}} \left( \ell(t_i) + \Trace\left( \mtx{L}_1(t_i)^{\top} \mtx{K}_1(t_i)^{\dagger} \mtx{L}_1(t_i) \right)  \right) $ \label{lin:4-nystromchebyshev-nystrom-pp}
    \EndFor
\end{algorithmic}
\end{algo}

The dominating cost in \refalg{alg:nystrom-chebyshev-pp} stems from the $(2m + 1)(n_{\mtx{\Psi}} + n_{\mtx{\Omega}})$ matrix-vector multiplications with $\mtx{A}$ and the cost associated with computing the estimate in each evaluation point $t_1, \dots, t_{n_t}$ given by $\mathcal{O}(m \log(m) + m (n_{\mtx{\Omega}}^2 + n_{\mtx{\Omega}} n_{\mtx{\Psi}} + 1) + (n_{\mtx{\Omega}}^3 + n_{\mtx{\Omega}} n_{\mtx{\Psi}}^2 + 1))$.

The algorithm also accomodates the two other estimators introduced in \refsec{sec:analysis}: For $n_{\mtx{\Omega}} = 0$ we recover the Girard-Hutchinson estimator \refequ{equ:hutchinson-trace-estimator}, and the complexity for each evaluation of the estimate decreases significantly. For $n_{\mtx{\Psi}} = 0$ we end up with the Nyström estimator \refequ{equ:nystrom-trace-estimator}, provided the implementation evaluates $0/0$ to $0$ in the second term on \reflin{lin:4-nystromchebyshev-nystrom-pp}. Notice that in this case, \refalg{alg:nystrom-chebyshev-pp} is equivalent to the \enquote{spectrum sweeping} method from \cite[Algorithm 5]{lin-2017-randomized-estimation}, but has a significantly lower computational complexity.

%Chebyshev $\mathcal{O}(n_t (m \log(m) + m (n_{\mtx{\Omega}}^2 + n_{\mtx{\Omega}} n_{\mtx{\Psi}} + 1) + n_{\mtx{\Omega}}^3) + mn^2(n_{\mtx{\Psi}} + n_{\mtx{\Omega}}))$.

%\todo{Complexity in terms of both $n_{\mtx{\Omega}}$ and $n_{\mtx{\Psi}}$.}
%With the cost of a matrix-vector product denoted by $c(n)$, and supposing $n_{\mtx{\Omega}} \approx n_{\mtx{\Psi}}$, we determine the computational complexity of the Chebyshev-Nyström++ method to be $\mathcal{O}(m \log(m) n_t + m n_{\mtx{\Omega}}^2 n + m n_t n_{\mtx{\Omega}}^2 +  m c(n) n_{\mtx{\Omega}} + n_t n_{\mtx{\Omega}}^3)$, with $\mathcal{O}(m n_t + n n_{\mtx{\Omega}} + n_{\mtx{\Omega}}^2 n_t)$ required additional storage.

%\todo{Explain modifications for other estimators, and improved complexity for quadratic trace estimator}

The pseudoinverses on \reflin{lin:4-nystromchebyshev-nystrom-pp} need to be computed with care. It is possible to reformulate the problem as a generalized eigenvalue problem and apply thresholding to the smallest eigenvalues of a certain matrix to enforce stability \cite{lin-2017-randomized-estimation, epperly-2022-theory-quantum}. Since this renders the algorithm significantly more complicated but only improved the accuracy slightly during our experiments, we will instead evaluate these expressions in \reflin{lin:4-nystromchebyshev-nystrom-pp} using a least-squares solver which likewise uses a truncation of the smallest eigenvalues to enforce better conditioning.

What we have found to be crucial for a successful approximation of the full spectral density was the detection of regions wherein lie no eigenvalues, i.e. regions where the spectral density nearly vanishes. For small smoothing parameters $\sigma$, the rapid decay of the Gaussian $g_{\sigma}$ will cause $\mtx{K}_1(t) \approx \mtx{\Omega}^{\top} g_{\sigma}(t\mtx{I}_n - \mtx{A}) \mtx{\Omega}$ to be close to the zero matrix for parameter values $t$ located in such regions. Hence, on \reflin{lin:4-nystromchebyshev-nystrom-pp} we would compute the pseudoinverse of an almost zero matrix, which can be extremely unstable (see \reffig{fig:zerocheck}). Notice that with $\mtx{K}_1(t)$ -- a matrix we need to assemble anyway -- we can cheaply compute $\Trace(\mtx{K}_1(t)) / n_{\mtx{\Omega}}$, which is the $n_{\mtx{\Omega}}$-query Girard-Hutchinson estimator. We may use it as a rough indicator whether $t$ lies within a region where the spectral density nearly vanishes, and subsequently set $\widetilde{\underline{\phi}}_{\sigma}^{(m)}(t) = 0$ instead of computing the full expression on \reflin{lin:4-nystromchebyshev-nystrom-pp}.

\begin{figure}[ht]
    \centering
    \input{plots/zerocheck.pgf}
    \caption{The difference the non-zero check can make when approximating a spectral density using \refalg{alg:nystrom-chebyshev-pp}. Here, we used a matrix described in \refsec{sec:results} and ran the Chebyshev-Nyström++ method with and without estimating if the matrix function is zero before taking its pseudo-inverse. We let $m=2000$, $n_{\mtx{\Omega}}=80$, $n_{\mtx{\Psi}}=0$, and $\sigma = 0.004$.}
    \label{fig:zerocheck}
\end{figure}

\color{black}

\subsubsection{Analysis}
\label{subsubsec:chebyshev-nystrom-analysis}

In this section, we will upper bound the output of the Chebyshev-Nyström++ method (see. \refalg{alg:nystrom-chebyshev-pp}). 
\begin{theorem}{Chebyshev-Nyström++ method}{chebyshev-nystrom}
    Let $\phi_{\sigma}$ be the smoothed spectral density of a symmetric matrix $\mtx{A} \in \mathbb{R}^{n \times n}$ whose spectrum is contained in $[-1, 1]$. Then the Chebyshev-Nyström++ estimate $\widetilde{\underline{\phi}}_{\sigma}^{(m)}$ with \emph{even} degree $m = \mathcal{O}(\sigma^{-1}(\log(\varepsilon^{-1}) + \log(\sigma^{-1})))$, \emph{even} $n_{\mtx{\Omega}} = n_{\mtx{\Psi}} = \mathcal{O}(\varepsilon^{-1}\log(\delta^{-1})^2)$, and smoothing parameter $0 < \sigma \ll 1$ satisfies with probability at least $1 - \delta$
    \begin{equation}
        \int_{-1}^{1} \left| \phi_{\sigma}(t) - \widetilde{\underline{\phi}}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t \leq \varepsilon
        \label{equ:chebyshev-nystrom-error}
        %\lVert  \phi_{\sigma} - \phi_{\sigma}^{(m)} \rVert _1 &\leq \frac{2\sqrt{2}}{\sigma^2} (1 + \sigma)^{-m}.
        %\label{equ:2-chebyshev-interpolation-error}
    \end{equation}
    for all $\varepsilon \in (0, 1)$ and $\delta \in (0, 1/2)$.
\end{theorem}

\begin{proof}
    The previously derived results can be applied to
    \begin{align}
        &\int_{-1}^{1} \left| \phi_{\sigma}(t) - \widetilde{\underline{\phi}}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t \notag \\
        &\leq \int_{-1}^{1} \left| \phi_{\sigma}(t) - \underline{\phi}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t + \int_{-1}^{1} \left| \underline{\phi}_{\sigma}^{(m)}(t) - \widetilde{\underline{\phi}}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t && \text{(triangle inequality)} \notag \\
        &\leq \int_{-1}^{1} \left| \phi_{\sigma}(t) - \underline{\phi}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t + \widetilde{\varepsilon} \int_{-1}^{1} \left| \underline{\phi}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t && \text{(\refthm{thm:nystrom-pp} w.p. $\geq 1 - \delta$)} \notag \\
        &\leq (1 + \widetilde{\varepsilon}) \int_{-1}^{1} \left| \phi_{\sigma}(t) - \underline{\phi}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t + \widetilde{\varepsilon} \int_{-1}^{1} \left| \phi_{\sigma}(t) \right|~\mathrm{d}t && \text{(triangle inequality)} \notag \\
        &\leq (1 + \widetilde{\varepsilon}) \underline{E}_{\sigma, m} + \widetilde{\varepsilon}. && \text{(\reflem{lem:non-negative-chebyshev-error}, normalization)} \notag \\
    \end{align}
    From \reflem{lem:non-negative-chebyshev-error} we see that if $0 < \sigma \ll 1$, choosing $m = \mathcal{O}(\sigma^{-1}(\log(\varepsilon^{-1}) + \log(\sigma^{-1})))$ ensures $\underline{E}_{\sigma, m} \leq \widetilde{\varepsilon}$, and thus, since $\widetilde{\varepsilon} \leq 1$,
    \begin{equation}
        \int_{-1}^{1} \left| \phi_{\sigma}(t) - \underline{\phi}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t \leq (1+\widetilde{\varepsilon}) \widetilde{\varepsilon} + \widetilde{\varepsilon} \leq 3 \widetilde{\varepsilon}
    \end{equation}
    from which the result follows by identifying $\varepsilon = 3 \widetilde{\varepsilon}$.
\end{proof}

%\todo{Maybe specialize \refthm{thm:nystrom} to spectral densities?}

%We now discuss what value to choose for $n_{\mtx{\Omega}}$ in the Nyström trace estimator introduced in \refsec{subsec:nystrom}. In order to guarantee an even better approximation error based on \refthm{thm:nystrom}, we analyze the numerical rank of the matrix function $g_{\sigma}(t\mtx{I}_n - \mtx{A})$.

In practice, the approximation error is often significantly better than guaranteed in \refthm{thm:chebyshev-nystrom}, particularly when $n_{\mtx{\Omega}}$ is chosen rather large. The reason is that for small $\sigma$ the singular values of the matrix function $g_{\sigma}(t\mtx{I}_n - \mtx{A})$ decay so quickly that already the Nyström approximation alone is highly accurate (cf. \refthm{thm:nystrom}). To establish a relation between the singular value decay of $g_{\sigma}(t\mtx{I}_n - \mtx{A})$ and what choice of $n_{\mtx{\Omega}}$ is needed for a significantly better approximation than ensured by \refthm{thm:chebyshev-nystrom}, we study the numerical rank of $g_{\sigma}(t\mtx{I}_n - \mtx{A}) / n$\footnote{The factor $1/n$ comes from the definition of the spectral density \refequ{equ:smooth-spectral-density} when it is written as the trace of a matrix function.}.

We can express the numerical rank of a matrix with respect to the nuclear norm in terms of its singular values $\sigma_1 \geq \dots \geq \sigma_n \geq 0$ as
\begin{equation}
    r_{\varepsilon} = \min \{1 \leq r \leq n: \sum_{j=r+1}^n \sigma_{j} \leq \varepsilon \},
    \label{equ:numerical-rank}
\end{equation}
where $\varepsilon$ is usually taken to be the double machine precision, i.e. $10^{-16}$.

%\begin{theorem}{Nyström estimator for parameter-dependent matrices}%{chebyshev-nystrom-raw}
%    Let $\phi_{\sigma}$ be the smooth spectral density of a symmetric %matrix $\mtx{A} \in \mathbb{R}^{n \times n}$ whose spectrum is %contained in $[-1, 1]$. Then the Chebyshev-Nyström++ estimate %$\widetilde{\phi}_{\sigma}^{(m)}$ with \emph{even} degree $m \in %\mathbb{N}$, $n_{\mtx{\Omega}} \geq \max(6, \max_{t \in [-1, 1]}(r_%{\varepsilon}(g_{\sigma}(t\mtx{I}_n - \mtx{A}))) + \log(\delta^%{-1}))$, $n_{\mtx{\Psi}} = 0$, and smoothing parameter $\sigma > 0$ %satisfies with for all $\gamma \geq 1$ with probability at least $1 %- \delta$
%    \begin{equation}
%        \int_{a}^{b} \left| \phi_{\sigma}(t) - \widetilde{\phi}_{\sigma}%(t) \right| ~\mathrm{d}t
%        \leq 2 e^2 (1 + r_{\varepsilon}) (\varepsilon + (n - r_%{\varepsilon}) \underline{E}_{\sigma, m}) + \underline{E}_%{\sigma, m}.
%    \end{equation}
%\end{theorem}
%
%\begin{proof}
%    \todo{TODO}
%\end{proof}

When using a Gaussian smoothing-kernel, we can explicitly write the singular values of $g_{\sigma}(t\mtx{I}_n - \mtx{A}) / n$ as
\begin{equation}
    \sigma_i(t) = g_{\sigma}(t - \lambda_{(i)}) / n = \frac{1}{n \sqrt{2 \pi \sigma^2}} e^{-\sfrac{(t - \lambda_{(i)})^2}{2 \sigma^2}},
    \label{equ:gaussian-kernel-eigenvalues}
\end{equation}
where $\lambda_{(1)}, \dots, \lambda_{(n)}$ denote the eigenvalues of $\mtx{A}$ sorted by increasing distance from spectral-parameter $t$, such that $\sigma_1(t) \geq \dots \geq \sigma_n(t)$. Consequently, by using the closed-form expression of the eigenvalues \refequ{equ:gaussian-kernel-eigenvalues}, we may generously upper bound the numerical rank by
\begin{equation}
    r_{\varepsilon}(g_{\sigma}(t\mtx{I}_n - \mtx{A}) / n) \leq \#\left\{1\leq i\leq n: |t - \lambda_i| < d_{\varepsilon, \sigma} \right\},
    \label{equ:gaussian-kernel-numerical-rank}
\end{equation}
with $d_{\varepsilon, \sigma} = \sigma \sqrt{-2 \log(\sqrt{2 \pi} \sigma \varepsilon)}$ and the eigenvalues $\lambda_1, \dots, \lambda_n$ of $\mtx{A}$ and the constants. The expression \refequ{equ:gaussian-kernel-numerical-rank} has a nice visual interpretation: The numerical rank of a matrix is at most equal to the number of eigenvalues which are closer to the parameter $t$ than $d_{\varepsilon, \sigma}$. This is illustrated in \reffig{fig:numerical-rank}.
\begin{figure}[ht]
    \centering
    \input{figures/numerical-rank.tex}
    \caption{The numerical rank of $g_{\sigma}(t\mtx{I}_n - \mtx{A}) / n$ can be approximately computed by counting the number of eigenvalues $\lambda_{(1)}, \dots, \lambda_{(n)}$ of the matrix $\mtx{A}$ which lie less than a constant $d_{\varepsilon, \sigma}$ away from $t$.}
    \label{fig:numerical-rank}
\end{figure}

If we additionally assume the eigenvalues of the matrix $\mtx{A}$ to be evenly distributed within $[-1, 1]$, that is, in any subinterval of fixed length in $[a, b]$, $-1 \leq a < b \leq 1$, we can expect to find roughly the same number of eigenvalues, then we can estimate the numerical rank of $g_{\sigma}(t\mtx{I}_n - \mtx{A}) / n$ to be
\begin{equation}
    r_{\varepsilon}(g_{\sigma}(t\mtx{I}_n - \mtx{A}) / n) \lessapprox n d_{\varepsilon, \sigma} \approx 6 n \sigma,
    \label{equ:gaussian-kernel-numerical-rank-uniform}
\end{equation}
provided we set $\varepsilon = 10^{-16}$ and choose $\sigma$ roughly between $10^{-8}$ and $1$. Hence, choosing $n_{\mtx{\Omega}} = \mathcal{O}(n \sigma)$ will yield approximations which can be significantly better than \refthm{thm:chebyshev-nystrom}.