### 2025-02-20 paper/re-pro-badge.tex
automatically reproduce paper

The code for our numerical experiments can be found on \url{https://github.com/FMatti/Rand-TRACE}. The figures and tables were generated from the commit \href{https://github.com/FMatti/Rand-TRACE/tree/7e1dc43}{7e1dc43} on 2025-02-20 at 14:53:12 UTC.