### 2025-02-20 paper/conclusion.tex
polish results and add conclusion

\section{Conclusion}
\label{sec:conclusion}

We have analyzed three stochastic trace estimators for constant symmetric matrices $\mtx{B} \in \mathbb{R}^{n \times n}$ when they are applied straightforwardly to a parameter-dependent symmetric matrix $\mtx{B}(t) \in \mathbb{R}^{n \times n}$. We have showed that in the $L^1$-norm, these estimators satisfy bounds which parallel existing results for constant matrices. The results for the estimators allow us to analyze the Nyström-Chebyshev++ method, an efficient technique for approximating the spectral density of a symmetric matrix. Further, we have proposed multiple improvements, both algorithmic and theoretical, to an existing implementation of this method.