### 2025-02-21 paper/spectraldensity.tex
review remainder of section 3

\section{Application to spectral density approximation}
\label{sec:application}

We now get back to the task of computing the spectral density $1/n \sum_{i=1}^{n} \delta(t - \lambda_i)$ of a real symmetric matrix $\mtx{A} \in \mathbb{R}^{n \times n}$ with eigenvalues $\lambda_1, \dots, \lambda_n$, where $\delta(\cdot)$ denotes the Dirac delta distribution. As explained in the introduction, finding accurate approximations to this distribution is challenging, and we will therefore work with
a smoothed version. %This still yields enough information
%. Furthermore, in most applications the exact location of each individual eigenvalue is not important, but more so their approximate locations relative to each other, such as eigenvalue clusters, spectral gaps, or outliers. In the following we will work with a smoothed version of the spectral density, which is significantly easier to approximate at the cost of losing some of the finer characteristics of the spectrum.

\subsection{Smoothed spectral density}
\label{subsec:spectral-density}

We recall that the smoothed spectral density $\phi_{\sigma}$ is defined as
\begin{equation}
    \phi_{\sigma}(t) = \frac{1}{n} \sum_{i=1}^{n} g_{\sigma}(t - \lambda_i) = \frac{1}{n} \Trace(g_{\sigma}(t\mtx{I}_n - \mtx{A}))
    \label{equ:smooth-spectral-density}
\end{equation}
for a smoothing kernel $g_{\sigma}$ parametrized by a smoothing parameter $\sigma > 0$. In this work, we choose a Gaussian smoothing kernel of width $\sigma > 0$:
\begin{equation}
    g_{\sigma}(s) = \frac{1}{\sigma \sqrt{2\pi}} e^{-\sfrac{s^2}{2\sigma^2}}.
    \label{equ:smoothing-kernel}
\end{equation}
The expression~\refequ{equ:smooth-spectral-density} gives rise to the task of estimating the trace of the parameter-dependent SPSD matrix $g_{\sigma}(t\mtx{I}_n - \mtx{A})$, which has been analyzed in \refsec{sec:analysis}.

Choosing the smoothing parameter $\sigma$ results in the following trade off: A larger $\sigma$ usually makes it  easier to approximate the matrix function $g_{\sigma}(t\mtx{I}_n - \mtx{A})$, whereas a smaller $\sigma$ allow one to stay closer to the original spectral density $\phi$. To quantify the latter, one can measure the error between $\phi$ and its smoothened version $\phi_{\sigma}$ by
\begin{equation}
    \sup_{f \in \mathcal{S}} \int_{a}^{b} f(t) (\phi(t) - \phi_{\sigma}(t))~\mathrm{d}t,
    \label{equ:error-metric}
\end{equation}
where $\mathcal{S}$ represents an appropriately chosen space of test functions. Among others, the reference~\cite{lin-2016-approximating-spectral} uses $\mathcal{S} = \{ f: f(t) \equiv g_{\sigma}(t - s), s \in [a, b]\}$, \cite{chen-2021-analysis-stochastic} uses $\mathcal{S} = \{f : f(t) = \Theta(s - t), s \in [a, b] \}$ with the Heaviside step function $\Theta$, and \cite{braverman-2022-sublinear-time,bhattacharjee-2024-improved-spectral} use $\mathcal{S} = \{f : |f(t) - f(s)| \leq |t - s| \}$ for which the error \refequ{equ:error-metric} is equivalent to the Wasserstein-1 distance between $\phi$ and $\phi_{\sigma}$.
When choosing 
the last metric,
assuming that the eigenvalues are somewhat uniformly distributed and $\sigma$ is rather small, a simple calculation allows one to conclude that the smoothing error \refequ{equ:error-metric} is of the order $\sigma$. Consequently, to ensure that the approximation deviates by at most a factor of $\varepsilon > 0$, one needs to choose $\sigma \approx \varepsilon$.

\subsection{Chebyshev approximation of smoothing kernel}
\label{subsec:chebyshev-approximation}

To estimate the trace in \refequ{equ:smooth-spectral-density} efficiently, we approximate the function 
$s\mapsto g_{\sigma}(t- s)$ by a
Chebyshev approximation $g_{\sigma}^{(m)}(t-s) = 
\sum_{l=0}^{m} \mu_l(t) T_l(s)$
with the approximation coefficients $\mu_l(t) \in \mathbb R$ and the Chebyshev polynomials
$T_l(s) = \cos(l \cdot \arccos(s))$ for $l = 0,1,\ldots,m$. We choose the coefficients $\mu_l(t)$ in such a way that $g_{\sigma}^{(m)}(t-s)$ interpolates $g_{\sigma}(t-s)$ at the Chebyshev nodes $s_i = \cos(\pi i / m),~i=0,1,\dots,m$. In this case, $g_{\sigma}^{(m)}$ is often referred to as the Chebyshev interpolant~\cite[Chapter 2]{trefethen-2020-approximation-theory}.
Assuming that the spectrum of $\mtx{A}$ is contained in $[-1,1]$, this gives rise to
%need to evaluate the involved matrix function $g_{\sigma}(t\mtx{I}_n - \mtx{A})$. However, doing this exactly would require us to first diagonalize $\mtx{A}$, which is often a prohibitively expensive operation. For algorithmic reasons (see \refsec{subsubsec:chebyshev-nystrom-implementation}), we seek for approximations of this matrix function in terms of affine linear approximations, of which the was determined to be most suitable due to its approximation error guarantees and ease of manipulation. In every $t \in \mathbb{R}$ we expand 
\begin{equation}
    g_{\sigma}(t\mtx{I}_n - \mtx{A}) \approx g_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A}) = \sum_{l=0}^{m} \mu_l(t) T_l(\mtx{A}).
    \label{equ:matrix-approximation}
\end{equation}
Inserting~\refequ{equ:matrix-approximation} into~\refequ{equ:smooth-spectral-density} yields the approximate smoothed spectral density
\begin{equation}
    \phi_{\sigma}(t) \approx \phi_{\sigma}^{(m)}(t) =  \frac{1}{n} \Trace(g_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A})).
    \label{equ:expanded-spectral-density}
\end{equation}
\begin{remark}
For a general symmetric matrix $\mtx{A}$, one first estimates an interval $[a,b]$ containing the spectrum of $\mtx{A}$~\cite{zhou-2011-bounding-spectrum}. The usual affine transform
$\tau(t) = \frac{2t - a - b}{b - a}$
then lets us define a matrix $\bar{\mtx{A}} = \tau(\mtx{A})$ for which the spectrum is contained in $[-1, 1]$ and, hence, the approximation \refequ{equ:matrix-approximation} is sensible.\footnote{Note that one has to adjust the smoothing parameter $\bar{\sigma} = 2 \sigma (b - a)^{-1}$ to obtain an undistorted approximation of the original spectrum when inverting the transformation.} This allows to assume, from now on and without loss of generality, that the spectrum of $\mtx{A}$ is contained in $[-1, 1]$.
\end{remark}

\subsubsection{Approximation error}
\label{subsubsec:approximation-error}

We now analyze the impact of the Chebyshev approximation of the Gaussian smoothing kernel $g_{\sigma}$ in the approximation of the spectral density $\phi_{\sigma}$. We start with a standard error estimate concerning the approximation of $g_{\sigma}$.

\begin{lemma}{Chebyshev approximation error}{chebyshev-error}
    The Chebyshev approximation $g_{\sigma}^{(m)}$ of the Gaussian smoothing kernel $g_{{\sigma}}$ from  \refequ{equ:smoothing-kernel} satisfies
    \begin{equation}
        \sup_{s,t \in [-1, 1]} \big| g_{\sigma}(t - s) - g_{\sigma}^{(m)}(t - s) \big| \leq 2\sqrt{\frac{2e}{\pi}} \frac{1}{\sigma^2} (1 + \sigma)^{-m} =: E_{\sigma, m}
        \label{equ:chebyshev-interpolation-sup-error-kernel}
        %\lVert  \phi_{\sigma} - \phi_{\sigma}^{(m)} \rVert _1 &\leq \frac{2\sqrt{2}}{\sigma^2} (1 + \sigma)^{-m}.
        %\label{equ:2-chebyshev-interpolation-error}
    \end{equation}
    for every degree $m \in \mathbb{N}$ and smoothing parameter $\sigma > 0$.
\end{lemma}
\begin{proof}
    The proof of this result is standard. % and follows the one given in the arXiv preprint of \cite[Theorem 2]{lin-2017-randomized-estimation}, but attains a slightly different dependence on $\sigma$. 
    Let $\mathcal{E}_{\chi} \subset \mathbb{C}$ denote the Bernstein ellipse  with foci $\{-1, +1\}$ and the sum of semi-axes equal to $\chi > 1$. By\cite[Theorem 8.2]{trefethen-2020-approximation-theory},
    \begin{equation}
        \sup_{s \in [-1, 1]} \big| g_{\sigma}(t - s) - g_{\sigma}^{(m)}(t - s) \big| \leq \frac{4}{\chi^m (\chi - 1)} \max_{z \in \mathcal{E}_{\chi}} |g_{\sigma}(t - z)|,
        \label{equ:bernstein-bound}
    \end{equation}
    where used the analyticity of $g_{\sigma}(t- \cdot)$.
    For $z = x + \mathrm{i} y \in \mathcal{E}_{\chi}$, we have
    \begin{equation}
    | g_{\sigma}(t - z) | 
    = \frac{1}{\sigma \sqrt{2 \pi}} \big| e^{-\sfrac{(t - z)^2}{2\sigma^2}} \big|
    = \frac{1}{\sigma \sqrt{2 \pi}} e^{-\sfrac{(t - x)^2}{2\sigma^2}}e^{\sfrac{y^2}{2\sigma^2}}
    \le \frac{1}{\sigma \sqrt{2 \pi}} e^{\sfrac{(\chi - \chi^{-1})^2}{8 \sigma^2}},
    %\leq \frac{1}{n \sigma \sqrt{2 \pi}} \max_{x + iy \in \mathcal{E}_{\chi}} e^{\sfrac{y^2}{2\sigma^2}} 
    \end{equation}
    where the last inequality uses 
    $e^{-\sfrac{(t - x)^2}{2\sigma^2}} \leq 1$ and $|y| \le (\chi - \chi^{-1}) / 2$.
    Choosing $\chi = 1 + \sigma$ and using $\chi - \chi^{-1} \leq 2\sigma$ completes the proof of~\refequ{equ:chebyshev-interpolation-sup-error-kernel}.
\end{proof}

For $0 < \sigma \ll 1$, \reflem{lem:chebyshev-error} implies that choosing a degree $m = \mathcal{O}(\sigma^{-1}(\log(\varepsilon^{-1}) + \log(\sigma^{-1})))$ ensures an error $\mathcal{O}(\varepsilon)$. The lemma also allows us to bound the $L^1$-error of the spectral density approximation
$\phi_{\sigma}^{(m)}$ defined in~\refequ{equ:expanded-spectral-density}:
\begin{align}
    \int_{-1}^{1} \left| \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) \right|~\mathrm{d}t 
    &= \int_{-1}^{1} \left| \frac{1}{n} \sum_{i=1}^n \big(g_{\sigma}(t - \lambda_i) - g_{\sigma}^{(m)}(t - \lambda_i)\big) \right|~\mathrm{d}t \notag \\
    &\leq \int_{-1}^{1} \max_{i = 1, \dots, n} \big| g_{\sigma}(t - \lambda_i) - g_{\sigma}^{(m)}(t - \lambda_i) \big|~\mathrm{d}t    \leq 2 E_{\sigma, m}.
    \label{equ:chebyshev-interpolation-spectral-density}
\end{align}

%\begin{theorem}{$L^1$-error of Chebyshev approximation of spectral density}{chebyshev-error}
%    Let $\phi_{\sigma}$ be the smoothed spectral density of a symmetric matrix $\mtx{A} \in \mathbb{R}^{n \times n}$ whose spectrum is contained in $[-1, 1]$. Then the spectal density with Chebyshev approximation $\phi_{\sigma}^{(m)}$ satisfies
%    \begin{equation}
%        \int_{-1}^{1} \left| \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) \right|~\mathrm{d}t \leq 2 E_{\sigma, m}
%        \label{equ:chebyshev-interpolation-spectral-density}
%        %\lVert  \phi_{\sigma} - \phi_{\sigma}^{(m)} \rVert _1 &\leq \frac{2\sqrt{2}}{\sigma^2} (1 + \sigma)^{-m}.
%        %\label{equ:2-chebyshev-interpolation-error}
%    \end{equation}
%    for all degrees $m \in \mathbb{N}$, smoothing parameters $\sigma > 0$, and the error $E_{\sigma, m}$ from \reflem{lem:chebyshev-error}.%, and $t \in \mathbb{R}$. %The same result holds for the approximation of a non-negative Chebyshev approximation of \emph{even} degree $m \in \mathbb{N}$ but with $\underline{E}_{\sigma, m}$.
%\end{theorem}
%
%\begin{proof}
%    We observe that for all $t \in [-1, 1]$
%    \begin{align}
%        &\left| \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) \right| \notag \\
%        &= \left| \frac{1}{n} \Trace(g_{\sigma}(t\mtx{I}_n - \mtx{A})) - \frac{1}{n} \Trace(g_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A})) \right|
%        && \text{(definitions \refequ{equ:smooth-spectral-density} and \refequ{equ:expanded-spectral-density})} \notag \\
%        &= \left| \frac{1}{n} \sum_{i=1}^n \left(g_{\sigma}(t - \lambda_i) - g_{\sigma}^{(m)}(t - \lambda_i)\right) \right|
%        && \text{($\lambda_1, \dots, \lambda_n$ eigenvalues of $\mtx{A}$)} \notag \\
%        &\leq \max_{i = 1, \dots, n} \left| g_{\sigma}(t - \lambda_i) - g_{\sigma}^{(m)}(t - \lambda_i) \right|
%        && \text{(conservative upper bound)} \notag \\
%        &\leq E_{\sigma, m}.% ~ \text{(or $\underline{E}_{\sigma, m}$)}.
%        && \text{(using \reflem{lem:chebyshev-error})} \notag \\
%    \end{align}
%    Then, Hölder's inequality allows us to conclude
%    \begin{equation}
%        \int_{-1}^{1} | \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) | ~\mathrm{d}t
%        \leq 2 \sup_{t \in [-1, 1]} \left| \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) \right|
%        \leq 2 E_{\sigma, m}.% ~ \text{(or $2 \underline{E}_{\sigma, m}$)}.
%    \end{equation}
%\end{proof}

\subsubsection{Preservation of non-negativity} \label{sec:preservnonneg}

In general, the Chebyshev approximation $g_{\sigma}^{(m)}$ does not inherit the  non-negativity of the smoothing kernel $g_{\sigma}$. Consequently, the matrix $g_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A})$ may become 
(slightly) indefinite. This causes issues when attempting to apply the Nyström++ trace estimator (\refalg{equ:nystrompp-trace-estimator}). While we have never observed this in our numerical experiments, the Nyström approximation may fail when the SPSD condition is not met; see also the discussion in~\cite{nakatsukasa-2023-randomized-lowrank}. On the theoretical side,~\refthm{thm:nystrom-pp}
critically relies on the SPSD condition.

To ensure a non-negative Chebyshev approximation, we suggest the following simple method: Given an even degree $m$, approximate the square root of the smoothing kernel up to degree $m/2$:
\begin{equation}
    \sqrt{g_{\sigma}}(t - s) \approx \sqrt{g_{\sigma}}^{(\sfrac{m}{2})}(t - s) = \sum_{l=0}^{m/2} \xi_l(t) T_l(s).
    \label{equ:square-root-chebyshev-approximation}
\end{equation}
When squaring this approximation, one obtains a polynomial of degree $m$, which admits an exact approximation
of the form 
$(\sum_{l=0}^{m/2} \xi_l(t) T_l(s))^2 = \sum_{l=0}^{m} \underline{\mu}_l(t) T_l(s)$.
Thus, we obtain the non-negative Chebyshev approximation
\begin{equation}
g_{\sigma}(t - s) \approx    \underline{g}_{\sigma}^{(m)}(t - s) = \sum_{l=0}^{m} \underline{\mu}_l(t) T_l(s).
    \label{equ:non-negative-chebyshev-approximation}
\end{equation}
%With a complexity of $\mathcal{O}(m \log(m))$ this procedure is computationally inexpensive and allows for very similar convergence bounds to \reflem{lem:chebyshev-error}, which we will derive \textcolor{red}{in the following?} in \refsec{subsubsec:chebyshev-nystrom-analysis}.
%To this extent, we will first analyze the non-negative Chebyshev approximation.
The corresponding non-negative approximation of the smoothed spectral density,
\begin{equation}
    \phi_{\sigma}(t) \approx \underline{\phi}_{\sigma}^{(m)}(t) := \frac{1}{n} \Trace(\underline{g}_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A})),
    \label{equ:non-negative-expanded-density}
\end{equation}
satisfies the following bound.

\begin{lemma}{$L^1$-error of non-negative approximate smooth spectral density}{non-negative-chebyshev-error}
Let $\phi_{\sigma}$ be the smoothed spectral density defined in~\refequ{equ:smooth-spectral-density}
for a symmetric matrix $\mtx{A} \in \mathbb{R}^{n \times n}$ with spectrum contained in $[-1, 1]$. Then the approximation $\underline{\phi}_{\sigma}^{(m)}$ defined in~\refequ{equ:non-negative-expanded-density} satisfies
    \begin{equation}
        \int_{-1}^{1} \big| \phi_{\sigma}(t) - \underline{\phi}_{\sigma}^{(m)}(t) \big|~\mathrm{d}t \leq 4 \sqrt{2} \left(1 + \sigma \sqrt{\pi} \cdot E_{\sqrt{2}\sigma, \sfrac{m}{2}}\right) E_{\sqrt{2}\sigma, \sfrac{m}{2}} =: \underline{E}_{\sigma, m}
        \label{equ:chebyshev-interpolation-sup-error-kernel-nonneg}
        %\lVert  \phi_{\sigma} - \phi_{\sigma}^{(m)} \rVert _1 &\leq \frac{2\sqrt{2}}{\sigma^2} (1 + \sigma)^{-m}.
        %\label{equ:2-chebyshev-interpolation-error}
    \end{equation}
    for every \emph{even} degree $m \in \mathbb{N}$, smoothing parameter $\sigma > 0$, and with $E_{\sigma, m}$ defined in \reflem{lem:chebyshev-error}.
\end{lemma}
\begin{proof}
    Using that the inequality $| a - b | \leq (2 \sqrt{a} + | \sqrt{a} - \sqrt{b} |)  | \sqrt{a} - \sqrt{b} |$ holds for any non-negative numbers $a, b$, it follows that
    \begin{equation}
        \big| g_{\sigma} - \underline{g}_{\sigma}^{(m)} \big| \leq \big( 2 \big| \sqrt{g_{\sigma}} \big| + \big| \sqrt{g_{\sigma}} - \sqrt{g_{\sigma}}^{(\sfrac{m}{2})} \big| \big) \big| \sqrt{g_{\sigma}} - \sqrt{g_{\sigma}}^{(\sfrac{m}{2})} \big|.
    \end{equation}
    Because of $\sqrt{g_{\sigma}} = \sqrt{2 \sigma \sqrt{2 \pi}} \cdot g_{\sqrt{2}\sigma}$,
we can apply \reflem{lem:chebyshev-error} with the substitutions
    $\sigma \gets \sqrt{2}\sigma$ and $m \gets m/2$ to bound 
    $\big| \sqrt{g_{\sigma}} - \sqrt{g_{\sigma}}^{(\sfrac{m}{2})} \big|$.
    Together with  $|\sqrt{g_{\sigma}}| \leq 1/\sqrt{\sigma \sqrt{2 \pi}}$, we get
    \begin{align}
        \big| g_{\sigma}(t - s) - \underline{g}_{\sigma}^{(m)}(t - s) \big| \notag 
        &\leq \Big( 2 \frac{1}{\sqrt{ \sigma \sqrt{2\pi}}} + \sqrt{2 \sigma \sqrt{2 \pi}} \cdot E_{\sqrt{2}\sigma, \sfrac{m}{2}}\Big)\sqrt{2 \sigma \sqrt{2 \pi}} \cdot E_{\sqrt{2}\sigma, \sfrac{m}{2}} \notag \\
        &= 2\sqrt{2} \big(1 + \sigma \sqrt{\pi} \cdot E_{\sqrt{2}\sigma, \sfrac{m}{2}}\big) E_{\sqrt{2}\sigma, \sfrac{m}{2}}
    \end{align}
    for every $s,t \in [-1, 1]$.
    Bounding the integral as in~\refequ{equ:chebyshev-interpolation-spectral-density} then yields the desired result.
\end{proof}

As the non-negative approximation~\refequ{equ:non-negative-chebyshev-approximation} ``wastes'' half of the degrees of freedom, it can be expected to be less accurate than the standard Chebyshev approximation. On the other hand, $\sqrt{g_{\sigma}} \propto g_{\sqrt{2}\sigma}$ is somewhat easier to approximate because it is less \enquote{pointy} compared to $g_{\sigma}$. Overall, it turns out that one needs to choose the degree of the non-negative approximation roughly a factor $\sqrt{2}$ larger to match the approximation error of the standard Chebyshev approximation (see also \reffig{fig:interpolation-issue}).

%In the following, we will show how the coefficients $\underline{\mu}_0(t), \dots, \underline{\mu}_m(t)$ of the non-negative Chebyshev approximation \refequ{equ:non-negative-chebyshev-approximation} can efficiently be computed and that this approximation exhibits similar convergence properties to the standard Chebyshev approximation.
%\todo{
%Maybe mention indefinite Nyström [Nakatsukasa], which won't give us desirable guarantees, and refinement of Chebyshev approximation [Francis Bach], which only slightly improves, but is much harder to implement. Side-note: it is possible to refine the coefficients of the Chebyshev approximation to get a better approximation in the supremum-norm over the set of all non-negative polynomials \cite%{fejer-1916-uber-trigonometrische}. The non-convex nature of this problem and the relatively small improvement we get with this technique made us decide to not pursue this further.
%}

%\subsection{Computation of approximation coefficients}

\subsubsection{Computation of Chebyshev coefficients}
\label{subsubsec:dct}

%\textcolor{red}{??? I don't understand this derivation. Don't you think that~\refequ{equ:discrete-cosine-transform} only holds approximately for the exact $g_\sigma$ and it should be something like $g_\sigma \approx g^{(m)} = ...$. I think there is a major confusion of two different approaches here:
%\begin{itemize}
% \item The Chebyshev \emph{approximation} of $f$ is the $\ell^2$-orthogonal projection of $f$ on the first $m+1$ Chebyshev polynomials. The coefficients are computed from $\ell^2$ inner products between the Chebyshev polynomials and $f$. This is probably what Lin is doing. 
% \item The Chebyshev \emph{approximation} used by Chebfun and possibly others proceeds differently. It first replaces (or rather pretends to replace) $f$ by the interpolation of $f$ in (some) Chebyshev nodes. This interpolation admits an exact truncated approximation and the coefficients can be computed from a Fourier transform (or even DCT?) of the interpolation data. This might match what is written below, but I am not sure whether the interpolation nodes are the same. Please check with Trefethen's book.
%\end{itemize}
%There are two issues now
%\begin{enumerate}
% \item I am now unsure how $g_\sigma^{(m)}$ is precisely defined. This should be spelled out clearer in the beginning of Section 3.2.
% \item There could be a mismatch between the $g_\sigma^{(m)}$ of Lemma 3.1 and the one we are actually using. This mismatch can most likely be resolved, because a result of the form of Lemma 3.1 also holds for the Chebyshev interpolation. If it is easy to resolve, please do so. If not, at least point this out here. Have a look at Trefethen's book; it might have the right error bound.
%\end{enumerate}
%???
It is well known that the coefficients of the Chebyshev approximation \refequ{equ:matrix-approximation} can --- in exact arithmetic --- be computed by inverting the (type I) discrete cosine transform (DCT)
\begin{equation}
    g_{\sigma}(t - \cos(s_i)) = \sum_{l=0}^{m} \mu_l(t) \cos\left(\frac{\pi i l}{m} \right)
    \label{equ:discrete-cosine-transform}
\end{equation}
between the approximation coefficients $\mu_l(t),~l=0,\dots,m$ and the function evaluations at the Chebyshev points $s_i = \cos(\pi i / m),~i=0,\dots,m$ \cite{baszenski-1997-fast-polynomial, plonka-2018-numerical-fourier, trefethen-2020-approximation-theory}. This can be accomplished in $\mathcal{O}(m \log(m))$ operations.%}

A Chebyshev approximation can be squared in $\mathcal{O}(m \log(m))$ operations by using the DCT and its inverse \cite{baszenski-1997-fast-polynomial}; see \refalg{alg:chebyshev-squaring}. This allows one to efficiently compute the non-negative Chebyshev approximation \refequ{equ:non-negative-chebyshev-approximation}, and it will find further use in \refsec{subsec:chebyshev-nystrom}.

\begin{algo}{Fast squaring of Chebyshev approximations}{chebyshev-squaring}
    \begin{algorithmic}[1]
        \Statex \textbf{Input:} Coefficients $\{ \mu_l \}_{l=0}^{m}$ of Chebyshev approximation $\sum_{l=0}^{m} \mu_l T_l$
        \Statex \textbf{Output:} Coefficients $\{ \nu_l \}_{l=0}^{2m}$ such that $\sum_{l=0}^{2m} \nu_l T_l = (\sum_{l=0}^{m} \mu_l T_l)^2$
        \State By appending $m$ zeros, define vector $\vct{\mu} = [\mu_0, \ldots, \mu_m,0,\ldots,0 ] \in \mathbb{R}^{2m + 1}$.
    \State Compute $\vct{f} = \operatorname{DCT}(\vct{\mu})$
    \State Compute $[\nu_0, \dots, \nu_{2m}] = \operatorname{DCT}^{-1}(\vct{f} \odot \vct{f})$
    \label{lin:inverse-DCT}
    \end{algorithmic}
\end{algo}
On \reflin{lin:inverse-DCT} of \refalg{alg:chebyshev-squaring}, the symbol $\odot$ denotes the entrywise product.

%\textcolor{red}{??? To be modified after the notion of Chebyshev approximation is clarified ???
In \cite{lin-2017-randomized-estimation}, a slightly different approach to approximating $g_{\sigma}$ is taken: Instead of Chebyshev interpolation,
the kernel $g_{\sigma}$ is approximated by a Chebyshev series~\cite[Chapter 4]{trefethen-2020-approximation-theory} truncated after term $m$. In constrast to interpolation, the expansion coefficents involve integrals, which need to be approximated by quadrature. By choosing $2m+1$ Chebyshev quadrature nodes, this approximation can be carried using the fast Fourier transform (FFT). Although interpolation and truncated expansion satisfy nearly identical convergence bounds~\cite[Theorem 8.2]{trefethen-2020-approximation-theory}, we observed the procedure from~\cite{lin-2017-randomized-estimation} to be significantly slower due to need of applying the FFT (instead of DCT) to longer vectors. See~\reftab{tab:chebyshev-timing-interpolation}, where we compare a benevolent implementation of \cite[Algorithm 1]{lin-2017-randomized-estimation} with our approach, both for the standard and the non-negative Chebyshev approximation.%}

\begin{table}[ht]
    \caption{Runtime in milliseconds for computing the coefficients of a degree-$m$ Chebyshev approximation to the smoothing kernel $g_{\sigma}$ in three different ways. \emph{FFT}: Using~\cite[Algorithm 1]{lin-2017-randomized-estimation}.  \emph{DCT}: Applying the discrete cosine transform directly to $g_\sigma$.  \emph{Non-negative DCT}: Applying the discrete cosine transform to $\sqrt{g_\sigma}$ together with \refalg{alg:chebyshev-squaring}. We use $\sigma=0.005$, $n_t=1000$ parameter values, and various degrees $m$. We average over 7 runs of the algorithms and repeat this $1000$ times to form the mean and standard deviation reported below.}
    \label{tab:chebyshev-timing-interpolation}
   \input{tables/interpolation.tex}
\end{table}

\subsection{Chebyshev-Nyström++ method for spectral density approximation}
\label{subsec:chebyshev-nystrom}

Combining the results of \refsec{sec:analysis} and \refsec{subsec:chebyshev-approximation}, we apply the Nyström++ estimator $\Nystrpp{\mtx{\Omega}}{\mtx{\Psi}}$ from~\refequ{equ:nystrompp-trace-estimator}
to the non-negative 
Chebyshev approximation, which yields the approximation
\begin{equation}
    \phi_{\sigma}(t) \approx \widetilde{\underline{\phi}}_{\sigma}^{(m)}(t) := \frac{1}{n} \Nystrpp{\mtx{\Omega}}{\mtx{\Psi}}(\mtx{B}(t)), \quad \mtx{B}(t) = \underline{g}_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A}).
    \label{equ:chebyshev-nystrom-formula}
\end{equation}

\subsubsection{Implementation}
\label{subsubsec:chebyshev-nystrom-implementation}

One usually aims at evaluating~\refequ{equ:chebyshev-nystrom-formula} for many values of $t$. In the following, we explain how this can be done efficiently by exploiting 
the well-known three-term recurrence relation of Chebyshev polynomials:
\begin{equation}
    T_0(s) = 1,\quad  T_1(s) = s,\quad 
    T_l(s) = 2 s T_{l-1}(s) - T_{l-2}(s) \text{ for $l \geq 2$.}
    \label{equ:chebyshev-recurrence}
\end{equation}

We start with the first term in the definition~\refequ{equ:nystrompp-trace-estimator} of $\Nystrpp{\mtx{\Omega}}{\mtx{\Psi}}$, which requires evaluating the trace of the 
Nyström approximation \refequ{equ:nystrom-approximation}:
\begin{equation}
    \frac{1}{n} \Trace \big((\mtx{B}(t) \mtx{\Omega}) (\mtx{\Omega}^{\top} \mtx{B}(t) \mtx{\Omega})^{\dagger} (\mtx{B}(t) \mtx{\Omega})^{\top} \big) =
    \frac{1}{n} \Trace \big((\underbrace{\mtx{\Omega}^{\top} \mtx{B}(t) \mtx{\Omega}}_{=: \mtx{K}_1(t)})^{\dagger} (\underbrace{\mtx{\Omega}^{\top} \mtx{B}(t)^2 \mtx{\Omega}}_{=: \mtx{K}_2(t)}) \big).
    %\frac{1}{n} \Trace((\mtx{\Omega}^{\top} \underline{g}_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A}) \mtx{\Omega})^{\dagger} (\mtx{\Omega}^{\top} \underline{g}_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A})^2 \mtx{\Omega})),
    \label{equ:cyclic-property}
\end{equation}
where we used the invariance of the trace under cyclic permutations.
Inserting the Chebyshev approximation  $\mtx{B}(t) = \underline{g}_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A})$ gives
\begin{equation}
    \mtx{K}_1(t) = \mtx{\Omega}^{\top} \underline{g}_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A}) \mtx{\Omega} = \sum_{l=0}^{m} \underline{\mu}_l(t) (\mtx{\Omega}^{\top} T_l(\mtx{A}) \mtx{\Omega}).
\end{equation}
Observing that $\mtx{\Omega}^{\top} T_l(\mtx{A}) \mtx{\Omega}$ is independent of $t$ allows us to precompute these (small) $n_{\mtx{\Omega}} \times n_{\mtx{\Omega}}$ matrices for $l=0, \dots, m$ using the Chebyshev recurrence \refequ{equ:chebyshev-recurrence}. For each value of $t$, we then only need to sum these matrices (multiplied by the corresponding coefficient $\underline{\mu}_l(t)$).

The squaring of Chebyshev approximations discussed in \refsec{sec:preservnonneg} allows us to express $\mtx{B}(t)^2 = \underline{g}_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A})^2 = \sum_{l=0}^{2m} \underline{\nu}_l(t) T_l(\mtx{A})$, where the coefficients $\underline{\nu}_l(t)$ can be cheaply computed from $\underline{\mu}_l(t)$ using~\refalg{alg:chebyshev-squaring}. Hence, if we precompute
$\mtx{\Omega}^{\top} T_l(\mtx{A}) \mtx{\Omega}$
also for $l=m+1, \dots, 2m$, this allows us to evaluate $\mtx{K}_2(t)$ cheaply, in an analogous way to $\mtx{K}_1(t)$. Alternatively, \cite{lin-2017-randomized-estimation} proposes to instead approximate the squared smoothing kernel $g_{\sigma}^2$ in a separate "auxiliary" Chebyshev series truncated after term $2m$. While this direct approximation of $g_{\sigma}^2$ might be more accurate than our squaring approach, it is \emph{inconsistent} in the sense that numerically computing the truncated Chebyshev series of $g_{\sigma}^2$ gives --- in general --- not the same result as squaring the truncated Chebyshev series of $g_{\sigma}$. We have observed that meticulously ensuring consistency between the approximations of $g_{\sigma}$ and $g_{\sigma}^2$ makes a visible difference in terms of accuracy (cf. \reffig{fig:interpolation-issue}). This consistency will also allow us to derive bounds for the estimator \refequ{equ:chebyshev-nystrom-formula}.

\begin{figure}[ht]
    \centering
    \input{plots/interpolation.pgf}
    \caption{Accuracy when computing the estimate \refequ{equ:cyclic-property} in three different ways. \emph{Inconsistent}: Applying an auxiliary truncated Chebyshev series to $g_{\sigma}^2$, as proposed in~\cite{lin-2017-randomized-estimation}. \emph{Consistent}: Squaring the Chebyshev approximation $g_{\sigma}^{(m)}$ using \refalg{alg:chebyshev-squaring}. \emph{Non-negative consistent}: Squaring the non-negative Chebyshev approximation $\underline{g}_{\sigma}^{(m)}$ with~\refalg{alg:chebyshev-squaring}. We use a \textcolor{red}{??? MORE DETAIL: At least, say what is $N$. Same in Figure 3.2 ???} coarse discretization of the Hamiltonian defined in~\refequ{equ:5-experiments-electronic-hamiltonian} in two dimensions, fix $n_{\mtx{\Omega}} = 80$, $\sigma = 0.005$, and compute the $L^1$ error from the approximation of the spectral density at $n_t = 100$ uniformly spaced values of $t$ in the interval $[-1, 1]$.}
    \label{fig:interpolation-issue}
\end{figure}

To treat the second term in the definition~\refequ{equ:nystrompp-trace-estimator} of $\Nystrpp{\mtx{\Omega}}{\mtx{\Psi}}$, we observe that 
\begin{align}
    &\frac{1}{n n_{\mtx{\Psi}}} \Trace \big(\mtx{\Psi}^{\top} (\mtx{B}(t) - (\mtx{B}(t) \mtx{\Omega}) (\mtx{\Omega}^{\top} \mtx{B}(t) \mtx{\Omega})^{\dagger} (\mtx{B}(t) \mtx{\Omega})^{\top}) \mtx{\Psi}^{\top} \big) \notag \\
    &= \frac{1}{n n_{\mtx{\Psi}}} \big( \underbrace{\Trace(\mtx{\Psi}^{\top} \mtx{B}(t)\mtx{\Psi})}_{=: \ell(t)} - \Trace((\underbrace{\mtx{\Psi}^{\top} \mtx{B}(t) \mtx{\Omega}}_{=: \mtx{L}_1(t)^{\top}}) (\underbrace{\mtx{\Omega}^{\top} \mtx{B}(t) \mtx{\Omega}}_{= \mtx{K}_1(t)^{\top}})^{\dagger} (\underbrace{\mtx{\Omega}^{\top} \mtx{B}(t) \mtx{\Psi}^{\top}}_{=: \mtx{L}_1(t)})) \big).
    \label{equ:cyclic-property-hutchinson}
\end{align}
Each of the annotated terms can be evaluated efficiently for many values of $t$, by precomputing quantities using the Chebyshev reccurrence 
$\mtx{B}(t) = \underline{g}_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A})$, as explained above.

\refalg{alg:nystrom-chebyshev-pp} summarizes the considerations of this section.

\begin{algo}{Chebyshev-Nyström++ method}{nystrom-chebyshev-pp}
\begin{algorithmic}[1]
    \Statex \textbf{Input:} Symmetric $\mtx{A} \in \mathbb{R}^{n \times n}$ with spectrum in $[-1, 1]$, points $\{t_i\}_{i=1}^{n_t} \subset \mathbb{R}$
    \Statex \textbf{Parameters:} Even approximation degree $m \in \mathbb{N}$, number of Girard-Hutchinson queries $n_{\mtx{\Psi}} \in \mathbb{N}_0$, Nyström sketch size $n_{\mtx{\Omega}} \in \mathbb{N}_0$,  smoothing parameter $\sigma > 0$
    \Statex \textbf{Output:} Spectral density approximation $\{\widetilde{\underline{\phi}}_{\sigma}^{(m)}(t_i)\}_{i=1}^{n_t}$
    \For {$i = 1, \dots, n_t$}
        \State Compute coefficients $\{\xi_l(t_i)\}_{l=0}^{m/2}$ of Chebyshev approx.~\refequ{equ:square-root-chebyshev-approximation} to $\sqrt{g_{\sigma}}$ with DCT
        \State Compute coefficients $\{\underline{\mu}_l(t_i)\}_{l=0}^{m}$ of non-negative Chebyshev approx. to $g_{\sigma}$ by applying \refalg{alg:chebyshev-squaring} to $\{\xi_l(t_i)\}_{l=0}^{m/2}$
        \State Compute coefficients $\{\underline{\nu}_l(t_i)\}_{l=0}^{2m}$ of squared non-negative Chebyshev approxim. by applying \refalg{alg:chebyshev-squaring} to $\{\underline{\mu}_l(t_i)\}_{l=0}^{m}$
    \EndFor
    %\State Compute the coefficients $\{\underline{\mu}_l(t_i)\}_{l=0}^{m}$ of the non-negative Chebyshev approximation \refequ{equ:non-negative-chebyshev-approximation} for all points $t_i$ \textcolor{red}{maybe more details needed (maybe compute $\xi$ for m/2, then twice Alg 3.3)}
    %\State Compute the coefficients $\{\nu_l(t_i)\}_{l=0}^{2m}$ of the square of the non-negative Chebyshev approximation with coefficients $\{\underline{\mu}_l(t_i)\}_{l=0}^{m}$ using \refalg{alg:chebyshev-squaring}%\refalg{alg:2-chebyshev-chebyshev-approximation}
    %\State Compute  for all $t_i$ using %\refalg{alg:3-nystrom-chebyshev-exponentiation}
    \State Generate Gaussian random matrices $\mtx{\Omega} \in \mathbb{R}^{n \times n_{\mtx{\Omega}}}$, $\mtx{\Psi} \in \mathbb{R}^{n \times n_{\mtx{\Psi}}}$%\glsfirst{sketching-matrix} $\in \mathbb{R}^{n \times n_{\mtx{\Omega}}}$
    %\State Generate standard Gaussian %\glsfirst{random-matrix} $\in \mathbb{R}^{n \times n_{\mtx{\Psi}}}$
    \State Initialize $[\mtx{V}_1, \mtx{V}_2, \mtx{V}_3] \gets [\mtx{0}_{n \times n_{\mtx{\Omega}}}, \mtx{\mtx{\Omega}}, \mtx{0}_{n \times n_{\mtx{\Omega}}}]$
    \State Initialize $[\mtx{W}_1, \mtx{W}_2, \mtx{W}_3] \gets [\mtx{0}_{n \times n_{\mtx{\Psi}}}, \mtx{\Psi}, \mtx{0}_{n \times n_{\mtx{\Psi}}}]$
    \State Initialize $[\mtx{K}_1(t_i), \mtx{K}_2(t_i)] \gets [\mtx{0}_{n_{\mtx{\Omega}} \times n_{\mtx{\Omega}}}, \mtx{0}_{n_{\mtx{\Omega}} \times n_{\mtx{\Omega}}}]$ for all $t_i$
    \State Initialize $[\mtx{L}_1(t_i), \ell(t_i)] \gets [\mtx{0}_{n_{\mtx{\Omega}} \times n_{\mtx{\Psi}}}, 0]$ for all $t_i$
    %\State Set $\breve{\phi}_{\sigma}^{(m)}(t_i) \gets 0$ for all $t_i$
    \For {$l = 0, \dots, 2m$}
    \State $[\mtx{X}, \mtx{Y}] \gets \mtx{\mtx{\Omega}}^{\top} [\mtx{V}_2, \mtx{W}_2]$  
    %\State $\mtx{X} \gets \mtx{\mtx{\Omega}}^{\top} \mtx{V}_2$
      %\State $\mtx{Y} \gets \mtx{\mtx{\Omega}}^{\top} \mtx{W}_2$
      \State $z \gets \Trace(\mtx{\Psi}^{\top} \mtx{W}_2)$
      \For {$i = 1, \dots, n_t$}
        \If {$l \leq m$}
            \State $\mtx{K}_1(t_i) \gets \mtx{K}_1(t_i) + \underline{\mu}_l(t_i) \mtx{X}$ \Comment{assembles $\mtx{\Omega}^{\top} \underline{g}_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A}) \mtx{\Omega}$}
            \State $\mtx{L}_1(t_i) \gets \mtx{L}_1(t_i) + \underline{\mu}_l(t_i) \mtx{Y}$ \Comment{assembles $\mtx{\Omega}^{\top} \underline{g}_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A}) \mtx{\Psi}$}
            \State $\ell(t_i) \gets \ell(t_i) + \underline{\mu}_l(t_i) z$ \Comment{assembles $\Trace(\mtx{\Psi}^{\top} \underline{g}_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A}) \mtx{\Psi})$}
        \EndIf
        \State $\mtx{K}_2(t_i) \gets \mtx{K}_2(t_i) + \underline{\nu}_l(t_i) \mtx{X}$ \Comment{assembles $\mtx{\Omega}^{\top} \underline{g}_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A})^2 \mtx{\Omega}$}
      \EndFor
      \State $[\mtx{V}_3, \mtx{W}_3] \gets (2 - \delta_{l0}) \mtx{A} [\mtx{V}_2, \mtx{W}_2] - [\mtx{V}_1, \mtx{W}_1]$ \Comment{Chebyshev recurrence}
      \State $[\mtx{V}_1, \mtx{W}_1] \gets [\mtx{V}_2, \mtx{W}_2]$
      \State $[\mtx{V}_2, \mtx{W}_2] \gets [\mtx{V}_3, \mtx{W}_3]$
      %\State $\mtx{V}_3 \gets (2 - \delta_{l0}) \mtx{A} \mtx{V}_2 - \mtx{V}_1$ %\Comment{Chebyshev recurrence \refequ{equ:2-chebyshev-chebyshev-recursion}}
      %\State $\mtx{V}_1 \gets \mtx{V}_2, \mtx{V}_2 \gets \mtx{V}_3$
      %\State $\mtx{W}_3 \gets (2 - \delta_{l0}) \mtx{A} \mtx{W}_2 - \mtx{W}_1$ %\Comment{Chebyshev recurrence \refequ{equ:2-chebyshev-chebyshev-recursion}}
      %\State $\mtx{W}_1 \gets \mtx{W}_2, \mtx{W}_2 \gets \mtx{W}_3$
    \EndFor
    \For {$i = 1, \dots, n_t$}
      \State $\widetilde{\underline{\phi}}_{\sigma}(t_i) \gets \frac{1}{n} \Trace\left( \mtx{K}_1(t_i)^{\dagger}\mtx{K}_2(t_i) \right) + \frac{1}{n n_{\mtx{\Psi}}} \left( \ell(t_i) + \Trace\left( \mtx{L}_1(t_i)^{\top} \mtx{K}_1(t_i)^{\dagger} \mtx{L}_1(t_i) \right)  \right) $ \label{lin:4-nystromchebyshev-nystrom-pp}
    \EndFor
\end{algorithmic}
\end{algo}

\refalg{alg:nystrom-chebyshev-pp} requires $(2m + 1)(n_{\mtx{\Psi}} + n_{\mtx{\Omega}})$ matrix-vector multiplications with $\mtx{A}$, which can be expected to dominate its cost. For each of the evaluation points $t_1, \dots, t_{n_t}$ another $\mathcal{O}(m \log(m) + m (n_{\mtx{\Omega}}^2 + n_{\mtx{\Omega}} n_{\mtx{\Psi}} + 1) + (n_{\mtx{\Omega}}^3 + n_{\mtx{\Omega}} n_{\mtx{\Psi}}^2 + 1))$ operations are needed, which is independent of $n$ -- thanks to the precomputations discussed above.

\refalg{alg:nystrom-chebyshev-pp} also accommodates the other two estimators discussed in \refsec{sec:analysis}: For $n_{\mtx{\Omega}} = 0$ we recover the Girard-Hutchinson estimator~\refequ{equ:hutchinson-trace-estimator}. For $n_{\mtx{\Psi}} = 0$ we recover the Nyström estimator~\refequ{equ:nystrom-trace-estimator}, provided that $0/0$ is evaluated to $0$ in the second term on \reflin{lin:4-nystromchebyshev-nystrom-pp}. In this case, \refalg{alg:nystrom-chebyshev-pp} essentially amounts to the \enquote{spectrum sweeping} method from \cite[Algorithm 5]{lin-2017-randomized-estimation}, featuring a significantly lower computational complexity.

%Chebyshev $\mathcal{O}(n_t (m \log(m) + m (n_{\mtx{\Omega}}^2 + n_{\mtx{\Omega}} n_{\mtx{\Psi}} + 1) + n_{\mtx{\Omega}}^3) + mn^2(n_{\mtx{\Psi}} + n_{\mtx{\Omega}}))$.

%\todo{Complexity in terms of both $n_{\mtx{\Omega}}$ and $n_{\mtx{\Psi}}$.}
%With the cost of a matrix-vector product denoted by $c(n)$, and supposing $n_{\mtx{\Omega}} \approx n_{\mtx{\Psi}}$, we determine the computational complexity of the Chebyshev-Nyström++ method to be $\mathcal{O}(m \log(m) n_t + m n_{\mtx{\Omega}}^2 n + m n_t n_{\mtx{\Omega}}^2 +  m c(n) n_{\mtx{\Omega}} + n_t n_{\mtx{\Omega}}^3)$, with $\mathcal{O}(m n_t + n n_{\mtx{\Omega}} + n_{\mtx{\Omega}}^2 n_t)$ required additional storage.

%\todo{Explain modifications for other estimators, and improved complexity for quadratic trace estimator}

The two pseudoinverses on \reflin{lin:4-nystromchebyshev-nystrom-pp} need to be computed with care. One way to ensure numerical stability is to 
consider an associated eigenvalue problem and threshold the smallest eigenvalues~\cite{lin-2017-randomized-estimation, epperly-2022-theory-quantum}. However, this renders the algorithm significantly more complicated and we found that it improves accuracy only slightly. Instead, we  use a least-squares solver when applying $\mtx{K}_1(t_i)^{\dagger}$ \textcolor{red}{??? Please add more detail here. Which Python function do you use? Is $10^{-5}$ your choice? ???}, which likewise truncates the smallest eigenvalues by treating them as zero if they are smaller than $10^{-5}$ times the largest eigenvalue.

\color{black}

What we have found to be crucial for a successful approximation of the full spectral density was the detection of regions wherein lie no eigenvalues, i.e. regions where the spectral density nearly vanishes. For small smoothing parameters $\sigma$, the rapid decay of the Gaussian $g_{\sigma}$ will cause $\mtx{K}_1(t) \approx \mtx{\Omega}^{\top} g_{\sigma}(t\mtx{I}_n - \mtx{A}) \mtx{\Omega}$ to be close to the zero matrix for parameter values $t$ located in such regions. Hence, on \reflin{lin:4-nystromchebyshev-nystrom-pp} we would compute the pseudoinverse of an almost zero matrix, which can be extremely unstable (see \reffig{fig:zerocheck}). Notice that with $\mtx{K}_1(t)$ --- a matrix we need to assemble anyway --- we can cheaply compute $\Trace(\mtx{K}_1(t)) / n_{\mtx{\Omega}}$, which is the $n_{\mtx{\Omega}}$-query Girard-Hutchinson estimator \refequ{equ:hutchinson-trace-estimator}. If this estimate is significantly smaller than $1/n$, in our experiments usually $10^{-5}$, we may use it as an indicator that $t$ lies within a region where the spectral density nearly vanishes, and therefore set $\widetilde{\underline{\phi}}_{\sigma}^{(m)}(t) = 0$ instead of computing the full expression on \reflin{lin:4-nystromchebyshev-nystrom-pp}.

\begin{figure}[ht]
    \centering
    \input{plots/zerocheck.pgf}
    \caption{The difference the non-zero check can make when approximating a spectral density using \refalg{alg:nystrom-chebyshev-pp}. Here, we use a coarse discretization of the Hamiltonian \refequ{equ:5-experiments-electronic-hamiltonian} in three dimensions and ran the Chebyshev-Nyström++ method with and without estimating if the matrix function is zero before taking its pseudo-inverse. We let $m=2000$, $n_{\mtx{\Omega}}=80$, $n_{\mtx{\Psi}}=0$, and $\sigma = 0.004$. We compute the $L^1$ error from the approximation of the spectral density at $n_t = 100$ uniformly spaced values of $t$ in the interval $[-1, 1]$.}
    \label{fig:zerocheck}
\end{figure}

\color{black}

\subsubsection{Analysis}
\label{subsubsec:chebyshev-nystrom-analysis}

\color{blue}

In this section, we analyze the accuracy of Chebyshev-Nyström++ estimate returned by~\refalg{alg:nystrom-chebyshev-pp}. 
\begin{theorem}{Error of Chebyshev-Nyström++ estimate}{chebyshev-nystrom}
    For a smoothing parameter $0 < \sigma \ll 1$ \textcolor{red}{??? Why do we need $\ll$ here? This is quite confusing because this is not a mathematical statement. Can this be simply replaced by $<$ or not? ???}, let $\phi_{\sigma}$ be the smoothed spectral density~\eqref{equ:smooth-spectral-density} of a symmetric matrix $\mtx{A} \in \mathbb{R}^{n \times n}$ with spectrum contained in $[-1, 1]$. Consider $\varepsilon \in (0, 3)$ and $\delta \in (0, 1/2)$.
    Then the Chebyshev-Nyström++ estimate $\widetilde{\underline{\phi}}_{\sigma}^{(m)}$ 
    satisfies the error bound
    \begin{equation}
        \int_{-1}^{1} \left| \phi_{\sigma}(t) - \widetilde{\underline{\phi}}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t \leq \varepsilon
        \label{equ:chebyshev-nystrom-error}
        %\lVert  \phi_{\sigma} - \phi_{\sigma}^{(m)} \rVert _1 &\leq \frac{2\sqrt{2}}{\sigma^2} (1 + \sigma)^{-m}.
        %\label{equ:2-chebyshev-interpolation-error}
    \end{equation}
    with probability at least $1 - \delta$ for $m = \mathcal{O}(\sigma^{-1}(\log(\varepsilon^{-1}) + \log(\sigma^{-1})))$  and $n_{\mtx{\Omega}} = n_{\mtx{\Psi}} = \mathcal{O}(\varepsilon^{-1}\log(\delta^{-1})^2)$.
\end{theorem}

\begin{proof}
Combining results derived above, we obtain that \color{black}
    \begin{align}
        &\int_{-1}^{1} \left| \phi_{\sigma}(t) - \widetilde{\underline{\phi}}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t \notag \\
        &\leq \int_{-1}^{1} \left| \phi_{\sigma}(t) - \underline{\phi}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t + \int_{-1}^{1} \left| \underline{\phi}_{\sigma}^{(m)}(t) - \widetilde{\underline{\phi}}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t &&  \notag \\
        &\leq \int_{-1}^{1} \left| \phi_{\sigma}(t) - \underline{\phi}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t + \widetilde{\varepsilon} \int_{-1}^{1} \left| \underline{\phi}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t && \text{(\refthm{thm:nystrom-pp} w.p. $\geq 1 - \delta$)} \notag \\
        &\leq (1 + \widetilde{\varepsilon}) \int_{-1}^{1} \left| \phi_{\sigma}(t) - \underline{\phi}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t + \widetilde{\varepsilon} \int_{-1}^{1} \left| \phi_{\sigma}(t) \right|~\mathrm{d}t &&  \notag \\
        &\leq (1 + \widetilde{\varepsilon}) \underline{E}_{\sigma, m} + \widetilde{\varepsilon}. && \text{(\reflem{lem:non-negative-chebyshev-error})} \notag 
    \end{align}
    From \reflem{lem:non-negative-chebyshev-error} we see that if \textcolor{red}{$0 < \sigma \ll 1$ ??? Why $\ll$? ???}, choosing $m = \mathcal{O}(\sigma^{-1}(\log(\widetilde{\varepsilon}^{-1}) + \log(\sigma^{-1})))$ ensures $\underline{E}_{\sigma, m} \leq \widetilde{\varepsilon}$, and thus, if $\widetilde{\varepsilon} \leq 1$,
    \begin{equation}
        \int_{-1}^{1} \left| \phi_{\sigma}(t) - \underline{\phi}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t \leq (1+\widetilde{\varepsilon}) \widetilde{\varepsilon} + \widetilde{\varepsilon} \leq 3 \widetilde{\varepsilon}
    \end{equation}
    from which the result follows by identifying $\varepsilon = 3 \widetilde{\varepsilon}$.
\end{proof}

%\todo{Maybe specialize \refthm{thm:nystrom} to spectral densities?}

%We now discuss what value to choose for $n_{\mtx{\Omega}}$ in the Nyström trace estimator introduced in \refsec{subsec:nystrom}. In order to guarantee an even better approximation error based on \refthm{thm:nystrom}, we analyze the numerical rank of the matrix function $g_{\sigma}(t\mtx{I}_n - \mtx{A})$.

In practice, the approximation error is often significantly better than guaranteed in \refthm{thm:chebyshev-nystrom}, particularly when $n_{\mtx{\Omega}}$ is chosen rather large. The reason is that for small $\sigma$ the singular values of the matrix function $g_{\sigma}(t\mtx{I}_n - \mtx{A})$ decay so quickly that already the Nyström approximation alone is highly accurate (cf. \refthm{thm:nystrom}). To establish a relation between the singular value decay of $g_{\sigma}(t\mtx{I}_n - \mtx{A})$ and what choice of $n_{\mtx{\Omega}}$ is needed for a significantly better approximation than ensured by \refthm{thm:chebyshev-nystrom}, we study the numerical rank of $g_{\sigma}(t\mtx{I}_n - \mtx{A}) / n$.\footnote{The factor $1/n$
does not change the conclusion, but is added to simplify the calculations and also reflects how we implement the method: We absorb the factor $1/n$ from the smoothed spectral density \refequ{equ:smooth-spectral-density} in the definition of $g_{\sigma}$, such that in reality, the Chebyshev approximation and the Nyström++ estimator are applied to $g_{\sigma}(t\mtx{I}_n - \mtx{A}) / n$.}

We can express the numerical rank of a matrix with respect to the nuclear norm in terms of its singular values $\sigma_1 \geq \dots \geq \sigma_n \geq 0$ as
\begin{equation}
    r_{\varepsilon_m} = \min \{1 \leq r \leq n: \sum_{j=r+1}^n \sigma_{j} \leq \varepsilon_m \},
    \label{equ:numerical-rank}
\end{equation}
where $\varepsilon_m$ is usually taken to be the double machine precision, i.e. $\varepsilon_m \approx 10^{-16}$.

%\begin{theorem}{Nyström estimator for parameter-dependent matrices}%{chebyshev-nystrom-raw}
%    Let $\phi_{\sigma}$ be the smooth spectral density of a symmetric %matrix $\mtx{A} \in \mathbb{R}^{n \times n}$ whose spectrum is %contained in $[-1, 1]$. Then the Chebyshev-Nyström++ estimate %$\widetilde{\phi}_{\sigma}^{(m)}$ with \emph{even} degree $m \in %\mathbb{N}$, $n_{\mtx{\Omega}} \geq \max(6, \max_{t \in [-1, 1]}(r_%{\varepsilon}(g_{\sigma}(t\mtx{I}_n - \mtx{A}))) + \log(\delta^%{-1}))$, $n_{\mtx{\Psi}} = 0$, and smoothing parameter $\sigma > 0$ %satisfies with for all $\gamma \geq 1$ with probability at least $1 %- \delta$
%    \begin{equation}
%        \int_{a}^{b} \left| \phi_{\sigma}(t) - \widetilde{\phi}_{\sigma}%(t) \right| ~\mathrm{d}t
%        \leq 2 e^2 (1 + r_{\varepsilon}) (\varepsilon + (n - r_%{\varepsilon}) \underline{E}_{\sigma, m}) + \underline{E}_%{\sigma, m}.
%    \end{equation}
%\end{theorem}
%
%\begin{proof}
%    \todo{TODO}
%\end{proof}

When using a Gaussian smoothing-kernel, we can explicitly write the singular values of $g_{\sigma}(t\mtx{I}_n - \mtx{A}) / n$ as
\begin{equation}
    \sigma_i(t) = g_{\sigma}(t - \lambda_{(i)}) / n = \frac{1}{n \sqrt{2 \pi \sigma^2}} e^{-\sfrac{\left(t - \lambda_{(i)}\right)^2}{2 \sigma^2}},
    \label{equ:gaussian-kernel-eigenvalues}
\end{equation}
where $\lambda_{(1)}, \dots, \lambda_{(n)}$ denote the eigenvalues of $\mtx{A}$ sorted by increasing distance from spectral-parameter $t$, such that $\sigma_1(t) \geq \dots \geq \sigma_n(t)$. Consequently, by using the closed-form expression of the eigenvalues \refequ{equ:gaussian-kernel-eigenvalues}, we may generously upper bound the numerical rank by
\begin{equation}
    r_{\varepsilon_m}(g_{\sigma}(t\mtx{I}_n - \mtx{A}) / n) \leq \#\left\{1\leq i\leq n: |t - \lambda_i| < d_{\varepsilon_m, \sigma} \right\},
    \label{equ:gaussian-kernel-numerical-rank}
\end{equation}
with $d_{\varepsilon_m, \sigma} = \sigma \sqrt{-2 \log(\sqrt{2 \pi} \sigma \varepsilon_m)}$ and the eigenvalues $\lambda_1, \dots, \lambda_n$ of $\mtx{A}$ and the constants. The expression \refequ{equ:gaussian-kernel-numerical-rank} has a nice visual interpretation: The numerical rank of a matrix is at most equal to the number of eigenvalues which are closer to the parameter $t$ than $d_{\varepsilon_m, \sigma}$. This is illustrated in \reffig{fig:numerical-rank}.
\begin{figure}[ht]
    \centering
    \input{figures/numerical-rank.tex}
    \caption{The numerical rank of $g_{\sigma}(t\mtx{I}_n - \mtx{A}) / n$ can be approximately computed by counting the number of eigenvalues $\lambda_{(1)}, \dots, \lambda_{(n)}$ of the matrix $\mtx{A}$ which lie less than a constant $d_{\varepsilon_m, \sigma}$ away from $t$.}
    \label{fig:numerical-rank}
\end{figure}

If we additionally assume the eigenvalues of the matrix $\mtx{A}$ to be evenly distributed within $[-1, 1]$, that is, in any subinterval of fixed length in $[a, b]$, $-1 \leq a < b \leq 1$, we can expect to find roughly the same number of eigenvalues, then we can estimate the numerical rank of $g_{\sigma}(t\mtx{I}_n - \mtx{A}) / n$ to be
\begin{equation}
    r_{\varepsilon_m}(g_{\sigma}(t\mtx{I}_n - \mtx{A}) / n) \lessapprox n d_{\varepsilon_m, \sigma} \approx 6 n \sigma,
    \label{equ:gaussian-kernel-numerical-rank-uniform}
\end{equation}
provided we set $\varepsilon_m = 10^{-16}$ and choose $\sigma$ roughly between $10^{-8}$ and $1$. Hence, choosing $n_{\mtx{\Omega}} = \mathcal{O}(n \sigma)$ will yield approximations which can be significantly better than \refthm{thm:chebyshev-nystrom}.