### 2025-02-21 paper/re-pro-badge.tex
automatically reproduce paper

The code for our numerical experiments can be found on \url{https://github.com/FMatti/Rand-TRACE}. The figures and tables were generated from the commit \href{https://github.com/FMatti/Rand-TRACE/tree/37ec0b7}{37ec0b7} on 2025-02-21 at 17:02:20 UTC.