### 2025-02-21 paper/figures/numerical-rank.tex
rework numerical results section and numerical rank explanation

\begin{tikzpicture}
    \fill[black!15!white] (-2.0, -0.15) rectangle (2.0, 0.15);
    \draw[thick, ->] (-5, 0) to (5, 0);
    \fill[black] (-4, 0) circle (0.075) node[below] {$\lambda_{(7)}$};
    \fill[black] (-3, 0) circle (0.075) node[below] {$\lambda_{(5)}$};
    \fill[black] (-2.35, 0) circle (0.075) node[below] {$\lambda_{(4)}$};
    \fill[black] (-0.8, 0) circle (0.075) node[below] {$\lambda_{(2)}$};
    \fill[black] (0.3, 0) circle (0.075) node[below] {$\lambda_{(1)}$};
    \fill[black] (1, 0) circle (0.075) node[below] {$\lambda_{(3)}$};
    \fill[black] (3, 0) circle (0.075) node[below] {$\lambda_{(6)}$};
    \fill[black] (4.2, 0) circle (0.075) node[below] {$\lambda_{(8)}$};
    \fill[black] (4.8, 0) circle (0.075) node[below] {$\lambda_{(9)}$};
    \fill[black] (-4, 0) circle (0.075) node[above] {$\sigma_{7}$};
    \fill[black] (-3, 0) circle (0.075) node[above] {$\sigma_{5}$};
    \fill[black] (-2.35, 0) circle (0.075) node[above] {$\sigma_{4}$};
    \fill[black] (-0.8, 1.3) circle (0.075) node[above left] {$\sigma_{2}$};
    \fill[black] (0.3, 2.28) circle (0.075) node[above right] {$\sigma_{1}$};
    \fill[black] (1, 0.91) circle (0.075) node[above right] {$\sigma_{3}$};
    \fill[black] (3, 0) circle (0.075) node[above] {$\sigma_{6}$};
    \fill[black] (4.2, 0) circle (0.075) node[above] {$\sigma_{8}$};
    \fill[black] (4.8, 0) circle (0.075) node[above] {$\sigma_{9}$};
    \draw[black, ultra thick] (-0.05, 0.15) to (-0.05, -0.55) node[below] {$t$};
    \draw[black, thick] (-2.0, 0.15) to (-2.0, -0.55) node[below] {$t - d$};
    \draw[black, thick] (2.0, 0.15) to (2.0, -0.55) node[below] {$t + d$};
    \draw[black, thick, dashed] (-0.8, 0) to (-0.8, 1.3);
    \draw[black, thick, dashed] (0.3, 0) to (0.3, 2.28);
    \draw[black, thick, dashed] (1, 0) to (1, 0.91);
    \draw[black!60!white, thick] (-5.000, 0.000) to (-4.899, 0.000) to (-4.798, 0.000) to (-4.697, 0.000) to (-4.596, 0.000) to (-4.495, 0.000) to (-4.394, 0.000) to (-4.293, 0.000) to (-4.192, 0.000) to (-4.091, 0.000) to (-3.990, 0.000) to (-3.889, 0.000) to (-3.788, 0.000) to (-3.687, 0.000) to (-3.586, 0.000) to (-3.485, 0.000) to (-3.384, 0.000) to (-3.283, 0.000) to (-3.182, 0.000) to (-3.081, 0.000) to (-2.980, 0.000) to (-2.879, 0.001) to (-2.778, 0.001) to (-2.677, 0.002) to (-2.576, 0.003) to (-2.475, 0.005) to (-2.374, 0.009) to (-2.273, 0.014) to (-2.172, 0.022) to (-2.071, 0.034) to (-1.970, 0.052) to (-1.869, 0.076) to (-1.768, 0.110) to (-1.667, 0.155) to (-1.566, 0.215) to (-1.465, 0.293) to (-1.364, 0.389) to (-1.263, 0.508) to (-1.162, 0.649) to (-1.061, 0.812) to (-0.960, 0.995) to (-0.859, 1.196) to (-0.758, 1.408) to (-0.657, 1.625) to (-0.556, 1.836) to (-0.455, 2.033) to (-0.354, 2.206) to (-0.253, 2.346) to (-0.152, 2.443) to (-0.051, 2.494) to (0.051, 2.494) to (0.152, 2.443) to (0.253, 2.346) to (0.354, 2.206) to (0.455, 2.033) to (0.556, 1.836) to (0.657, 1.625) to (0.758, 1.408) to (0.859, 1.196) to (0.960, 0.995) to (1.061, 0.812) to (1.162, 0.649) to (1.263, 0.508) to (1.364, 0.389) to (1.465, 0.293) to (1.566, 0.215) to (1.667, 0.155) to (1.768, 0.110) to (1.869, 0.076) to (1.970, 0.052) to (2.071, 0.034) to (2.172, 0.022) to (2.273, 0.014) to (2.374, 0.009) to (2.475, 0.005) to (2.576, 0.003) to (2.677, 0.002) to (2.778, 0.001) to (2.879, 0.001) to (2.980, 0.000) to (3.081, 0.000) to (3.182, 0.000) to (3.283, 0.000) to (3.384, 0.000) to (3.485, 0.000) to (3.586, 0.000) to (3.687, 0.000) to (3.788, 0.000) to (3.889, 0.000) to (3.990, 0.000) to (4.091, 0.000) to (4.192, 0.000) to (4.293, 0.000) to (4.394, 0.000) to (4.495, 0.000) to (4.596, 0.000) to (4.697, 0.000) to (4.798, 0.000) to (4.899, 0.000) to (5.000, 0.000);
    \node[black!60!white] at (-1.3, 2.3) {$g_{\sigma}(t - \cdot)$};
\end{tikzpicture}