### 2025-02-25 paper/re-pro-badge.tex
automatically reproduce paper

The code for our numerical experiments can be found on \url{https://github.com/FMatti/parameter-trace}. The figures and tables were generated from the commit \href{https://github.com/FMatti/parameter-trace/tree/334d5c6}{334d5c6} on 2025-02-25 at 19:12:59 UTC.